### 2021-07-21 paper/paper.tex
add paper template
Co-authored-by: Michael Schlottke-Lakemper <michael@sloede.com>

% JuliaCon proceedings template
\documentclass{juliacon}
\setcounter{page}{1}

\begin{document}

\input{header}

\maketitle

\begin{abstract}
TODO: abstract %TODO
\end{abstract}

\section{Introduction}

TODO %TODO
\begin{itemize}
  \item Basic design of Trixi.jl and its capabilities
  \item PID comparison of Fluxo and Trixi.jl
  \item Discussion on Julia for numerical simulation science
  \item Prepare repro repo including scripts for Trixi.jl and Fluxo
  \item Cite \cite{bezanson2017julia,schlottkelakemper2021purely,rackauckas2017differentialequations,revels2016forward}
\end{itemize}

\clearpage
\section*{Some guidelines}
\label{sec:some_guide}

The following notes may help you achieve the best effects with the
\verb juliacon  class file.

\subsection{Writing Julia code}

A special environment is already defined for Julia code,
built on top of \textit{listings} and \textit{jlcode}.

\begin{verbatim}
\begin{lstlisting}[language = Julia]
using Plots

x = -3.0:0.01:3.0
y = rand(length(x))
plot(x, y)
\end{lstlisting}
\end{verbatim}
\begin{lstlisting}[language = Julia]
using Plots

x = -3.0:0.01:3.0
y = rand(length(x))
plot(x, y)
\end{lstlisting}

\subsection{Double Column Figure and Tables}
\label{subsub:double_fig_tab}
For generating the output of figures and tables in double column
we can use the following coding:

\begin{enumerate}
\item For Figures:
\begin{verbatim}
\begin{figure*}...\end{figure*}
\end{verbatim}
\item For landscape figures:
\begin{verbatim}
\begin{sidewaysfigure*}...\end{sidewaysfigure*}
\end{verbatim}
\item For Tables:
\begin{verbatim}
\begin{table*}...\end{table*}
\end{verbatim}
\item For landscape tables:
\begin{verbatim}
\begin{sidewaystable*}...\end{sidewaystable*}
\end{verbatim}
\end{enumerate}

\subsection{Enunciations}
\label{subsub:enunciation}

The \verb juliacon   class file generates the enunciations with the help of
the following commands:
\begin{verbatim}
\begin{theorem}...\end{theorem}
\begin{strategy}...\end{strategy}
\begin{property}...\end{property}
\begin{proposition}...\end{proposition}
\begin{lemma}...\end{lemma}
\begin{example}...\end{example}
\begin{proof}...\end{proof}
\begin{definition}...\end{definition}
\begin{algorithm}...\end{algorithm}
\begin{remark}...\end{remark}
\end{verbatim}
The above-mentioned coding can also include optional arguments
such as
\begin{verbatim}
\begin{theorem}[...]. Example for theorem:
\begin{theorem}[Generalized Poincare Conjecture]
Four score and seven ... created equal.
\end{theorem}
\end{verbatim}

\begin{theorem}[Generalized Poincare Conjecture]
Four score and seven years ago our fathers brought forth,
upon this continent, a new nation, conceived in Liberty,
 and dedicated to the proposition that all men are
created equal.
\end{theorem}

\subsection{Balancing column at last page}
\label{subsub:Balance}
For balancing the both column length at last page use:
\begin{verbatim}
\vadjust{\vfill\pagebreak}
\end{verbatim}

%\vadjust{\vfill\pagebreak}

at appropriate place in your \TeX{} file or in bibliography file.

\subsection{Handling references}
\label{subsub:references}
References are most easily (and correctly) generated using the
BIBTEX, which is easily invoked via
\begin{verbatim}
\bibliographystyle{juliacon}
\bibliography{ref}
\end{verbatim}
When submitting the document source (.tex) file to external
parties, the ref.bib file should be sent with it.
Don't forget to cite \cite{bezanson2017julia}.

% References
\input{bib.tex}

\end{document}

% Inspired by the International Journal of Computer Applications template

### 2021-07-21 paper/paper.tex
add draft of PID figures for Euler

@@
 \setcounter{page}{1}
+
+% Use multiple images in a single environment
+\usepackage{subcaption}
+
+% Plotting in tikz images
+\usepackage{pgfplots}
+% Silence warnings
+\pgfplotsset{compat=1.16}
+% Allow external generation of images to speed up compilation
+% \usetikzlibrary{external}
+% \tikzexternalize[prefix=../figures/tikz/]
 
@@
 \begin{abstract}
-TODO: abstract %TODO
+TODO: abstract (at most 600 characters, written in plain English with no symbol
+nor formula + references etc.) %TODO
 \end{abstract}
@@
   \item Cite \cite{bezanson2017julia,schlottkelakemper2021purely,rackauckas2017differentialequations,revels2016forward}
+  \item Change title to something like ``Adaptive high-order numerical simulations
+  of hyperbolic PDEs with Trixi.jl. A case study of Julia for scientific computing''?
 \end{itemize}
+
+\section{Performance comparison with FLUXO}
+
+FLUXO\footnote{\url{https://gitlab.com/project-fluxo/fluxo}} is an open source
+MPI Fortran code implementing discontinuous Galerkin methods on unstructured hex
+meshes in three space dimensions for advection diffusion equations. It also
+implements the same kind of modern DG methods based on flux differencing
+to achieve entropy conservation/dissipation or kinetic energy preservation.
+Thus, both Trixi.jl and FLUXO share a common set of features, enabling a reasonable
+comparison of their performance.
+
+Here, we compare the serial performance of Trixi.jl and FLUXO when solving a
+hyperbolic PDE in three space dimensions on curvilinear hex meshes.
+
+TODO: Andrew, please describe the setups that you used.
+
+\begin{figure}
+\centering
+  \begin{subfigure}{\linewidth}
+    \begin{tikzpicture}[
+      font=\footnotesize
+      ]
+      \begin{axis}[
+          xlabel={Polynomial degree},
+          % ylabel={Time per RHS call and DOF (seconds)},
+          ylabel={Time/RHS/DOF {[sec]}},
+          width=\textwidth,
+          legend columns=-1,
+          % legend to name=place_legend_here,
+          legend pos=north west,
+          ymin=0.0,
+          ymax=2.2e-6,
+          grid=major,
+        ]
+        % \addplot[blue, mark=square*, solid]
+        %   table [x index=0, y index=1]{../data/pids_euler_weak.dat};
+        %   \addlegendentry{\texttt{TreeMesh} (Trixi.jl)}
+        \addplot[blue, mark=*, solid]
+          table [x index=0, y index=3]{../data/pids_euler_weak.dat};
+          \addlegendentry{\footnotesize FLUXO}
+        \addplot[red, mark=diamond*, solid]
+          table [x index=0, y index=2]{../data/pids_euler_weak.dat};
+          % \addlegendentry{\texttt{StructuredMesh} (Trixi.jl)}
+          \addlegendentry{\footnotesize Trixi.jl}
+        \node at (axis cs:13,3.5e-7) {Weak form};
+
+        % \addplot[blue, mark=square*, dashed]
+        %   table [x index=0, y index=1]{../data/pids_euler_ranocha.dat};
+        \addplot[blue, mark=*, dashed]
+          table [x index=0, y index=3]{../data/pids_euler_ranocha.dat};
+        \addplot[red, mark=diamond*, dashed]
+          table [x index=0, y index=2]{../data/pids_euler_ranocha.dat};
+        \node [rotate=23] at (axis cs:9,8.2e-7) {Flux differencing};
+        \draw (axis cs:9,9.2e-7) -- (axis cs:7.5,10.2e-7);
+        \draw (axis cs:9,7.2e-7) -- (axis cs:8.5,5.2e-7);
+      \end{axis}
+    \end{tikzpicture}%
+    \caption{Absolute run times.}
+  \end{subfigure}%
+  \\
+  \begin{subfigure}{\linewidth}
+    \begin{tikzpicture}[
+      font=\footnotesize
+      ]
+      \begin{axis}[
+          xlabel={Polynomial degree},
+          % ylabel={Time per RHS call and DOF (relative)},
+          ylabel={Time relative to FLUXO {[-]}},
+          width=\textwidth,
+          ymin=0.0,
+          ymax=1.1,
+          ytick={0,0.25,0.5,0.75,1},
+          grid=major,
+        ]
+        % \addplot[blue, mark=square*, solid]
+        %   table [x index=0, y index=1]{../data/pids_euler_weak_relative.dat};
+        \addplot[red, mark=diamond*, solid]
+          table [x index=0, y index=2]{../data/pids_euler_weak_relative.dat};
+        \addplot[blue, mark=*, solid]
+          table [x index=0, y index=3]{../data/pids_euler_weak_relative.dat};
+
+        % \addplot[blue, mark=square*, dashed]
+        %   table [x index=0, y index=1]{../data/pids_euler_ranocha_relative.dat};
+        \addplot[red, mark=diamond*, dashed]
+          table [x index=0, y index=2]{../data/pids_euler_ranocha_relative.dat};
+        \addplot[blue, mark=*, dashed]
+          table [x index=0, y index=3]{../data/pids_euler_ranocha_relative.dat};
+
+        % \node at (axis cs:10,0.6) {Will be improved soon};
+      \end{axis}
+    \end{tikzpicture}
+    \caption{Run time relative to FLUXO.}
+  \end{subfigure}%
+  \caption{Run time per right-hand side evaluation and degree of freedom for
+           different DG discretizations of the 3D compressible Euler equations.}
+\end{figure}
 
@@
 \label{sec:some_guide}
+
+Full paper: 5--10 pages.
+\begin{quote}
+  Compared to an extended abstract, a full paper presents more of the background
+  and context motivating the work. It compares the work to other approaches taken
+  in the field and gives some additional insights on the conference contribution.
+  Use cases back up the work by showing how it can be used.
+\end{quote}
 

### 2021-07-21 paper/paper.tex
first brief description of PID results

@@
 TODO: Andrew, please describe the setups that you used.
+
+The results of this comparison are visualized in Figure~\ref{fig:PID-Euler}.
+Clearly, Trixi.jl is a bit more than 2x faster than the Fortran code FLUXO
+while both implement the same numerical methods. This demonstrates the
+suitability of Julia for this kind of simulation-focused scientific computing.
+
+Note that we do not want to state that Julia is generically faster than Fortran,
+C, or C++. In contrast, we would like to emphasize that Julia can be at least
+as fast as code written in these traditional scientific computing languages.
+We believe that it will be possible to get the same efficiency by rewriting
+FLUXO appropriately. However, the code introspection and profiling tools available
+in Julia make it easier for us to optimize Julia code compared to other languages
+we are familiar with (having a diverse team with a strong background in C, C++,
+and Fortran).
 
@@
            different DG discretizations of the 3D compressible Euler equations.}
+  \label{fig:PID-Euler}
 \end{figure}

### 2021-07-21 paper/paper.tex
add MPI.jl reference

@@
 
+\section{Design of Trixi.jl and feature overview}
+
+TODO %TODO
+
 \section{Performance comparison with FLUXO}
@@
 
-TODO: Andrew, please describe the setups that you used.
+TODO: Andrew, please describe the setups that you used. %TODO
 
@@
 \end{figure}
+
+\section{Assessment of Julia for simulation-focused scientific computing}
+
+TODO %TODO
+\begin{itemize}
+  \item Some nice features
+  \item Compilation time can be annoying but also allows cool features and performance
+  \item Extendability: Multi-physics with 350 lines of code \cite{schlottkelakemper2021purely}
+  \item MPI: To be evaluated but the basic tools are already there \cite{byrne2021mpi}
+\end{itemize}
+
+\section{Summary and conclusions}
+
+TODO %TODO
+
+\section*{Acknowledgments}
+
+Funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation)
+under Germany's Excellence Strategy EXC 2044-390685587, Mathematics Münster:
+Dynamics-Geometry-Structure.
+
+TODO %TODO
 

### 2021-07-21 paper/paper.tex
additional TODO notes

@@
 
-\section{Introduction}
+\section*{TODO}
 
-TODO %TODO
 \begin{itemize}
@@
   \item Prepare repro repo including scripts for Trixi.jl and Fluxo
-  \item Cite \cite{bezanson2017julia,schlottkelakemper2021purely,rackauckas2017differentialequations,revels2016forward}
+  \item Cite \cite{bezanson2017julia,schlottkelakemper2021purely,rackauckas2017differentialequations,revels2016forward,ranocha2021sbp}
   \item Change title to something like ``Adaptive high-order numerical simulations
-  of hyperbolic PDEs with Trixi.jl. A case study of Julia for scientific computing''?
+  of hyperbolic PDEs with Trixi.jl: A case study of Julia for scientific computing''?
+  \item Restructure non-conservative stuff and add MHD benchmarks?
 \end{itemize}
 
-\section{Design of Trixi.jl and feature overview}
+\section{Introduction}
+
+TODO %TODO
+
+\section{Design of Trixi.jl and overview of its features}
 
@@
 
-TODO %TODO
+TODO: Add your funding sources (if any) %TODO
 
@@
 
-%\vadjust{\vfill\pagebreak}
+% \vadjust{\vfill\pagebreak}
 
 at appropriate place in your \TeX{} file or in bibliography file.
-
-\subsection{Handling references}
-\label{subsub:references}
-References are most easily (and correctly) generated using the
-BIBTEX, which is easily invoked via
-\begin{verbatim}
-\bibliographystyle{juliacon}
-\bibliography{ref}
-\end{verbatim}
-When submitting the document source (.tex) file to external
-parties, the ref.bib file should be sent with it.
-Don't forget to cite \cite{bezanson2017julia}.
 

### 2021-07-21 paper/paper.tex
tweak PID figure slightly

@@
         ]
+        \addplot[blue, mark=*, solid]
+          table [x index=0, y index=3]{../data/pids_euler_weak.dat};
+          \addlegendentry{FLUXO}
         % \addplot[blue, mark=square*, solid]
@@
         %   \addlegendentry{\texttt{TreeMesh} (Trixi.jl)}
-        \addplot[blue, mark=*, solid]
-          table [x index=0, y index=3]{../data/pids_euler_weak.dat};
-          \addlegendentry{\footnotesize FLUXO}
         \addplot[red, mark=diamond*, solid]
@@
           % \addlegendentry{\texttt{StructuredMesh} (Trixi.jl)}
-          \addlegendentry{\footnotesize Trixi.jl}
+          \addlegendentry{Trixi.jl}
         \node at (axis cs:13,3.5e-7) {Weak form};
 
+        \addplot[blue, mark=*, dashed]
+          table [x index=0, y index=3]{../data/pids_euler_ranocha.dat};
         % \addplot[blue, mark=square*, dashed]
         %   table [x index=0, y index=1]{../data/pids_euler_ranocha.dat};
-        \addplot[blue, mark=*, dashed]
-          table [x index=0, y index=3]{../data/pids_euler_ranocha.dat};
         \addplot[red, mark=diamond*, dashed]
@@
           width=\textwidth,
+          legend columns=-1,
+          legend pos=south west,
           ymin=0.0,
@@
         ]
+        \addplot[blue, mark=*, solid, forget plot]
+          table [x index=0, y index=3]{../data/pids_euler_weak_relative.dat};
         % \addplot[blue, mark=square*, solid]
@@
           table [x index=0, y index=2]{../data/pids_euler_weak_relative.dat};
-        \addplot[blue, mark=*, solid]
-          table [x index=0, y index=3]{../data/pids_euler_weak_relative.dat};
+          \addlegendentry{Weak form}
 
+        \addplot[blue, mark=*, dashed, forget plot]
+          table [x index=0, y index=3]{../data/pids_euler_ranocha_relative.dat};
         % \addplot[blue, mark=square*, dashed]
@@
           table [x index=0, y index=2]{../data/pids_euler_ranocha_relative.dat};
-        \addplot[blue, mark=*, dashed]
-          table [x index=0, y index=3]{../data/pids_euler_ranocha_relative.dat};
-
-        % \node at (axis cs:10,0.6) {Will be improved soon};
+          \addlegendentry{Flux differencing}
       \end{axis}

### 2021-07-21 paper/paper.tex
mention TreeMesh

@@
 \setcounter{page}{1}
+
+% Physicists rule
+\usepackage{siunitx}
+% \sisetup{output-exponent-marker=\text{e}}
+% \sisetup{tight-spacing=true}
 
@@
 The results of this comparison are visualized in Figure~\ref{fig:PID-Euler}.
-Clearly, Trixi.jl is a bit more than 2x faster than the Fortran code FLUXO
+Clearly, Trixi.jl is more than 2x faster than the Fortran code FLUXO
 while both implement the same numerical methods. This demonstrates the
 suitability of Julia for this kind of simulation-focused scientific computing.
-
-Note that we do not want to state that Julia is generically faster than Fortran,
-C, or C++. In contrast, we would like to emphasize that Julia can be at least
-as fast as code written in these traditional scientific computing languages.
-We believe that it will be possible to get the same efficiency by rewriting
-FLUXO appropriately. However, the code introspection and profiling tools available
-in Julia make it easier for us to optimize Julia code compared to other languages
-we are familiar with (having a diverse team with a strong background in C, C++,
-and Fortran).
 
@@
           width=\textwidth,
+          height=0.66\textwidth,
           legend columns=-1,
@@
         ]
-        \addplot[blue, mark=*, solid]
+        \addplot[blue, mark=*, solid] % FLUXO, weak form
           table [x index=0, y index=3]{../data/pids_euler_weak.dat};
           \addlegendentry{FLUXO}
-        % \addplot[blue, mark=square*, solid]
+%         \addplot[orange, mark=square*, solid] % TreeMesh, weak form
 %           table [x index=0, y index=1]{../data/pids_euler_weak.dat};
 %           \addlegendentry{\texttt{TreeMesh} (Trixi.jl)}
-        \addplot[red, mark=diamond*, solid]
+        \addplot[red, mark=diamond*, solid] % StructuredMesh, weak form
           table [x index=0, y index=2]{../data/pids_euler_weak.dat};
@@
 
-        \addplot[blue, mark=*, dashed]
+        \addplot[blue, mark=*, dashed] % FLUXO, flux differencing
           table [x index=0, y index=3]{../data/pids_euler_ranocha.dat};
-        % \addplot[blue, mark=square*, dashed]
+%         \addplot[orange, mark=square*, dashed] % TreeMesh, flux differencing
 %           table [x index=0, y index=1]{../data/pids_euler_ranocha.dat};
-        \addplot[red, mark=diamond*, dashed]
+        \addplot[red, mark=diamond*, dashed] % StructuredMesh, flux differencing
           table [x index=0, y index=2]{../data/pids_euler_ranocha.dat};
@@
           ymax=1.1,
-          ytick={0,0.25,0.5,0.75,1},
+          ytick={0, 0.2, 0.4, 0.5, 0.6, 0.8, 1},
           grid=major,
         ]
-        \addplot[blue, mark=*, solid, forget plot]
+        \addplot[blue, mark=*, solid, forget plot] % FLUXO, weak form
           table [x index=0, y index=3]{../data/pids_euler_weak_relative.dat};
-        % \addplot[blue, mark=square*, solid]
+%         \addplot[orange, mark=square*, solid, forget plot] % TreeMesh, weak form
 %           table [x index=0, y index=1]{../data/pids_euler_weak_relative.dat};
-        \addplot[red, mark=diamond*, solid]
+        \addplot[red, mark=diamond*, solid] % StructuredMesh, weak form
           table [x index=0, y index=2]{../data/pids_euler_weak_relative.dat};
@@
 
-        \addplot[blue, mark=*, dashed, forget plot]
+        \addplot[blue, mark=*, dashed, forget plot] % FLUXO, flux differencing
           table [x index=0, y index=3]{../data/pids_euler_ranocha_relative.dat};
-        % \addplot[blue, mark=square*, dashed]
+%         \addplot[orange, mark=square*, dashed, forget plot] % TreeMesh, flux differencing
 %           table [x index=0, y index=1]{../data/pids_euler_ranocha_relative.dat};
-        \addplot[red, mark=diamond*, dashed]
+        \addplot[red, mark=diamond*, dashed] % StructuredMesh, flux differencing
           table [x index=0, y index=2]{../data/pids_euler_ranocha_relative.dat};
@@
 \end{figure}
+
+Note that we do not want to state that Julia is generically faster than Fortran,
+C, or C++. In contrast, we would like to emphasize that Julia can be at least
+as fast as code written in these traditional scientific computing languages.
+We believe that it will be possible to get the same efficiency by rewriting
+FLUXO appropriately. However, the code introspection and profiling tools available
+in Julia make it easier for us to optimize Julia code compared to other languages
+we are familiar with (having a diverse team with a strong background in C, C++,
+and Fortran).
+
+Finally, the different mesh types available in Trixi.jl allow further optimizations.
+For example, if the computational domain is essentially a cube (or square, line)
+without need for curvilinear coordinates, the Cartesian \lstinline{TreeMesh} can
+be used. Depending on the relative computational effort of the discretization
+(such as the polynomial degree and the choice of the volume terms), the
+Cartesian \lstinline{TreeMesh} can be ca.\ \SI{10}{\percent}--\SI{25}{\percent}
+more efficient than the curvilinear meshes in Trixi.jl.
 

### 2021-07-21 paper/paper.tex
enable eprint fields
draft of abstract

@@
 
+% Macros
+\newcommand{\trixi}{Trixi.jl}
+
 \begin{document}
@@
 \begin{abstract}
-TODO: abstract (at most 600 characters, written in plain English with no symbol
-nor formula + references etc.) %TODO
+% Abstract: at most 600 characters, written in plain English with no symbol
+% nor formula + references etc.
+We present Trixi.jl, a Julia library for adaptive high-order numerical simulations
+of hyperbolic partial differential equations (PDEs). Utilizing Julia's strengths,
+Trixi.jl is extendable, easy to use, and fast. We describe the main design choices
+that allow satisfying these goals and compare Trixi.jl with an established open
+source Fortran code for hyperbolic PDEs using the same numerical methods.
+We conclude with an assessment of Julia for simulation-focused scientific
+computing, an area that is still dominated by traditional high-performance
+computing languages such as C, C++, and Fortran.
 \end{abstract}
@@
 \begin{itemize}
-  \item Basic design of Trixi.jl and its capabilities
-  \item PID comparison of Fluxo and Trixi.jl
-  \item Discussion on Julia for numerical simulation science
-  \item Prepare repro repo including scripts for Trixi.jl and Fluxo
+  \item Basic design of \trixi and its capabilities
+  \item PID comparison of Fluxo and \trixi
+  \item Discussion: Julia for numerical simulation science
+  \item Prepare repro repo including scripts for \trixi and Fluxo
   \item Cite \cite{bezanson2017julia,schlottkelakemper2021purely,rackauckas2017differentialequations,revels2016forward,ranocha2021sbp}
   \item Change title to something like ``Adaptive high-order numerical simulations
-  of hyperbolic PDEs with Trixi.jl: A case study of Julia for scientific computing''?
+  of hyperbolic PDEs with \trixi: A case study of Julia for scientific computing''?
   \item Restructure non-conservative stuff and add MHD benchmarks?
@@
 
+\trixi\footnote{https://github.com/trixi-framework/Trixi.jl} provides adaptive
+high-order numerical simulations of hyperbolic partial differential equations (PDEs).
+in Julia \cite{bezanson2017julia}.
+
 TODO %TODO
 
-\section{Design of Trixi.jl and overview of its features}
+\section{Design of \trixi and overview of its features}
 
@@
 to achieve entropy conservation/dissipation or kinetic energy preservation.
-Thus, both Trixi.jl and FLUXO share a common set of features, enabling a reasonable
+Thus, both \trixi and FLUXO share a common set of features, enabling a reasonable
 comparison of their performance.
 
-Here, we compare the serial performance of Trixi.jl and FLUXO when solving a
+Here, we compare the serial performance of \trixi and FLUXO when solving a
 hyperbolic PDE in three space dimensions on curvilinear hex meshes.
@@
 The results of this comparison are visualized in Figure~\ref{fig:PID-Euler}.
-Clearly, Trixi.jl is more than 2x faster than the Fortran code FLUXO
+Clearly, \trixi is more than 2x faster than the Fortran code FLUXO
 while both implement the same numerical methods. This demonstrates the
@@
 %           table [x index=0, y index=1]{../data/pids_euler_weak.dat};
-%           \addlegendentry{\texttt{TreeMesh} (Trixi.jl)}
+%           \addlegendentry{\texttt{TreeMesh} (\trixi)}
         \addplot[red, mark=diamond*, solid] % StructuredMesh, weak form
           table [x index=0, y index=2]{../data/pids_euler_weak.dat};
-          % \addlegendentry{\texttt{StructuredMesh} (Trixi.jl)}
-          \addlegendentry{Trixi.jl}
+          % \addlegendentry{\texttt{StructuredMesh} (\trixi)}
+          \addlegendentry{\trixi}
         \node at (axis cs:13,3.5e-7) {Weak form};
@@
 
-Finally, the different mesh types available in Trixi.jl allow further optimizations.
+Finally, the different mesh types available in \trixi allow further optimizations.
 For example, if the computational domain is essentially a cube (or square, line)
@@
 Cartesian \lstinline{TreeMesh} can be ca.\ \SI{10}{\percent}--\SI{25}{\percent}
-more efficient than the curvilinear meshes in Trixi.jl.
+more efficient than the curvilinear meshes in \trixi.
+
+TODO: Plot of this comparison with the \lstinline{TreeMesh}?
 

### 2021-07-21 paper/paper.tex
change title (preliminary)

@@
   \item Cite \cite{bezanson2017julia,schlottkelakemper2021purely,rackauckas2017differentialequations,revels2016forward,ranocha2021sbp}
-  \item Change title to something like ``Adaptive high-order numerical simulations
-  of hyperbolic PDEs with \trixi: A case study of Julia for scientific computing''?
+  \item The title of our talk at JuliaCon is
+  ``Adaptive and extendable numerical simulations with Trixi.jl''.
+  Shall we use something like
+  ``Adaptive high-order numerical simulations of hyperbolic PDEs with Trixi.jl:
+  A case study of Julia for scientific computing'' for this paper?
   \item Restructure non-conservative stuff and add MHD benchmarks?

### 2021-07-21 paper/paper.tex
WIP: overview of Trixi.jl

@@
 
-% Physicists rule
+% Mathematics is necessary
+\usepackage{amsmath}
+\allowdisplaybreaks
+
+% Physics rules
 \usepackage{siunitx}
@@
 % Macros
-\newcommand{\trixi}{Trixi.jl}
+\usepackage{xspace}
+\newcommand{\ca}[0]{{ca.\@}\xspace}
+\newcommand{\cf}[0]{{cf.\@}\xspace}
+\newcommand{\eg}[0]{{e.g.\@}\xspace}
+\newcommand{\ie}[0]{{i.e.\@}\xspace}
+\newcommand{\etc}[0]{{etc.\@}\xspace}
+\newcommand{\trixi}{Trixi.jl\xspace}
 
@@
   \item Restructure non-conservative stuff and add MHD benchmarks?
+  \item Shall we re-create the schematic overview in tikz?
 \end{itemize}
@@
 \section{Design of \trixi and overview of its features}
+
+\trixi is designed as a library of high-order methods for hyperbolic PDEs of
+the form
+\begin{equation}
+\label{eq:hcl}
+  \partial_t u(t, x) + \sum_{j=1}^d \partial_{x_j} f^j(u) = s(t, x, u),
+  \quad t \in (0, T), x \in \Omega,
+\end{equation}
+in $d \in \{1, 2, 3\}$ space dimensions. Here, the independent variables are
+time $t$ and space coordinates $x \in \Omega \subset \mathbb{R}^d$. The conserved
+quantities are denoted as $u$, \eg, mass, momentum, and energy for the compressible
+Euler equations of an ideal gas.
+The physical model is specified by the fluxes $f^j$ and the source term $s$.
+In addition, suitable initial and boundary conditions (ICs, BCs) are required.
+\trixi also handles non-conservative source terms containing derivatives of the
+conserved quantities as in the shallow water equations or magnetohydrodynamics
+with divergence cleaning.
+
+\subsection{High-level overview of the structure}
+
+\trixi is based on the method of lines. Thus, a discretization of \eqref{eq:hcl}
+is obtained in two steps. First, a spatial semidiscretization is created. Next,
+the resulting ordinary differential equation (ODE) is solved using a time
+integration method. Currently, \trixi focuses on the spatial semidiscretization
+and uses mostly Runge-Kutta methods implemented in OrdinaryDiffEq.jl, a part
+of DifferentialEquations.jl \cite{rackauckas2017differentialequations}.
+
+\begin{figure*}
+  \includegraphics[width=\linewidth]{../figures/trixi_global_overview}
+  \caption{Schematic overview of the basic components of \trixi and how they
+           interact with each other.}
+  \label{fig:trixi_global_overview}
+\end{figure*}
+
+Figure~\ref{fig:trixi_global_overview} presents an overview of the basic
+components of \trixi. The most important structure is the semidiscretization,
+which bundles all information about the spatial part. The mathematical-physical
+model is determined by the \lstinline{equations}, the \lstinline{initial_condition},
+\lstinline{boundary_conditions}, and possible \lstinline{source_terms}. The
+\lstinline{solver} describes purely numerical parameters determining the specific
+discretization method such as discontinuous Galerkin or finite difference
+methods, kinetic energy preserving or shock capturing approaches. Finally, the
+\lstinline{mesh} has a necessarily hybrid role including information about the
+spatial domain and its discretization.
+
+Adding information about the time span $[0, T]$, a semidiscretization of \trixi
+can be converted into an ODE problem, which can be solved by methods from
+OrdinaryDiffEq.jl. The flexible callback infrastructure of this ODE library
+allows us to provide extended functionality for \trixi without modifying any
+main loop. In particular, diverse tasks such as input/output operations, adaptive
+mesh refinement (AMR), and positivity preserving limiting are implemented using
+callbacks.
+
+\subsection{Main features of \trixi}
 
@@
 (such as the polynomial degree and the choice of the volume terms), the
-Cartesian \lstinline{TreeMesh} can be ca.\ \SI{10}{\percent}--\SI{25}{\percent}
+Cartesian \lstinline{TreeMesh} can be \ca \SI{10}{\percent}--\SI{25}{\percent}
 more efficient than the curvilinear meshes in \trixi.
@@
 \begin{itemize}
-  \item Some nice features
-  \item Compilation time can be annoying but also allows cool features and performance
-  \item Extendability: Multi-physics with 350 lines of code \cite{schlottkelakemper2021purely}
+  \item Packages can really be used together at low (or no) cost
+  \item Reproducibility
+  \item Good software development practices (CI, tests, version management)
+  \item Relatively easy to integrate existing C libraries (p4est, HDF5)
+  \item more nice features
+  \item Can be extended: Multi-physics with 350 lines of code
+  \item Compilation time can be annoying but also allows cool features and performance \cite{schlottkelakemper2021purely}
   \item MPI: To be evaluated but the basic tools are already there \cite{byrne2021mpi}
@@
 TODO: Add your funding sources (if any) %TODO
+
+% References
+\input{bib.tex}
 
@@
 
-% References
-\input{bib.tex}
-
 \end{document}

### 2021-07-21 paper/paper.tex
list main features

@@
 
+\subsection{Main features of \trixi}
+
+As of version v0.3.47 (July 2021), \trixi comes with the following features.
+\begin{itemize}
+  \item 1D, 2D, and 3D simulations on line/quad/hex meshes
+  \begin{itemize}
+    \item Cartesian and curvilinear meshes
+    \item Conforming and non-conforming meshes
+    \item Structured and unstructured meshes
+    \item Hierarchical quadtree/octree grid with adaptive refinement
+  \end{itemize}
+
+  \item High-order nodal discontinuous Galerkin methods
+  \begin{itemize}
+    \item Kinetic energy preserving and entropy-stable methods
+    \item Entropy-stable shock-capturing
+    \item Positivity-preserving limiting
+  \end{itemize}
+
+  \item Multiple governing equations
+  \begin{itemize}
+    \item Compressible Euler equations (optionally with self-gravity)
+    \item Magnetohydrodynamics equations
+    \item Acoustic perturbation equations
+    \item Hyperbolic diffusion for elliptic problems
+    \item Lattice-Boltzmann equations (D2Q9 and D3Q27 schemes)
+    \item Several scalar conservation laws (advection, Burgers)
+  \end{itemize}
+
+  \item Integration with the Julia package ecosystem and external tools
+  \begin{itemize}
+    \item Time integration methods from OrdinaryDiffEq.jl
+    \item Automatic differentiation with ForwardDiff.jl
+    \item In-situ visualization with Plots.jl
+    \item Postprocessing with ParaView/Visit via Trixi2Vtk.jl
+  \end{itemize}
+\end{itemize}
+
 \subsection{High-level overview of the structure}
@@
 
-\begin{figure*}
+\begin{figure*}[htbp]
   \includegraphics[width=\linewidth]{../figures/trixi_global_overview}
@@
 
-\subsection{Main features of \trixi}
+\subsection{Review of the main design criteria}
 

### 2021-07-21 paper/paper.tex
WIP: main design choices

@@
 
+\newcommand{\todo}[1]{{\color{red}#1}}
+
 \begin{document}
@@
 
-\section*{TODO}
-
-\begin{itemize}
-  \item Basic design of \trixi and its capabilities
-  \item PID comparison of Fluxo and \trixi
-  \item Discussion: Julia for numerical simulation science
-  \item Prepare repro repo including scripts for \trixi and Fluxo
-  \item Cite \cite{bezanson2017julia,schlottkelakemper2021purely,rackauckas2017differentialequations,revels2016forward,ranocha2021sbp}
-  \item The title of our talk at JuliaCon is
-  ``Adaptive and extendable numerical simulations with Trixi.jl''.
-  Shall we use something like
-  ``Adaptive high-order numerical simulations of hyperbolic PDEs with Trixi.jl:
-  A case study of Julia for scientific computing'' for this paper?
-  \item Restructure non-conservative stuff and add MHD benchmarks?
-  \item Shall we re-create the schematic overview in tikz?
-\end{itemize}
-
 \section{Introduction}
@@
 
-TODO %TODO
+\todo{TODO} %TODO
 
@@
 
-\subsection{Review of the main design criteria}
+\subsection{Review of the main design choices}
 
-TODO %TODO
+The design of \trixi was guided by extensibility, ease of use, and efficiency
+(in this particular order). Thus, it will likely be possible to make (some parts)
+even faster while sacrificing simplicity. Nevertheless, \trixi is already quite
+fast, even compared to established high-performance open source software,
+as described in details in Section~\ref{sec:performance-comparison}. In this
+section, we focus mostly on the first two goals.
+
+\subsubsection{\trixi is a library}
+
+The most important design decision was to create \trixi as a library. In the
+context of simulation-focused scientific computing, there are a lot of (open
+or closed source) codes that are designed as monolithic applications that can
+simulate a specific setup. They can often be configured in limited amounts by
+compile time options and parameter files. In contrast, users of \trixi do not
+need to compile anything or learn the syntax of parameter files. They just load
+the library and write their simulation setups in Julia. Such scripts describing
+a simulation setup for \trixi are called elixirs. As of version v0.3.43, \trixi
+comes with \ca 200 example elixirs. A basic elixir could look as follows.
+
+\begin{lstlisting}[language = Julia]
+# Load all required libraries
+using Trixi, OrdinaryDiffEq, Plots
+
+# Set up a 2D linear advection problem with
+# advection velocity `a`
+a = (1.0, 1.0)
+equations = LinearScalarAdvectionEquation2D(a)
+
+# Choose an initial condition coming with Trixi.jl
+ic = initial_condition_convergence_test
+
+# Set up a discontinuous Galerkin spectral element
+# method with given polynomial degree and surface
+# flux implemented in Trixi.jl
+solver = DGSEM(polydeg=3,
+               surface_flux=flux_lax_friedrichs)
+
+# Create a uniformely refined mesh with periodic
+# boundary conditions in a square domain
+coordinates_min = (-1.0, -1.0) # lower left
+coordinates_max = ( 1.0,  1.0) # upper right
+mesh = TreeMesh(coordinates_min, coordinates_max,
+                initial_refinement_level=4,
+                n_cells_max=10^5, periodicity=true)
+
+# Create semidiscretization with all spatial
+# discretization-related components
+semi = SemidiscretizationHyperbolic(
+  mesh, equations, ic, solver)
+
+# Create an ODE problem from the semidiscretization
+# with time span from 0.0 to 1.0
+ode = semidiscretize(semi, (0.0, 1.0))
+
+# Evolve the ODE problem in time using `solve` from
+# OrdinaryDiffEq with adaptive time stepping
+sol = solve(ode, RDPK3SpFSAL49(), abstol=1.0e-6,
+  reltol=1.0e-6, save_everystep=false)
+
+# Plot the numerical solution at the final time
+plot(sol)
+\end{lstlisting}
+
+\subsubsection{Functions are pure Julia functions}
+
+Since functions are first-class citizens in Julia, they can be passed around
+and used efficiently. Thus, everything that acts like a function can be used
+in \trixi, \eg, to define initial/boundary conditions. While such a flexibility
+is not too uncommon, two-point numerical fluxes are also just functions with a
+specified signature. Thus, users can write their own numerical fluxes in their
+own code and they will work and be as efficient as if they were implemented
+directly in \trixi.
+
+\subsubsection{Ingredients use common interfaces and can be exchanged}
+
+Abstractions such as the \lstinline{mesh} and the \lstinline{equations} are
+based on a common interface. Using multiple dispatch in Julia,
+
+\todo{TODO} %TODO
+
+\subsubsection{New physics can be specified with minimal effort}
+
+Instead of providing only the very basic discretization ingredients such as
+other open source frameworks for PDEs, \trixi includes also some widely used
+physical parts to make it easier to use.
+
+\todo{TODO} %TODO
+
+\subsubsection{Adding new functionality does not require changing existing code}
+
+\todo{TODO} %TODO
+
+\subsubsection{There is no spooky action at a distance}
+
+\todo{TODO} %TODO
 
 \section{Performance comparison with FLUXO}
+\label{sec:performance-comparison}
 
@@
 
-TODO: Andrew, please describe the setups that you used. %TODO
+\todo{TODO: Andrew, please describe the setups that you used} %TODO
 
@@
 
-TODO: Plot of this comparison with the \lstinline{TreeMesh}?
+\todo{TODO: Plot of this comparison with the \lstinline{TreeMesh}?}
 
@@
 
-TODO %TODO
+\todo{TODO} %TODO
 \begin{itemize}
@@
 
+We would like to thank all other contributors to and users of \trixi for their
+help, encouragement, and discussions.
 Funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation)
@@
 
-TODO: Add your funding sources (if any) %TODO
+\todo{TODO: Add your funding sources (if any)} %TODO
 
@@
 \input{bib.tex}
+
+\todo{ % TODO
+\begin{itemize}
+  \item Prepare repro repo including scripts for \trixi and Fluxo
+  \item Cite \cite{schlottkelakemper2021purely,revels2016forward,ranocha2021sbp}
+  \item The title of our talk at JuliaCon is
+  ``Adaptive and extendable numerical simulations with Trixi.jl''.
+  Shall we use something like
+  ``Adaptive high-order numerical simulations of hyperbolic PDEs with Trixi.jl:
+  A case study of Julia for scientific computing'' for this paper?
+  \item Restructure non-conservative stuff and add MHD benchmarks?
+  \item Shall we re-create the schematic overview in tikz?
+  \item Shall we compare \trixi to other open source frameworks such as MFEM, NGSolve, DealII, DUNE, FENICS, Firedrake, ClawPack, ...?
+\end{itemize}}
 

### 2021-07-22 paper/paper.tex
finish first draft of main design choices

@@
 
-\todo{TODO} %TODO
+\todo{TODO %TODO
+\begin{itemize}
+  \item Why we created \trixi
+  \item Overview of the paper \etc
+\end{itemize}}
 
@@
 is not too uncommon, two-point numerical fluxes are also just functions with a
-specified signature. Thus, users can write their own numerical fluxes in their
+specified signature. Thus, users can implement numerical fluxes in their
 own code and they will work and be as efficient as if they were implemented
@@
 
-Abstractions such as the \lstinline{mesh} and the \lstinline{equations} are
-based on a common interface. Using multiple dispatch in Julia,
-
-\todo{TODO} %TODO
+Abstractions such as (variants of) the \lstinline{solver}, the \lstinline{mesh},
+and the \lstinline{equations} are based on a common interface. Using multiple
+dispatch in Julia, internal implementations are specialized accordingly. For
+example, changing the volume terms of the DG discretization from the standard
+weak form to entropy stable and/or shock capturing methods can be achieved by
+passing one additional parameter to the \lstinline{DGSEM} constructor in the
+example above.
 
@@
 other open source frameworks for PDEs, \trixi includes also some widely used
-physical parts to make it easier to use.
-
-\todo{TODO} %TODO
-
-\subsubsection{Adding new functionality does not require changing existing code}
-
-\todo{TODO} %TODO
+physical parts to make it easier to use. Nevertheless, users are not restricted
+to the physics models bundled in \trixi. To setup a new type of \lstinline{equations},
+it is only necessary to create an appropriate \lstinline{struct} bundling all
+parameters and implementing pointwise operations such as the calculation of
+the fluxes $f^j$ in \eqref{eq:hcl} or two-point numerical fluxes. Due to Julia's
+design and just-ahead-of-time compiler, these fluxes can be inlined into the
+library functions of \trixi. Thus, the physics can be separated completely from
+the \lstinline{solver} but are still very efficient.
 
@@
 
-\todo{TODO} %TODO
+In traditional monolithic code bases, it is often necessary to implement a new
+feature for all combinations of possible (compile time) options. This effort
+can make it hard to experiment with new ideas. In contrast, Julia's dynamic
+nature and just-ahead-of-time compiler allow us to implement only the features
+that are strictly necessary. For example, if a user wants to simulate a new
+physics model only on Cartesian grids with smooth solutions, there is no need
+to implement anything for curvilinear coordinates or shock capturing approaches.
 
@@
 
+\subsection{Description of the setup}
+
 Here, we compare the serial performance of \trixi and FLUXO when solving a
@@
 \todo{TODO: Andrew, please describe the setups that you used} %TODO
+
+\subsection{Results of the performance comparison}
 
@@
 
-\todo{TODO} %TODO
+\todo{TODO %TODO
 \begin{itemize}
@@
   \item MPI: To be evaluated but the basic tools are already there \cite{byrne2021mpi}
-\end{itemize}
+\end{itemize}}
 

### 2021-07-22 paper/paper.tex
WIP: assessment of Julia

@@
 to implement anything for curvilinear coordinates or shock capturing approaches.
+In addition, it allows to extend \trixi step by step with new capabilities.
+For example, there are ongoing efforts to incorporate summation by parts finite
+difference methods via SummationByPartsOperators.jl \cite{ranocha2021sbp} and
+discontinuous Galerkin methods on simplices via StartUpDG.jl\footnote{\url{https://github.com/jlchan/StartUpDG.jl}} in \trixi. This becomes feasible since not every
+feature combination needs to be considered and it is totally fine to start with
+simple variants on a subset of the available meshes.
 
@@
 \todo{TODO: Andrew, please describe the setups that you used} %TODO
+
+The code and detailed information necessary to reproduce these numerical
+experiments is available in the accompanying repository \cite{ranocha2021adaptiveRepro}.
 
@@
 
+Julia was designed from scratch to be useful for numerical computing
+\cite{bezanson2017julia}. Its usefulness has been demonstrated for example for
+optimization \cite{dunning2017jump}, data science including
+big data\footnote{Celeste project: \url{https://github.com/jeff-regier/Celeste.jl},
+see \url{https://juliacomputing.com/case-studies/celeste/} (accessed 2021-07-22)},
+machine learning \cite{innes2018fashionable}, and scientific machine learning
+(SciML) \cite{pal2021opening}. There are also mature libraries for ODE problems
+\cite{rackauckas2017differentialequations}, small scale spectral approximations
+\cite{olver2014practical}, and classical finite element methods \cite{badia2020gridap}.
+In addition, there are some packages dedicated to solve specific time-dependent
+PDEs such as \cite{ramadhan2020oceananigans, constantinou2021geophysicalflows}.
+However, there does not seem to be a general framework of high-order methods
+or hyperbolic PDEs in Julia. Thus, \trixi is an ideal candidate for a case study
+of Julia for simulation-focused scientific computing, an area where codes written
+in classical programming languages such as C and Fortran are still dominant
+\cite{krais2021flexi, parsani2021ssdc}.
+
+\subsection{What works well}
+
+\subsubsection{Julia encourages good software development practices}
+
+Many scientists implementing numerical methods often have no real experience or
+education in software development. Julia and its ecosystem support these
+researches by making it easy to set up unit and regression tests, since a
+testing framework is included in the standard library and default package
+generation setups prepare the file structure accordingly. This allows test
+driven development and continuous integration (CI), which makes code restructuring
+and optimization feasible.
+
+In addition, the package manager encourages a unified form of semantic versioning
+across the Julia ecosystem, which lays the basis for the following three
+observations.
+
+\subsubsection{It is easy to set up reproducible numerical experiments}
+
+The package manager makes it easy to reproduce the exact environment, including
+binary dependencies. We have used this feature for all of our papers based on
+\trixi \cite{schlottkelakemper2021purely, ranocha2021preventing}, including this
+one \cite{ranocha2021adaptiveRepro}.
+This facilitates code sharing and reproducible research in computational science,
+which is arguably important but not mainstream yet \cite{barnes2010publish,
+donoho2010invitation, leveque2013top}.
+
+\subsubsection{External C libraries can be integrated relatively easily}
+
+\cite{byrne2021mpi}
+
+\todo{TODO: p4est \cite{burstedde2011p4est} (HDF5)} %TODO
+
+\subsubsection{Packages can really be used together at low (or no) cost}
+
+\todo{TODO} %TODO
+
+\subsubsection{The code base is simple enough to be useful for students}
+
+\todo{TODO} %TODO
+
+\subsubsection{Existing features can be extended and combined easily}
+
+Multi-physics with 350 lines of code \cite{schlottkelakemper2021purely}
+\todo{TODO} %TODO
+
+\subsubsection{\trixi can be extended from outside}
+
+\todo{TODO} %TODO
+
+\subsection{What is still difficult or unknown}
+
 \todo{TODO %TODO
 \begin{itemize}
-  \item Packages can really be used together at low (or no) cost
-  \item Reproducibility
-  \item Good software development practices (CI, tests, version management)
-  \item Relatively easy to integrate existing C libraries (p4est, HDF5)
-  \item more nice features
-  \item Can be extended: Multi-physics with 350 lines of code
-  \item Compilation time can be annoying but also allows cool features and performance \cite{schlottkelakemper2021purely}
+  \item Compilation time can be annoying (serial compiler!) but also allows cool features and performance
   \item MPI: To be evaluated but the basic tools are already there \cite{byrne2021mpi}
@@
 \begin{itemize}
+  \item Full paper: 5--10 pages
   \item Prepare repro repo including scripts for \trixi and Fluxo
@@
   \item Shall we compare \trixi to other open source frameworks such as MFEM, NGSolve, DealII, DUNE, FENICS, Firedrake, ClawPack, ...?
-\end{itemize}}
-
-\clearpage
-\section*{Some guidelines}
-\label{sec:some_guide}
-
-Full paper: 5--10 pages.
+\end{itemize}
 \begin{quote}
@@
   Use cases back up the work by showing how it can be used.
-\end{quote}
-
-The following notes may help you achieve the best effects with the
-\verb juliacon  class file.
-
-\subsection{Writing Julia code}
-
-A special environment is already defined for Julia code,
-built on top of \textit{listings} and \textit{jlcode}.
-
-\begin{verbatim}
-\begin{lstlisting}[language = Julia]
-using Plots
-
-x = -3.0:0.01:3.0
-y = rand(length(x))
-plot(x, y)
-\end{lstlisting}
-\end{verbatim}
-\begin{lstlisting}[language = Julia]
-using Plots
-
-x = -3.0:0.01:3.0
-y = rand(length(x))
-plot(x, y)
-\end{lstlisting}
-
-\subsection{Double Column Figure and Tables}
-\label{subsub:double_fig_tab}
-For generating the output of figures and tables in double column
-we can use the following coding:
-
-\begin{enumerate}
-\item For Figures:
-\begin{verbatim}
-\begin{figure*}...\end{figure*}
-\end{verbatim}
-\item For landscape figures:
-\begin{verbatim}
-\begin{sidewaysfigure*}...\end{sidewaysfigure*}
-\end{verbatim}
-\item For Tables:
-\begin{verbatim}
-\begin{table*}...\end{table*}
-\end{verbatim}
-\item For landscape tables:
-\begin{verbatim}
-\begin{sidewaystable*}...\end{sidewaystable*}
-\end{verbatim}
-\end{enumerate}
-
-\subsection{Enunciations}
-\label{subsub:enunciation}
-
-The \verb juliacon   class file generates the enunciations with the help of
-the following commands:
-\begin{verbatim}
-\begin{theorem}...\end{theorem}
-\begin{strategy}...\end{strategy}
-\begin{property}...\end{property}
-\begin{proposition}...\end{proposition}
-\begin{lemma}...\end{lemma}
-\begin{example}...\end{example}
-\begin{proof}...\end{proof}
-\begin{definition}...\end{definition}
-\begin{algorithm}...\end{algorithm}
-\begin{remark}...\end{remark}
-\end{verbatim}
-The above-mentioned coding can also include optional arguments
-such as
-\begin{verbatim}
-\begin{theorem}[...]. Example for theorem:
-\begin{theorem}[Generalized Poincare Conjecture]
-Four score and seven ... created equal.
-\end{theorem}
-\end{verbatim}
-
-\begin{theorem}[Generalized Poincare Conjecture]
-Four score and seven years ago our fathers brought forth,
-upon this continent, a new nation, conceived in Liberty,
- and dedicated to the proposition that all men are
-created equal.
-\end{theorem}
-
-\subsection{Balancing column at last page}
-\label{subsub:Balance}
-For balancing the both column length at last page use:
-\begin{verbatim}
-\vadjust{\vfill\pagebreak}
-\end{verbatim}
-
-% \vadjust{\vfill\pagebreak}
-
-at appropriate place in your \TeX{} file or in bibliography file.
+\end{quote}}
 

### 2021-07-22 paper/paper.tex
first draft of assessment

@@
 
-\todo{TODO %TODO
+\todo{TODO %TODO Introduction
 \begin{itemize}
@@
 
-\todo{TODO: Andrew, please describe the setups that you used} %TODO
+\todo{TODO: Andrew, please describe the setups that you used} %TODO Andrew
 
@@
 
-\cite{byrne2021mpi}
+We have written a Julia wrapper\footnote{\url{https://github.com/trixi-framework/P4est.jl}}
+of the C library \texttt{p4est} \cite{burstedde2011p4est}. In our experience,
+it is relatively easy to do so in Julia and the package manager in combination
+with BinaryBuilder.jl makes it convenient to distribute the necessary binaries.
+In particular, users do not have to compile C libraries on their own system
+and the binaries are available on all major platforms including Linux, Mac OS,
+and Windows.
 
-\todo{TODO: p4est \cite{burstedde2011p4est} (HDF5)} %TODO
+Similarly, there is no need to compile HDF5 on a user system due to the
+availability of HDF5.jl. There is still work to be done how to improve the
+experience in combination with MPI on high performance computing (HPC) clusters,
+but some promising approaches exist \cite{byrne2021mpi}.
 
@@
 
-\todo{TODO} %TODO
+Due to Julia's package manager, the cost of using external dependencies is low
+enough to not be a problem. Traditionally, many scientific simulation codes tend
+to reduce the number of external dependencies as much as possible due to the
+complexity of handling different versions and making everything work together.
+Such an approach often leads to a lot of code duplication. For example, many
+CFD codes implement their own time integration methods, arguing that the spatial
+part is much more complex. While this is often true, it is still more efficient
+in terms of human workload to be able to reuse existing implementations. This
+allows expert on time integration methods to develop specialized algorithms and
+implement them in open source software while practitioners or researchers focusing
+on spatial semidiscretizations ca benefit immediately by changing only one line
+of code. For example, optimized time integration methods were developed in
+\cite{ranocha2021optimized} and implemented in OrdinaryDiffEq.jl. Thus, they
+can be used with \trixi by changing a single line of code. In contrast, researchers
+working without external dependencies would need to use a day or two to digest
+all the details and implement the methods in their own codes, making it much
+less likely that they will do this and benefit from recent algorithmic developments.
+
+An additional example is given by automatic differentiation. Due to dispatch,
+it is conveniently possible to use forward mode automatic differentiation
+\cite{revels2016forward} to compute Jacobians of semidiscretizations of nonlinear
+conservation laws with \trixi or to differentiate through a complete simulation
+including time integration methods from OrdinaryDiffEq.jl.
 
@@
 
-\todo{TODO} %TODO
+Due to the package manager, \trixi and all related ODE and visualization tools
+can be installed with a single command, reducing the overhead of exploring the
+code. In addition, Julia's expressiveness and the ability to work with a restricted
+subset of all possible features have enabled more than 18 students to use \trixi
+for their coursework or theses. In our experience, the time required to get
+started with traditional monolithic code bases is so long that students either
+choose to build their own implementations specialized on their tasks or can only
+work on toy problems since the remaining time is rather restricted.
 
@@
 
-Multi-physics with 350 lines of code \cite{schlottkelakemper2021purely}
-\todo{TODO} %TODO
-
-\subsubsection{\trixi can be extended from outside}
-
-\todo{TODO} %TODO
+Due to Julia's high-level programming approach, dynamic typing, and multiple
+dispatch, it is easy to combine existing functionality efficiently. For example,
+the single-physics solvers for hyperbolic PDEs of \trixi were extended to
+a multi-physics setup for the compressible Euler equations with self-gravity
+with roughly 350 lines of code \cite{schlottkelakemper2021purely}.
+In addition, \trixi can be extended from the outside without modifying the main
+source code, which makes it easy to set up new simulation approaches and analyze
+existing ones, \eg by fluctuation simulations \cite{ranocha2021preventing}.
 
@@
 
-\todo{TODO %TODO
-\begin{itemize}
-  \item Compilation time can be annoying (serial compiler!) but also allows cool features and performance
-  \item MPI: To be evaluated but the basic tools are already there \cite{byrne2021mpi}
-\end{itemize}}
+\subsubsection{Compilation times can be annoying}
+
+Since Julia uses a serial just-ahead-of-time compiler that does currently not cache
+code between different Julia sessions, the initial compilation time can be
+annoying. For example, the time to finish the first simulation in a fresh Julia
+session and plot the results on a notebook can take between 30 seconds and a
+minute. Having compiled the code, the second simulation and plot take less than
+0.05 seconds. Thus, switching to Julia requires adapting the workflow compared
+to other languages such as keeping a REPL session active for a longer time and
+using packages such as Revise.jl\footnote{\url{https://github.com/timholy/Revise.jl}}.
+This is particularly problematic in HPC environments. There are tools such as
+PackageCompiler.jl\footnote{\url{https://github.com/JuliaLang/PackageCompiler.jl}}
+that can help with these tasks but a good workflow usable for development and
+deployment still has to be found.
+
+\subsubsection{MPI-based parallel simulations need to be established}
+
+MPI: To be evaluated but the basic tools are already there \cite{byrne2021mpi}
+
+\todo{TODO: Michael, could you please write something here? You are much more
+experienced with MPI and HPC} %TODO Michael
 
@@
 
-\todo{TODO} %TODO
+\todo{TODO %TODO Summary
+\begin{itemize}
+  \item Learning a new programming language has some overhead but it pays off
+\end{itemize}}
 
@@
 
-\todo{TODO: Add your funding sources (if any)} %TODO
+\todo{TODO: Add your funding sources that need to be here (if any)} %TODO Funding
 

### 2021-07-22 paper/paper.tex
adapt TODO notes
Julia is fast

@@
 
+\subsubsection{Julia is fast}
+
+As demonstrated in Section~\ref{sec:performance-comparison}, there is no reason
+why Julia should be generically slower than traditional high-performance programming
+languages such as C, C++, and Fortran. A minor exception of this rule is that
+Julia is build on LLVM, which does not perform the same optimizations as most
+optimizing C/C++/Fortran compilers in use nowadays. In particular, LLVM does
+not use fused multiply add (FMA) instructions by default while GCC and Intel
+compilers do. Thus, one needs to opt-in to these optimizations explicitly in
+Julia. Luckily, that is relatively easy to do due to Julia's code manipulation
+techniques which enable the macro \lstinline{@muladd}\footnote{\url{https://github.com/SciML/MuladdMacro.jl}}.
+
 \subsubsection{Julia encourages good software development practices}
@@
   \item Prepare repro repo including scripts for \trixi and Fluxo
-  \item Cite \cite{schlottkelakemper2021purely,revels2016forward,ranocha2021sbp}
   \item The title of our talk at JuliaCon is
@@
   \item Shall we compare \trixi to other open source frameworks such as MFEM, NGSolve, DealII, DUNE, FENICS, Firedrake, ClawPack, ...?
+  \item Are there other aspects we should discuss?
 \end{itemize}

### 2021-07-27 paper/paper.tex
update TODO note
update Trixi version after merging Jesse's PR

@@
 
-As of version v0.3.47 (July 2021), \trixi comes with the following features.
+% TODO: Update Trixi version
+As of version v0.3.50 (July 2021), \trixi comes with the following features.
 \begin{itemize}
-  \item 1D, 2D, and 3D simulations on line/quad/hex meshes
+  \item 1D, 2D, and 3D simulations on line/quad/hex/simplex meshes
   \begin{itemize}
@@
 
-  \item High-order nodal discontinuous Galerkin methods
+  \item High-order nodal and modal discontinuous Galerkin methods
   \begin{itemize}
@@
 difference methods via SummationByPartsOperators.jl \cite{ranocha2021sbp} and
-discontinuous Galerkin methods on simplices via StartUpDG.jl\footnote{\url{https://github.com/jlchan/StartUpDG.jl}} in \trixi. This becomes feasible since not every
+more discontinuous Galerkin methods on simplices via StartUpDG.jl\footnote{\url{https://github.com/jlchan/StartUpDG.jl}} in \trixi. This becomes feasible since not every
 feature combination needs to be considered and it is totally fine to start with
@@
   A case study of Julia for scientific computing'' for this paper?
-  \item Restructure non-conservative stuff and add MHD benchmarks?
+  \item Restructure non-conservative stuff and add MHD benchmarks
+  \item Add new features (Jesse's mesh and solver) to the list and update version, numer of examples, features, \etc
   \item Shall we re-create the schematic overview in tikz?

### 2021-07-27 paper/paper.tex
add references besard2018juliagpu, omlin2020solving

@@
 \cite{bezanson2017julia}. Its usefulness has been demonstrated for example for
+GPU programming \cite{besard2018juliagpu, omlin2020solving},
 optimization \cite{dunning2017jump}, data science including

### 2021-07-27 paper/paper.tex
update TODO notes

@@
 
-% TODO: Update Trixi version
+% TODO: Update Trixi version below
 As of version v0.3.50 (July 2021), \trixi comes with the following features.
@@
 the library and write their simulation setups in Julia. Such scripts describing
-a simulation setup for \trixi are called elixirs. As of version v0.3.43, \trixi
+a simulation setup for \trixi are called elixirs. As of version v0.3.50, \trixi
 comes with \ca 200 example elixirs. A basic elixir could look as follows.
+% TODO: Update Trixi version above
 
@@
 
-\todo{TODO: Andrew, please describe the setups that you used} %TODO Andrew
+\todo{TODO: Andrew, could you please describe the setups that you used? %TODO Andrew
+\begin{itemize}
+  \item Physical setup
+  \item Numerical methods/parameters
+  \item Software versions, compiler \etc
+  \item Hardware
+\end{itemize}}
 
@@
 \todo{TODO: Plot of this comparison with the \lstinline{TreeMesh}?}
+
+\todo{TODO: Add MHD results?}
 
@@
 \begin{itemize}
+  \item Julia is suited for simulation-focused scientific computing but some
+        things need to be improved/demonstrated for HPC
+  \item Julia allows us to do some cool stuff in \trixi
   \item Learning a new programming language has some overhead but it pays off
@@
   \item Full paper: 5--10 pages
-  \item Prepare repro repo including scripts for \trixi and Fluxo
+  \item Prepare repro repo including scripts for \trixi and FLUXO
   \item The title of our talk at JuliaCon is
@@
   \item Restructure non-conservative stuff and add MHD benchmarks
-  \item Add new features (Jesse's mesh and solver) to the list and update version, numer of examples, features, \etc
   \item Shall we re-create the schematic overview in tikz?

### 2021-07-27 paper/paper.tex
update TODO notes as discussed in the Trixi meeting

@@
     \item Hierarchical quadtree/octree grid with adaptive refinement
+    \item Forests of qudtrees/octrees with \texttt{p4est} \cite{burstedde2011p4est}
+          via P4est.jl\footnote{\url{https://github.com/trixi-framework/P4est.jl}}
   \end{itemize}
 
-  \item High-order nodal and modal discontinuous Galerkin methods
+  \item High-order discontinuous Galerkin methods
   \begin{itemize}
@@
 and used efficiently. Thus, everything that acts like a function can be used
-in \trixi, \eg, to define initial/boundary conditions. While such a flexibility
+in \trixi, \eg, to define initial/boundary conditions. While the flexibility
+to change initial/boundary conditions
 is not too uncommon, two-point numerical fluxes are also just functions with a
@@
 
+\todo{TODO: Mention multiple dispatch} %TODO
+
 \section{Performance comparison with FLUXO}
@@
 comparison of their performance.
+
+\todo{TODO: Be careful how to formulate this, be more neutral} %TODO
+
+\todo{TODO: Point out that both codes have features the other ones does not} %TODO
 
@@
 
-\subsubsection{External C libraries can be integrated relatively easily}
+\subsubsection{External libraries can be integrated relatively easily}
 
@@
 but some promising approaches exist \cite{byrne2021mpi}.
+
+\todo{TODO: Mention also Fortran libraries/codes (KROME, HOHQMesh)} %TODO
 
@@
   \item Shall we re-create the schematic overview in tikz?
-  \item Shall we compare \trixi to other open source frameworks such as MFEM, NGSolve, DealII, DUNE, FENICS, Firedrake, ClawPack, ...?
+  \item Shall we compare \trixi (in terms of structure, not performance!)
+        to other open source frameworks such as MFEM,
+        NGSolve, DealII, DUNE, FENICS, Firedrake, ClawPack, ...? Definitely not
+        explicitly to not make people angry, but maybe implicitly without
+        mentioning their names?
+        \begin{itemize}
+          \item Everything in Julia, not a two language problem!
+          \item Other things? Gregor will think about it!
+        \end{itemize}
   \item Are there other aspects we should discuss?
+  \item Check affiliations
+  \item Maybe include some cool pictures? Give a flavor of the kind of simulations
+        you can do with \trixi? Pictures trigger people! For example after the
+        list of main features?
+  \item Summary of community, collaborators, and users: GitHub statistics, contributors,
+        Vikram and Chandrashekar, Steven Frankel and Philipp Öffner (maybe), ...
+        Related to the relatively large number of students!
+        $\to$ Jesse will make a PR!
+  \item A simple example where we really use multiple dispatch?
+        \begin{itemize}
+          \item Just pass functions around
+          \item Dispatch on parameters of DG struct
+        \end{itemize}
 \end{itemize}

### 2021-07-27 paper/paper.tex
mention wrapping Fortran libraries

@@
 can make it hard to experiment with new ideas. In contrast, Julia's dynamic
-nature and just-ahead-of-time compiler allow us to implement only the features
+nature, multiple dispatch, and just-ahead-of-time compilation allow us to implement only the features
 that are strictly necessary. For example, if a user wants to simulate a new
@@
 feature combination needs to be considered and it is totally fine to start with
-simple variants on a subset of the available meshes.
-
-\todo{TODO: Mention multiple dispatch} %TODO
+simple variants on a subset of the available meshes, selected via multiple dispatch.
 
@@
 to achieve entropy conservation/dissipation or kinetic energy preservation.
-Thus, both \trixi and FLUXO share a common set of features, enabling a reasonable
-comparison of their performance.
+At the same time, both codes also have features the other one does not, \eg
+parabolic equations and MPI in FLUXO or multiple mesh types in \trixi.
+Nevertheless, both \trixi and FLUXO share a common set of features, enabling a
+reasonable comparison of their performance.
 
 \todo{TODO: Be careful how to formulate this, be more neutral} %TODO
-
-\todo{TODO: Point out that both codes have features the other ones does not} %TODO
 
@@
 
-\todo{TODO: Mention also Fortran libraries/codes (KROME, HOHQMesh)} %TODO
+It is also possible to wrap Fortran libraries/tools in Julia. For example, we
+have created the wrappers KROME.jl\footnote{\url{https://github.com/trixi-framework/KROME.jl}}
+for KROME, a package to embed chemistry in astrophysical simulations
+\cite{grassi2014krome}, and HOHQMesh.jl\footnote{\url{https://github.com/trixi-framework/HOHQMesh.jl}}
+for HOHQMesh\footnote{\url{https://github.com/trixi-framework/HOHQMesh}}, a high order
+hex-quad mesh generator written in Fortran.
 

### 2021-07-27 paper/paper.tex
mention two-language problem

@@
 library functions of \trixi. Thus, the physics can be separated completely from
-the \lstinline{solver} but are still very efficient.
+the \lstinline{solver} but are still very efficient. In particular, this allows
+to reuse the same numerical fluxes for discontinuous Galerkin methods, finite
+difference methods, and variants of finite volume methods.
 
@@
 At the same time, both codes also have features the other one does not, \eg
-parabolic equations and MPI in FLUXO or multiple mesh types in \trixi.
-Nevertheless, both \trixi and FLUXO share a common set of features, enabling a
-reasonable comparison of their performance.
+parabolic equations and MPI in FLUXO or multiple mesh types and more physics setups
+in \trixi. Nevertheless, both \trixi and FLUXO share a common set of features,
+enabling a reasonable comparison of their performance.
 
-\todo{TODO: Be careful how to formulate this, be more neutral} %TODO
+\todo{TODO: Be careful how to formulate this, be neutral and avoid making people angry} %TODO
 
@@
 with BinaryBuilder.jl makes it convenient to distribute the necessary binaries.
-In particular, users do not have to compile C libraries on their own system
+In particular, users do not have to compile libraries on their own system
 and the binaries are available on all major platforms including Linux, Mac OS,
@@
 Similarly, there is no need to compile HDF5 on a user system due to the
-availability of HDF5.jl. There is still work to be done how to improve the
-experience in combination with MPI on high performance computing (HPC) clusters,
-but some promising approaches exist \cite{byrne2021mpi}.
+availability of HDF5.jl\footnote{\url{https://github.com/JuliaIO/HDF5.jl}}.
+There is still work to be done how to improve the experience in combination
+with MPI on high performance computing (HPC) clusters, but some promising
+approaches exist \cite{byrne2021mpi}.
 
@@
 including time integration methods from OrdinaryDiffEq.jl.
+
+\subsubsection{Julia solves the two-language problem}
+
+Many libraries for simulation-focused scientific computing are written in
+languages such as C, C++, or Fortran to get decent performance. On top of these
+low-level details, high-level wrappers are often provided to make it easier
+for users to apply the algorithms for their problems. In contrast, \trixi is
+written completely in Julia.
+
+Such a difference between the programming languages used in the front- and
+back-ends lead to a well-known barrier between user and developers of a library.
+Julia was designed from the beginning to make it possible to write both simple
+high-level code and highly efficient compute kernels, thereby solving the
+two-language problem --- at least to a large extend. Highly efficient low-level
+code will of course still look different from simple high-level code, but the
+barrier between them is much smaller. In particular, this makes it easier for
+users to contribute to a package and act as developers. This also contributes
+to the following observation.
 
@@
   \item Check affiliations
+  \item Check order of authors
   \item Maybe include some cool pictures? Give a flavor of the kind of simulations

### 2021-08-01 paper/paper.tex
update MPI TODO note
more positive aspects of julia

@@
 
-  \item High-order discontinuous Galerkin methods
+  \item High-order matrix-free discontinuous Galerkin methods
   \begin{itemize}
@@
 passing one additional parameter to the \lstinline{DGSEM} constructor in the
-example above.
+example above. Moreover, there is no hidden global state and multiple instances
+of similar structures can be instantiated simultaneously. In particular, multiple
+semidiscretizations (in possibly different spatial dimensions) can be created
+and used in the same code.
 
@@
 
+Julia itself and most packages are developed on GitHub. The community provides
+a lot of tools for using corresponding CI setups, preparing and publishing
+documentation\footnote{\url{https://github.com/JuliaDocs/Documenter.jl}},
+and related tasks. This setup also encourages other good software development
+practices that we use in \trixi such as mandatory code review before pull requests
+are merged.
+
 In addition, the package manager encourages a unified form of semantic versioning
@@
 
+\subsubsection{Julia is free}
+
+Julia itself is released under the MIT license and ships some GPL-licensed
+third-party software (that can optionally be disabled for commercial purposes).
+Many packages in the Julia ecosystem follow this approach and are freely available
+under the permissive MIT license, including \trixi. This allows using Julia
+without needing to pay for commercial software. In particular, students can
+work with the software on their private computers without restrictions.
+
 \subsection{What is still difficult or unknown}
@@
 \todo{TODO: Michael, could you please write something here? You are much more
-experienced with MPI and HPC} %TODO Michael
+experienced with MPI and HPC. Be aware of \cite{omlin2020solving} to not step
+on anybodies toes...} %TODO Michael
 

### 2021-08-01 paper/paper.tex
first draft of summary

@@
 \trixi\footnote{https://github.com/trixi-framework/Trixi.jl} provides adaptive
-high-order numerical simulations of hyperbolic partial differential equations (PDEs).
+high-order numerical simulations of hyperbolic partial differential equations (PDEs)
 in Julia \cite{bezanson2017julia}.
@@
 
-\todo{TODO %TODO Summary
-\begin{itemize}
-  \item Julia is suited for simulation-focused scientific computing but some
-        things need to be improved/demonstrated for HPC
-  \item Julia allows us to do some cool stuff in \trixi
-  \item Learning a new programming language has some overhead but it pays off
-\end{itemize}}
+We have presented \trixi, a Julia library for adaptive high-order numerical
+simulations of hyperbolic PDEs. As researchers in numerical analysis and
+scientific computing, our goals were to create a framework that is extendable,
+easy to use, and fast (in this particular order). Making use of Julia's strengths,
+we have been successful based on recent publications making use of \trixi,
+the number of students and researchers working with \trixi, and serial performance
+comparisons with established Fortran codes. Having worked on \trixi for a bit
+more than one year allows us to give an assessment of Julia for simulation-focused
+scientific computing.
+
+Based on our experience, we consider Julia to be suitable for simulation-focused
+scientific computing, in particular for hyperbolic PDEs, computational fluid
+dynamics, and related problems --- at least on the scale of shared memory
+parallelism. The scalability of high-order methods for hyperbolic PDEs written
+in Julia to high performance computing applications still needs to be demonstrated.
+However, we do not expect unsolvable problems there since Julia is not generically
+slower than traditional compiled HPC programming languages. Nevertheless, it
+appears to be more complicated to scale Julia to HPC while still keeping the
+code as simple, clean, and flexible as it is now.
+
+Learning a new programming language requires of course some effort and needs some
+time. Nevertheless, it pays off in our experience and we do not want to miss the
+nice features of \trixi enabled by Julia and its ecosystem. Thus, while there is
+of course no need to switch to Julia if people are satisfied with their existing
+tools, we encourage researchers to try out Julia and stay for a while. The Julia
+community (on the Julia Discourse forum\footnote{\url{https://discourse.julialang.org/}}
+or Slack/Zulip workspaces) is usually welcoming and helpful, both for new and
+experienced users. Making use of existing learning material and asking questions
+appears to be important since a new programming style is necessary to make full
+use of all of Julia's strengths.
 
@@
   \item Prepare repro repo including scripts for \trixi and FLUXO
-  \item The title of our talk at JuliaCon is
-  ``Adaptive and extendable numerical simulations with Trixi.jl''.
-  Shall we use something like
-  ``Adaptive high-order numerical simulations of hyperbolic PDEs with Trixi.jl:
-  A case study of Julia for scientific computing'' for this paper?
   \item Restructure non-conservative stuff and add MHD benchmarks
@@
         \begin{itemize}
-          \item Everything in Julia, not a two language problem!
-          \item Other things? Gregor will think about it!
+          \item Gregor will think about it!
         \end{itemize}

### 2021-08-01 paper/paper.tex
first draft of introduction

@@
+% Allow breaking URLs at hyphens
+\PassOptionsToPackage{hyphens}{url}
+
 % JuliaCon proceedings template
@@
 
-\trixi\footnote{https://github.com/trixi-framework/Trixi.jl} provides adaptive
+We are broadly interested in simulation-focused scientific computing, in particular
+in hyperbolic partial differential equations (PDEs), computational fluid dynamics (CFD),
+and related problems. The focus of our research ranges from specific applications
+in CFD over general multi-physics coupling strategies to the required high-order
+numerical methods themselves. In addition, we are academics and involved in teaching
+students in these areas of science. Thus, we would like to have a code that is
+\begin{enumerate}
+  \item extendable to be useful for research and development
+  \item easy to apply/understand to be useful for students/collaborators
+  \item fast enough to be useful not only for small toy problems
+\end{enumerate}
+Additionally, we are greedy\footnote{See \url{https://julialang.org/blog/2012/02/why-we-created-julia/}} and want to have all these features in a single code.
+Roughly one year of collaborative work has resulted in the current version of
+\trixi\footnote{https://github.com/trixi-framework/Trixi.jl}, providing adaptive
 high-order numerical simulations of hyperbolic partial differential equations (PDEs)
-in Julia \cite{bezanson2017julia}.
+in Julia \cite{bezanson2017julia}. Starting as an experiment, we have been able
+to reach more and more of our goals with \trixi.
 
-\todo{TODO %TODO Introduction
-\begin{itemize}
-  \item Why we created \trixi
-  \item Overview of the paper \etc
-\end{itemize}}
+\todo{TODO: Jesse wants to write about the \trixi community} %TODO: Jesse
+
+In this article, we present an overview of the main features and design decisions
+of \trixi in Section~\ref{sec:design-of-trixi}, laying the ground for an extendable
+and easy-to-use framework of high-order methods for hyperbolic PDEs. Next, we
+compare the serial performance with an established HPC Fortran code in
+Section~\ref{sec:performance-comparison}, demonstrating that Julia is not generically
+slower than traditional HPC languages (and can even be faster in this particular case).
+Thereafter, we present an assessment of Julia for simulation-focused scientific
+computing based on our experience with \trixi in Section~\ref{sec:assessment-of-julia}.
+Finally, we summary our findings and conclusions in Section~\ref{sec:summary}.
 
 \section{Design of \trixi and overview of its features}
+\label{sec:design-of-trixi}
 
@@
 \section{Assessment of Julia for simulation-focused scientific computing}
+\label{sec:assessment-of-julia}
 
@@
 \section{Summary and conclusions}
+\label{sec:summary}
 

### 2021-08-02 paper/paper.tex
added funding and resources for ARW to acknowledgements

@@
 Dynamics-Geometry-Structure.
+Andrew Winters was funded through Vetenskapsr{\aa}det, Sweden grant
+agreement 2020-03642 VR.
+
+The performance computations were enabled by resources provided by the Swedish National
+Infrastructure for Computing (SNIC) at Tetralith partially funded by the Swedish
+Research Council through grant agreement no. 2018-05973.
 

### 2021-08-02 paper/paper.tex
paper edits

--- 2021-08-01 paper/paper.tex
@@
 Trixi.jl is extendable, easy to use, and fast. We describe the main design choices
-that allow satisfying these goals and compare Trixi.jl with an established open
+that enable these features and compare Trixi.jl with an established open
 source Fortran code for hyperbolic PDEs using the same numerical methods.
@@
 and related problems. The focus of our research ranges from specific applications
-in CFD over general multi-physics coupling strategies to the required high-order
-numerical methods themselves. In addition, we are academics and involved in teaching
-students in these areas of science. Thus, we would like to have a code that is
+in CFD to general multi-physics coupling strategies to the high-order numerical methods 
+such simulations are built upon. In addition, we are academics and involved in teaching 
+students in these areas of science. Thus, we would like to have a 
+code that is
 \begin{enumerate}
@@
   \item easy to apply/understand to be useful for students/collaborators
-  \item fast enough to be useful not only for small toy problems
+  \item fast enough to be useful beyond small toy problems
 \end{enumerate}
-Additionally, we are greedy\footnote{See \url{https://julialang.org/blog/2012/02/why-we-created-julia/}} and want to have all these features in a single code.
+Additionally, we are greedy\footnote{See \url{https://julialang.org/blog/2012/02/why-we-created-julia/}} 
+and wish to include all these features within a single code.
 Roughly one year of collaborative work has resulted in the current version of
@@
 to reach more and more of our goals with \trixi.
-
-\todo{TODO: Jesse wants to write about the \trixi community} %TODO: Jesse
 
@@
   \partial_t u(t, x) + \sum_{j=1}^d \partial_{x_j} f^j(u) = s(t, x, u),
-  \quad t \in (0, T), x \in \Omega,
+  \quad t \in (0, T),
 \end{equation}
@@
     \item Hierarchical quadtree/octree grid with adaptive refinement
-    \item Forests of qudtrees/octrees with \texttt{p4est} \cite{burstedde2011p4est}
+    \item Forests of quadtrees/octrees with \texttt{p4est} \cite{burstedde2011p4est}
           via P4est.jl\footnote{\url{https://github.com/trixi-framework/P4est.jl}}
@@
 
-Adding information about the time span $[0, T]$, a semidiscretization of \trixi
-can be converted into an ODE problem, which can be solved by methods from
-OrdinaryDiffEq.jl. The flexible callback infrastructure of this ODE library
+Once information about the time span $[0, T]$ is provided, a semidiscretization
+in \trixi can be converted into an ODE problem, which can be solved by methods 
+from OrdinaryDiffEq.jl. The flexible callback infrastructure of this ODE library
 allows us to provide extended functionality for \trixi without modifying any
@@
 The design of \trixi was guided by extensibility, ease of use, and efficiency
-(in this particular order). Thus, it will likely be possible to make (some parts)
-even faster while sacrificing simplicity. Nevertheless, \trixi is already quite
+(in this particular order). Thus, it is likely possible to make (some parts) even 
+faster at the cost of simplicity and readibility. Nevertheless, \trixi is already quite
 fast, even compared to established high-performance open source software,
@@
 or closed source) codes that are designed as monolithic applications that can
-simulate a specific setup. They can often be configured in limited amounts by
-compile time options and parameter files. In contrast, users of \trixi do not
-need to compile anything or learn the syntax of parameter files. They just load
-the library and write their simulation setups in Julia. Such scripts describing
-a simulation setup for \trixi are called elixirs. As of version v0.3.50, \trixi
-comes with \ca 200 example elixirs. A basic elixir could look as follows.
+simulate a specific setup. These codes can often be configured to a limited extent 
+by specifying compile time options or editing parameter files. In contrast, users of 
+\trixi do not need to compile anything or learn parameter file syntax. Instead, they can
+just load the library and write their simulation setups in Julia. Scripts which describe
+a simulation setup for \trixi are referred to as ``elixirs''. As of version v0.3.50, \trixi
+comes with \ca 200 example elixirs. A basic elixir might look as follows:
 % TODO: Update Trixi version above
@@
 library functions of \trixi. Thus, the physics can be separated completely from
-the \lstinline{solver} but are still very efficient. In particular, this allows
+the \lstinline{solver} while retaining efficiency. In particular, this allows
 to reuse the same numerical fluxes for discontinuous Galerkin methods, finite
@@
 difference methods via SummationByPartsOperators.jl \cite{ranocha2021sbp} and
-more discontinuous Galerkin methods on simplices via StartUpDG.jl\footnote{\url{https://github.com/jlchan/StartUpDG.jl}} in \trixi. This becomes feasible since not every
-feature combination needs to be considered and it is totally fine to start with
-simple variants on a subset of the available meshes, selected via multiple dispatch.
+discontinuous Galerkin methods on simplices via StartUpDG.jl\footnote{\url{https://github.com/jlchan/StartUpDG.jl}} 
+in \trixi. This becomes feasible since not every feature combination needs to be considered, 
+enabling developers to introduce variants of existing solvers on only a subset of the available
+mesh types via multiple dispatch.
 
@@
 we are familiar with (having a diverse team with a strong background in C, C++,
-and Fortran).
+and Fortran). \todo{from Jesse: what does this mean?}
 
@@
 \cite{olver2014practical}, and classical finite element methods \cite{badia2020gridap}.
-In addition, there are some packages dedicated to solve specific time-dependent
+In addition, there are some packages dedicated to solving specific time-dependent
 PDEs such as \cite{ramadhan2020oceananigans, constantinou2021geophysicalflows}.
-However, there does not seem to be a general framework of high-order methods
-or hyperbolic PDEs in Julia. Thus, \trixi is an ideal candidate for a case study
+However, there does not seem to be a general framework for high-order methods
+for hyperbolic PDEs in Julia. Thus, \trixi is an ideal candidate for a case study
 of Julia for simulation-focused scientific computing, an area where codes written
@@
 and related tasks. This setup also encourages other good software development
-practices that we use in \trixi such as mandatory code review before pull requests
+practices that we use in \trixi, such as mandatory code review before pull requests
 are merged.
@@
 complexity of handling different versions and making everything work together.
-Such an approach often leads to a lot of code duplication. For example, many
+Such an approach typically leads to significant code duplication. For example, many
 CFD codes implement their own time integration methods, arguing that the spatial
-part is much more complex. While this is often true, it is still more efficient
+part is much more complex \todo{should we cite something for this?}. While this is often true, it is still more efficient
 in terms of human workload to be able to reuse existing implementations. This
@@
 of code. For example, optimized time integration methods were developed in
-\cite{ranocha2021optimized} and implemented in OrdinaryDiffEq.jl. Thus, they
+\cite{ranocha2021optimized} and implemented in OrdinaryDiffEq.jl. These methods
 can be used with \trixi by changing a single line of code. In contrast, researchers
-working without external dependencies would need to use a day or two to digest
-all the details and implement the methods in their own codes, making it much
-less likely that they will do this and benefit from recent algorithmic developments.
+working without external dependencies would need to spend a day or two to digest
+details before even beginning implementation. This additional overhead in developer
+time makes it less likely that researchers would implement these methods in their own 
+codes and benefit from recent algorithmic developments.
 
@@
 high-level code and highly efficient compute kernels, thereby solving the
-two-language problem --- at least to a large extend. Highly efficient low-level
+two-language problem --- at least to a significant extent. Highly efficient low-level
 code will of course still look different from simple high-level code, but the
 barrier between them is much smaller. In particular, this makes it easier for
-users to contribute to a package and act as developers. This also contributes
-to the following observation.
+users to transition to developers and contribute to a package. This also contributes
+to the following observation:
 
-\subsubsection{The code base is simple enough to be useful for students}
+\subsubsection{The code base is simple enough to be useful for new users}
 
@@
 work on toy problems since the remaining time is rather restricted.
+
+Another anecdotal example of ease of use is the preprint \cite{singh2021linear}. 
+While the authors utilized \trixi for some numerical experiments, they did so 
+independently of the development team (who learned that the authors were 
+using \trixi only \textit{after} reading the preprint!).
 
@@
 Many packages in the Julia ecosystem follow this approach and are freely available
-under the permissive MIT license, including \trixi. This allows using Julia
+under the permissive MIT license, including \trixi. This allows programmers to use Julia
 without needing to pay for commercial software. In particular, students can
@@
 in Julia to high performance computing applications still needs to be demonstrated.
-However, we do not expect unsolvable problems there since Julia is not generically
-slower than traditional compiled HPC programming languages. Nevertheless, it
-appears to be more complicated to scale Julia to HPC while still keeping the
+However, we do not expect this to remain an unsolvable problem since Julia is not 
+generically slower than traditional compiled HPC programming languages. Nevertheless, 
+it appears to be more complicated to scale Julia to HPC while still keeping the
 code as simple, clean, and flexible as it is now.
@@
 Learning a new programming language requires of course some effort and needs some
-time. Nevertheless, it pays off in our experience and we do not want to miss the
+time. Nevertheless, it has paid off in our experience and we do not want to miss the
 nice features of \trixi enabled by Julia and its ecosystem. Thus, while there is
@@
 
+Jesse Chan acknowledge support from the US National Science Foundation under 
+awards DMS-1719818 and DMS-1943186. 
 % References
@@
         list of main features?
-  \item Summary of community, collaborators, and users: GitHub statistics, contributors,
-        Vikram and Chandrashekar, Steven Frankel and Philipp Öffner (maybe), ...
-        Related to the relatively large number of students!
-        $\to$ Jesse will make a PR!
   \item A simple example where we really use multiple dispatch?

### 2021-08-03 paper/paper.tex
Update paper/paper.tex
Co-authored-by: Hendrik Ranocha <ranocha@users.noreply.github.com>
applying changes as discussed on the PR

@@
   \partial_t u(t, x) + \sum_{j=1}^d \partial_{x_j} f^j(u) = s(t, x, u),
-  \quad t \in (0, T),
+  \quad t \in (0, T), x \in \Omega, 
 \end{equation}
@@
 as fast as code written in these traditional scientific computing languages.
-We believe that it will be possible to get the same efficiency by rewriting
-FLUXO appropriately. However, the code introspection and profiling tools available
-in Julia make it easier for us to optimize Julia code compared to other languages
-we are familiar with (having a diverse team with a strong background in C, C++,
-and Fortran). \todo{from Jesse: what does this mean?}
+%We believe that it will be possible to get the same efficiency by rewriting
+%FLUXO appropriately. 
+Together, the members of our team have strong backgrounds in C, C++, and 
+Fortran. Nevertheless, our Julia code is faster than the established HPC Fortran 
+code FLUXO for this example. The reason is not that Fortran or Julia are generically 
+faster; in contrast, we expect to be able to achieve similar performance from either 
+language. Instead, \trixi owes its performance optimizations in part to the code 
+introspection and profiling tools available in Julia. Similar tools used to optimize the 
+performance in other languages are usually not as easy to use as their Julia 
+equivalents (if they exist at all).
 

### 2021-08-03 paper/paper.tex
JC paper edits (#5)
* paper edits

* Update paper/ref.bib

Co-authored-by: Hendrik Ranocha <ranocha@users.noreply.github.com>

* Update paper/paper.tex

Co-authored-by: Hendrik Ranocha <ranocha@users.noreply.github.com>

* applying changes as discussed on the PR

Co-authored-by: Jesse Chan <jesse.chan@rice.edu>
Co-authored-by: Hendrik Ranocha <ranocha@users.noreply.github.com>
fix URL

@@
 Roughly one year of collaborative work has resulted in the current version of
-\trixi\footnote{https://github.com/trixi-framework/Trixi.jl}, providing adaptive
+\trixi\footnote{\url{https://github.com/trixi-framework/Trixi.jl}}, providing adaptive
 high-order numerical simulations of hyperbolic partial differential equations (PDEs)
@@
 CFD codes implement their own time integration methods, arguing that the spatial
-part is much more complex \todo{should we cite something for this?}. While this is often true, it is still more efficient
+part is much more complex. While this is often true, it is still more efficient
 in terms of human workload to be able to reuse existing implementations. This
@@
 
-Jesse Chan acknowledge support from the US National Science Foundation under 
+Jesse Chan acknowledges support from the US National Science Foundation under 
 awards DMS-1719818 and DMS-1943186. 

### 2021-08-03 paper/paper.tex
minor edits to all sections (except 3) and bib

--- 2021-08-02 paper/paper.tex
@@
 Trixi.jl is extendable, easy to use, and fast. We describe the main design choices
-that enable these features and compare Trixi.jl with an established open
+that allow satisfying these goals and compare Trixi.jl with an established open
 source Fortran code for hyperbolic PDEs using the same numerical methods.
@@
 We are broadly interested in simulation-focused scientific computing, in particular
-in hyperbolic partial differential equations (PDEs), computational fluid dynamics (CFD),
+numerical approximations for hyperbolic partial differential equations (PDEs), computational fluid dynamics (CFD),
 and related problems. The focus of our research ranges from specific applications
-in CFD to general multi-physics coupling strategies to the high-order numerical methods 
-such simulations are built upon. In addition, we are academics and involved in teaching 
-students in these areas of science. Thus, we would like to have a 
-code that is
+in CFD over general multi-physics coupling strategies to the required high-order
+numerical methods themselves. In addition, we are academics and involved in teaching
+students in these areas of science. Thus, we would like to have a code that is
 \begin{enumerate}
@@
   \item easy to apply/understand to be useful for students/collaborators
-  \item fast enough to be useful beyond small toy problems
+  \item fast enough to be useful not only for small toy problems
 \end{enumerate}
-Additionally, we are greedy\footnote{See \url{https://julialang.org/blog/2012/02/why-we-created-julia/}} 
-and wish to include all these features within a single code.
+Additionally, we are greedy\footnote{See \url{https://julialang.org/blog/2012/02/why-we-created-julia/}} and want to have all these features in a single code.
 Roughly one year of collaborative work has resulted in the current version of
-\trixi\footnote{https://github.com/trixi-framework/Trixi.jl}, providing adaptive
+\trixi\footnote{https://github.com/trixi-framework/Trixi.jl}, which provides adaptive
 high-order numerical simulations of hyperbolic partial differential equations (PDEs)
@@
 to reach more and more of our goals with \trixi.
+
+\todo{TODO: Jesse wants to write about the \trixi community} %TODO: Jesse
 
@@
   \partial_t u(t, x) + \sum_{j=1}^d \partial_{x_j} f^j(u) = s(t, x, u),
-  \quad t \in (0, T),
+  \quad t \in (0, T), x \in \Omega,
 \end{equation}
@@
     \item Hierarchical quadtree/octree grid with adaptive refinement
-    \item Forests of quadtrees/octrees with \texttt{p4est} \cite{burstedde2011p4est}
+    \item Forests of qudtrees/octrees with \texttt{p4est} \cite{burstedde2011p4est}
           via P4est.jl\footnote{\url{https://github.com/trixi-framework/P4est.jl}}
@@
 
-  \item High-order matrix-free discontinuous Galerkin methods
+  \item High-order matrix-free discontinuous Galerkin (DG) methods
   \begin{itemize}
@@
     \item Lattice-Boltzmann equations (D2Q9 and D3Q27 schemes)
-    \item Several scalar conservation laws (advection, Burgers)
+    \item Several scalar conservation laws (\eg, advection, Burgers)
   \end{itemize}
@@
 
-\trixi is based on the method of lines. Thus, a discretization of \eqref{eq:hcl}
+\trixi is built from the method of lines. Thus, a discretization of \eqref{eq:hcl}
 is obtained in two steps. First, a spatial semidiscretization is created. Next,
@@
   \includegraphics[width=\linewidth]{../figures/trixi_global_overview}
-  \caption{Schematic overview of the basic components of \trixi and how they
-           interact with each other.}
+  \caption{Schematic overview of the basic components in \trixi and how they
+           interact.}
   \label{fig:trixi_global_overview}
@@
 components of \trixi. The most important structure is the semidiscretization,
-which bundles all information about the spatial part. The mathematical-physical
+which bundles all information about the spatial approximation. The mathematical-physical
 model is determined by the \lstinline{equations}, the \lstinline{initial_condition},
@@
 
-Once information about the time span $[0, T]$ is provided, a semidiscretization
-in \trixi can be converted into an ODE problem, which can be solved by methods 
-from OrdinaryDiffEq.jl. The flexible callback infrastructure of this ODE library
+Adding information about the time span $[0, T]$, a semidiscretization of \trixi
+can be converted into an ODE problem, which can be solved by methods from
+OrdinaryDiffEq.jl. The flexible callback infrastructure of this ODE library
 allows us to provide extended functionality for \trixi without modifying any
@@
 The design of \trixi was guided by extensibility, ease of use, and efficiency
-(in this particular order). Thus, it is likely possible to make (some parts) even 
-faster at the cost of simplicity and readibility. Nevertheless, \trixi is already quite
+(in this particular order). Thus, it will likely be possible to make (some parts)
+even faster while sacrificing simplicity. Nevertheless, \trixi is already quite
 fast, even compared to established high-performance open source software,
-as described in details in Section~\ref{sec:performance-comparison}. In this
+as described in detail in Section~\ref{sec:performance-comparison}. In this
 section, we focus mostly on the first two goals.
@@
 The most important design decision was to create \trixi as a library. In the
-context of simulation-focused scientific computing, there are a lot of (open
-or closed source) codes that are designed as monolithic applications that can
-simulate a specific setup. These codes can often be configured to a limited extent 
-by specifying compile time options or editing parameter files. In contrast, users of 
-\trixi do not need to compile anything or learn parameter file syntax. Instead, they can
-just load the library and write their simulation setups in Julia. Scripts which describe
-a simulation setup for \trixi are referred to as ``elixirs''. As of version v0.3.50, \trixi
-comes with \ca 200 example elixirs. A basic elixir might look as follows:
+context of simulation-focused scientific computing, there are many (open
+or closed source) codes designed as monolithic applications that can
+simulate a specific setup. Such monolithic codes can often be configured in limited amounts by
+compile time options and parameter files. In contrast, users of \trixi do not
+need to compile anything or learn the syntax of parameter files. They just load
+the library and write their simulation setups in Julia. Such scripts describing
+a simulation setup for \trixi are called elixirs. As of version v0.3.50, \trixi
+comes with \ca 200 example elixirs. A basic elixir could look as follows.
 % TODO: Update Trixi version above
@@
 Abstractions such as (variants of) the \lstinline{solver}, the \lstinline{mesh},
-and the \lstinline{equations} are based on a common interface. Using multiple
+and the \lstinline{equations} use a common interface. Utilizing multiple
 dispatch in Julia, internal implementations are specialized accordingly. For
@@
 passing one additional parameter to the \lstinline{DGSEM} constructor in the
-example above. Moreover, there is no hidden global state and multiple instances
+example above. Moreover, there is no hidden global state. This means that multiple instances
 of similar structures can be instantiated simultaneously. In particular, multiple
@@
 
-Instead of providing only the very basic discretization ingredients such as
-other open source frameworks for PDEs, \trixi includes also some widely used
-physical parts to make it easier to use. Nevertheless, users are not restricted
+Instead of providing only the very basic discretization ingredients, as with
+other open source frameworks for PDEs, \trixi also includes some widely used
+physical systems and analysis routines, such as computation of the integrated kinetic energy
+for the compressible Euler equations,
+to make it easier to use. Nevertheless, users are not restricted
 to the physics models bundled in \trixi. To setup a new type of \lstinline{equations},
@@
 library functions of \trixi. Thus, the physics can be separated completely from
-the \lstinline{solver} while retaining efficiency. In particular, this allows
-to reuse the same numerical fluxes for discontinuous Galerkin methods, finite
+the \lstinline{solver} but remain computationally efficient. In particular, this allows
+a user to reuse the same numerical fluxes for discontinuous Galerkin methods, finite
 difference methods, and variants of finite volume methods.
@@
 feature for all combinations of possible (compile time) options. This effort
-can make it hard to experiment with new ideas. In contrast, Julia's dynamic
+can make it difficult to experiment with new ideas. In contrast, Julia's dynamic
 nature, multiple dispatch, and just-ahead-of-time compilation allow us to implement only the features
@@
 to implement anything for curvilinear coordinates or shock capturing approaches.
-In addition, it allows to extend \trixi step by step with new capabilities.
+In addition, it allows a user to extend \trixi step by step with new capabilities.
 For example, there are ongoing efforts to incorporate summation by parts finite
 difference methods via SummationByPartsOperators.jl \cite{ranocha2021sbp} and
-discontinuous Galerkin methods on simplices via StartUpDG.jl\footnote{\url{https://github.com/jlchan/StartUpDG.jl}} 
-in \trixi. This becomes feasible since not every feature combination needs to be considered, 
-enabling developers to introduce variants of existing solvers on only a subset of the available
-mesh types via multiple dispatch.
+more discontinuous Galerkin methods on simplex elements via StartUpDG.jl\footnote{\url{https://github.com/jlchan/StartUpDG.jl}} in \trixi.
+This is feasible because the modular design of \trixi gives developers the flexibility to use
+a subset of the available meshes/methods, selected via multiple dispatch,
+to test their newly implemented feature(s).
 
@@
 we are familiar with (having a diverse team with a strong background in C, C++,
-and Fortran). \todo{from Jesse: what does this mean?}
+and Fortran).
 
@@
 \cite{olver2014practical}, and classical finite element methods \cite{badia2020gridap}.
-In addition, there are some packages dedicated to solving specific time-dependent
+In addition, there are some Julia packages dedicated to solve specific time-dependent
 PDEs such as \cite{ramadhan2020oceananigans, constantinou2021geophysicalflows}.
-However, there does not seem to be a general framework for high-order methods
-for hyperbolic PDEs in Julia. Thus, \trixi is an ideal candidate for a case study
+However, there does not seem to be a general framework of high-order methods
+or hyperbolic PDEs in Julia. Thus, \trixi is an ideal candidate for a case study
 of Julia for simulation-focused scientific computing, an area where codes written
@@
 languages such as C, C++, and Fortran. A minor exception of this rule is that
-Julia is build on LLVM, which does not perform the same optimizations as most
+Julia is built on LLVM, which does not perform the same optimizations as most
 optimizing C/C++/Fortran compilers in use nowadays. In particular, LLVM does
@@
 compilers do. Thus, one needs to opt-in to these optimizations explicitly in
-Julia. Luckily, that is relatively easy to do due to Julia's code manipulation
+Julia. Luckily, that is relatively easy due to Julia's code manipulation
 techniques which enable the macro \lstinline{@muladd}\footnote{\url{https://github.com/SciML/MuladdMacro.jl}}.
@@
 education in software development. Julia and its ecosystem support these
-researches by making it easy to set up unit and regression tests, since a
+researchers by making it easy to set up unit and regression tests, since a
 testing framework is included in the standard library and default package
@@
 driven development and continuous integration (CI), which makes code restructuring
-and optimization feasible.
+and optimization straightforward.
 
 Julia itself and most packages are developed on GitHub. The community provides
-a lot of tools for using corresponding CI setups, preparing and publishing
+many tools to use corresponding CI setups to prepare and publish
 documentation\footnote{\url{https://github.com/JuliaDocs/Documenter.jl}},
 and related tasks. This setup also encourages other good software development
-practices that we use in \trixi, such as mandatory code review before pull requests
+practices that we use in \trixi such as mandatory code review before pull requests
 are merged.
@@
 
-The package manager makes it easy to reproduce the exact environment, including
-binary dependencies. We have used this feature for all of our papers based on
+The package manager makes it easy to reproduce the exact run environment, including
+binary dependencies, used to generate numerical results. We have used this feature for all of our papers based on
 \trixi \cite{schlottkelakemper2021purely, ranocha2021preventing}, including this
@@
 This facilitates code sharing and reproducible research in computational science,
-which is arguably important but not mainstream yet \cite{barnes2010publish,
+which is arguably important but not yet mainstream \cite{barnes2010publish,
 donoho2010invitation, leveque2013top}.
@@
 
-We have written a Julia wrapper\footnote{\url{https://github.com/trixi-framework/P4est.jl}}
+To incorporate some of the adaptive mesh capabilities into \trixi
+we have written a Julia wrapper\footnote{\url{https://github.com/trixi-framework/P4est.jl}}
 of the C library \texttt{p4est} \cite{burstedde2011p4est}. In our experience,
@@
 In particular, users do not have to compile libraries on their own system
-and the binaries are available on all major platforms including Linux, Mac OS,
+and the binaries are available on all major platforms including Linux, macOS,
 and Windows.
@@
 
-\subsubsection{Packages can really be used together at low (or no) cost}
+\subsubsection{Packages can be used together at low (or no) cost}
 
@@
 complexity of handling different versions and making everything work together.
-Such an approach typically leads to significant code duplication. For example, many
+Such an approach often leads to a lot of code duplication. For example, many
 CFD codes implement their own time integration methods, arguing that the spatial
-part is much more complex \todo{should we cite something for this?}. While this is often true, it is still more efficient
+part is much more complex. While this is often true, it is still more efficient
 in terms of human workload to be able to reuse existing implementations. This
-allows expert on time integration methods to develop specialized algorithms and
+allows experts on time integration methods to develop specialized algorithms and
 implement them in open source software while practitioners or researchers focusing
@@
 of code. For example, optimized time integration methods were developed in
-\cite{ranocha2021optimized} and implemented in OrdinaryDiffEq.jl. These methods
-can be used with \trixi by changing a single line of code. In contrast, researchers
-working without external dependencies would need to spend a day or two to digest
-details before even beginning implementation. This additional overhead in developer
-time makes it less likely that researchers would implement these methods in their own 
-codes and benefit from recent algorithmic developments.
+\cite{ranocha2021optimized} and implemented in OrdinaryDiffEq.jl. Thus, they
+can be used with \trixi by changing a single line of code in an elixir file. In contrast, researchers
+working without external dependencies would need to use a day or two to digest
+all the details and implement such time integration methods in their own codes, making it much
+less likely that they will do this and benefit from recent algorithmic developments.
 
@@
 Such a difference between the programming languages used in the front- and
-back-ends lead to a well-known barrier between user and developers of a library.
+back-ends lead to a well-known barrier between a user and the developers of a library.
 Julia was designed from the beginning to make it possible to write both simple
 high-level code and highly efficient compute kernels, thereby solving the
-two-language problem --- at least to a significant extent. Highly efficient low-level
+two-language problem --- at least to a large extent. Highly efficient low-level
 code will of course still look different from simple high-level code, but the
 barrier between them is much smaller. In particular, this makes it easier for
-users to transition to developers and contribute to a package. This also contributes
-to the following observation:
+users to contribute to a package and act as developers. This also contributes
+to the following observation.
 
-\subsubsection{The code base is simple enough to be useful for new users}
+\subsubsection{The code base is simple enough to be useful for students}
 
@@
 work on toy problems since the remaining time is rather restricted.
-
-Another anecdotal example of ease of use is the preprint \cite{singh2021linear}. 
-While the authors utilized \trixi for some numerical experiments, they did so 
-independently of the development team (who learned that the authors were 
-using \trixi only \textit{after} reading the preprint!).
 
@@
 Many packages in the Julia ecosystem follow this approach and are freely available
-under the permissive MIT license, including \trixi. This allows programmers to use Julia
-without needing to pay for commercial software. In particular, students can
+under the permissive MIT license, including \trixi. This allows using Julia
+without the need to pay for commercial software. In particular, students can
 work with the software on their private computers without restrictions.
@@
 
-Since Julia uses a serial just-ahead-of-time compiler that does currently not cache
+Since Julia uses a serial just-ahead-of-time compiler that currently does not cache
 code between different Julia sessions, the initial compilation time can be
-annoying. For example, the time to finish the first simulation in a fresh Julia
+annoying. For example, the time to finish a first \trixi simulation in a fresh Julia
 session and plot the results on a notebook can take between 30 seconds and a
-minute. Having compiled the code, the second simulation and plot take less than
+minute. Having compiled the code, the second simulation and plot takes less than
 0.05 seconds. Thus, switching to Julia requires adapting the workflow compared
@@
 the number of students and researchers working with \trixi, and serial performance
-comparisons with established Fortran codes. Having worked on \trixi for a bit
+comparisons with an established Fortran code. Having developed, from scratch, \trixi for a bit
 more than one year allows us to give an assessment of Julia for simulation-focused
@@
 in Julia to high performance computing applications still needs to be demonstrated.
-However, we do not expect this to remain an unsolvable problem since Julia is not 
-generically slower than traditional compiled HPC programming languages. Nevertheless, 
-it appears to be more complicated to scale Julia to HPC while still keeping the
+However, we do not expect unsolvable problems there, as Julia is not generically
+slower than traditional compiled HPC programming languages. Nevertheless, it
+appears to be more complicated to scale Julia to HPC while still keeping the
 code as simple, clean, and flexible as it is now.
@@
 Learning a new programming language requires of course some effort and needs some
-time. Nevertheless, it has paid off in our experience and we do not want to miss the
+time. Nevertheless, it pays off in our experience and we do not want to miss the
 nice features of \trixi enabled by Julia and its ecosystem. Thus, while there is
@@
 
-Jesse Chan acknowledge support from the US National Science Foundation under 
-awards DMS-1719818 and DMS-1943186. 
 % References
@@
         list of main features?
+  \item Summary of community, collaborators, and users: GitHub statistics, contributors,
+        Vikram and Chandrashekar, Steven Frankel and Philipp Öffner (maybe), ...
+        Related to the relatively large number of students!
+        $\to$ Jesse will make a PR!
   \item A simple example where we really use multiple dispatch?

### 2021-08-03 paper/paper.tex

@@
   \item easy to apply/understand to be useful for students/collaborators
-  \item fast enough to be useful not only for small toy problems
+  \item fast enough to be useful beyond small toy problems
 \end{enumerate}
-Additionally, we are greedy\footnote{See \url{https://julialang.org/blog/2012/02/why-we-created-julia/}} and want to have all these features in a single code.
+Additionally, we are greedy\footnote{See \url{https://julialang.org/blog/2012/02/why-we-created-julia/}} 
+and wish to include all these features within a single code.
 Roughly one year of collaborative work has resulted in the current version of
-\trixi\footnote{https://github.com/trixi-framework/Trixi.jl}, which provides adaptive
+\trixi\footnote{\url{https://github.com/trixi-framework/Trixi.jl}}, providing adaptive
 high-order numerical simulations of hyperbolic partial differential equations (PDEs)
@@
 context of simulation-focused scientific computing, there are many (open
-or closed source) codes designed as monolithic applications that can
-simulate a specific setup. Such monolithic codes can often be configured in limited amounts by
-compile time options and parameter files. In contrast, users of \trixi do not
-need to compile anything or learn the syntax of parameter files. They just load
-the library and write their simulation setups in Julia. Such scripts describing
-a simulation setup for \trixi are called elixirs. As of version v0.3.50, \trixi
-comes with \ca 200 example elixirs. A basic elixir could look as follows.
+or closed source) codes that are designed as monolithic applications that can
+simulate a specific setup. These codes can often be configured to a limited extent 
+by specifying compile time options or editing parameter files. In contrast, users of 
+\trixi do not need to compile anything or learn parameter file syntax. Instead, they can
+just load the library and write their simulation setups in Julia. Scripts which describe
+a simulation setup for \trixi are referred to as ``elixirs''. As of version v0.3.50, \trixi
+comes with \ca 200 example elixirs. A basic elixir might look as follows:
 % TODO: Update Trixi version above
@@
 difference methods via SummationByPartsOperators.jl \cite{ranocha2021sbp} and
-more discontinuous Galerkin methods on simplex elements via StartUpDG.jl\footnote{\url{https://github.com/jlchan/StartUpDG.jl}} in \trixi.
-This is feasible because the modular design of \trixi gives developers the flexibility to use
+discontinuous Galerkin methods on simplex elements via StartUpDG.jl\footnote{\url{https://github.com/jlchan/StartUpDG.jl}} in \trixi.
+This is feasible because the modular design of \trixi gives users the flexibility to select
 a subset of the available meshes/methods, selected via multiple dispatch,
@@
 as fast as code written in these traditional scientific computing languages.
-We believe that it will be possible to get the same efficiency by rewriting
-FLUXO appropriately. However, the code introspection and profiling tools available
-in Julia make it easier for us to optimize Julia code compared to other languages
-we are familiar with (having a diverse team with a strong background in C, C++,
-and Fortran).
+%We believe that it will be possible to get the same efficiency by rewriting
+%FLUXO appropriately. 
+Together, the members of our team have strong backgrounds in C, C++, and 
+Fortran. Nevertheless, our Julia code is faster than the established HPC Fortran 
+code FLUXO for this example. The reason is not that Fortran or Julia are generically 
+faster; in contrast, we expect to be able to achieve similar performance from either 
+language. Instead, \trixi owes its performance optimizations in part to the code 
+introspection and profiling tools available in Julia. Similar tools used to optimize the 
+performance in other languages are usually not as easy to use as their Julia 
+equivalents (if they exist at all).
 
@@
 \cite{olver2014practical}, and classical finite element methods \cite{badia2020gridap}.
-In addition, there are some Julia packages dedicated to solve specific time-dependent
+In addition, there are some packages dedicated to solving specific time-dependent
 PDEs such as \cite{ramadhan2020oceananigans, constantinou2021geophysicalflows}.
-However, there does not seem to be a general framework of high-order methods
-or hyperbolic PDEs in Julia. Thus, \trixi is an ideal candidate for a case study
+However, there does not seem to be a general framework for high-order methods
+for hyperbolic PDEs in Julia. Thus, \trixi is an ideal candidate for a case study
 of Julia for simulation-focused scientific computing, an area where codes written
@@
 complexity of handling different versions and making everything work together.
-Such an approach often leads to a lot of code duplication. For example, many
+Such an approach typically leads to significant code duplication. For example, many
 CFD codes implement their own time integration methods, arguing that the spatial
@@
 of code. For example, optimized time integration methods were developed in
-\cite{ranocha2021optimized} and implemented in OrdinaryDiffEq.jl. Thus, they
-can be used with \trixi by changing a single line of code in an elixir file. In contrast, researchers
-working without external dependencies would need to use a day or two to digest
-all the details and implement such time integration methods in their own codes, making it much
-less likely that they will do this and benefit from recent algorithmic developments.
+\cite{ranocha2021optimized} and implemented in OrdinaryDiffEq.jl. These methods
+can be used with \trixi by changing a single line of code. In contrast, researchers
+working without external dependencies would need to spend a day or two to digest
+details before even beginning implementation. This additional overhead in developer
+time makes it less likely that researchers would implement these methods in their own 
+codes and benefit from recent algorithmic developments.
 
@@
 high-level code and highly efficient compute kernels, thereby solving the
-two-language problem --- at least to a large extent. Highly efficient low-level
+two-language problem --- at least to a significant extent. Highly efficient low-level
 code will of course still look different from simple high-level code, but the
 barrier between them is much smaller. In particular, this makes it easier for
-users to contribute to a package and act as developers. This also contributes
-to the following observation.
+users to transition to developers and contribute to a package. This also contributes
+to the following observation:
 
-\subsubsection{The code base is simple enough to be useful for students}
+\subsubsection{The code base is simple enough to be useful for new users}
 
@@
 Many packages in the Julia ecosystem follow this approach and are freely available
-under the permissive MIT license, including \trixi. This allows using Julia
-without the need to pay for commercial software. In particular, students can
+under the permissive MIT license, including \trixi. This allows programmers to use Julia
+without needing to pay for commercial software. In particular, students can
 work with the software on their private computers without restrictions.
@@
 in Julia to high performance computing applications still needs to be demonstrated.
-However, we do not expect unsolvable problems there, as Julia is not generically
-slower than traditional compiled HPC programming languages. Nevertheless, it
-appears to be more complicated to scale Julia to HPC while still keeping the
+However, we do not expect this to remain an unsolvable problem since Julia is not 
+generically slower than traditional compiled HPC programming languages. Nevertheless, 
+it appears to be more complicated to scale Julia to HPC while still keeping the
 code as simple, clean, and flexible as it is now.
@@
 
+Jesse Chan acknowledges support from the US National Science Foundation under 
+awards DMS-1719818 and DMS-1943186. 
 % References

### 2021-08-03 paper/paper.tex
add DG and entropy references

--- 2021-08-03 paper/paper.tex
@@
 % TODO: Update Trixi version below
-As of version v0.3.50 (July 2021), \trixi comes with the following features.
+As of version v0.3.50 (July 2021), \trixi concentrates mainly on discontinuous
+Galerkin (DG) methods \cite{hesthaven2007nodal, kopriva2009implementing}.
+In particular, it has a focus on entropy-conservative and -dissipative methods
+\cite{tadmor1987numerical, lefloch2002fully, fisher2013discretely,
+ranocha2018comparison, chen2017entropy}. Currently, \trixi comes with the
+following features.
 \begin{itemize}

### 2021-08-03 paper/paper.tex
add KHI figures

@@
 \end{itemize}
+
+Some of the main features of \trixi are demonstrated in the following figures.
+Figure~\ref{fig:kelvin_helmholtzs} demonstrates entropy-dissipative shock-capturing
+methods with adaptive mesh refinement applied to a classical Kelvin-Helmholtz
+flow instability problem.
+
+\begin{figure}[htbp]
+  \begin{subfigure}{0.53\linewidth}
+    \includegraphics[width=\textwidth]{../figures/kelvin_helmholtz_density_t2}
+    \caption{Density at time $t = 2$.}
+  \end{subfigure}%
+  \hspace*{\fill}
+  \begin{subfigure}{0.47\linewidth}
+    \includegraphics[width=\textwidth]{../figures/kelvin_helmholtz_mesh_t2}
+    \caption{Mesh at time $t = 2$.}
+  \end{subfigure}%
+  \\
+  \begin{subfigure}{0.53\linewidth}
+    \includegraphics[width=\textwidth]{../figures/kelvin_helmholtz_density_t3}
+    \caption{Density at time $t = 3$.}
+  \end{subfigure}%
+  \hspace*{\fill}
+  \begin{subfigure}{0.47\linewidth}
+    \includegraphics[width=\textwidth]{../figures/kelvin_helmholtz_mesh_t3}
+    \caption{Mesh at time $t = 3$.}
+  \end{subfigure}%
+  \caption{Numerical solutions of a Kelvin-Helmholtz instability using
+           entropy-dissipative shock-capturing methods and adaptive mesh
+           refinement for the compressible Euler equations.}
+  \label{fig:kelvin_helmholtzs}
+\end{figure}
+
+\todo{TODO: Andrew, add gingerbread man example for acoustics} %TODO: Andrew
+
+The detailed physical setups and numerical parameters as well as all code
+necessary to reproduce these figures are available in the reproducibility
+repository for this article \cite{ranocha2021adaptiveRepro}.
 

### 2021-08-03 figure_kelvin_helmholtz/elixir_euler_kelvin_helmholtz_instability_amr.jl

@@ -1,0 +1,110 @@
+
+using OrdinaryDiffEq
+using Trixi
+
+###############################################################################
+# semidiscretization of the compressible Euler equations
+gamma = 1.4
+equations = CompressibleEulerEquations2D(gamma)
+
+"""
+    initial_condition(x, t, equations::CompressibleEulerEquations2D)
+
+A version of the classical Kelvin-Helmholtz instability based on
+- Andrés M. Rueda-Ramírez, Gregor J. Gassner (2021)
+  A Subcell Finite Volume Positivity-Preserving Limiter for DGSEM Discretizations
+  of the Euler Equations
+  [arXiv: 2102.06017](https://arxiv.org/abs/2102.06017)
+"""
+function initial_condition(x, t, equations::CompressibleEulerEquations2D)
+  # change discontinuity to tanh
+  # typical resolution 128^2, 256^2
+  # domain size is [-1,+1]^2
+  slope = 15
+  amplitude = 0.02
+  B = tanh(slope * x[2] + 7.5) - tanh(slope * x[2] - 7.5)
+  rho = 0.5 + 0.75 * B
+  v1 = 0.5 * (B - 1)
+  v2 = 0.1 * sin(2 * pi * x[1])
+  p = 1.0
+  return prim2cons(SVector(rho, v1, v2, p), equations)
+end
+
+surface_flux = flux_lax_friedrichs
+volume_flux  = flux_chandrashekar
+polydeg = 3
+basis = LobattoLegendreBasis(polydeg)
+indicator_sc = IndicatorHennemannGassner(equations, basis,
+                                         alpha_max=0.002,
+                                         alpha_min=0.0001,
+                                         alpha_smooth=true,
+                                         variable=density_pressure)
+volume_integral = VolumeIntegralShockCapturingHG(indicator_sc;
+                                                 volume_flux_dg=volume_flux,
+                                                 volume_flux_fv=surface_flux)
+solver = DGSEM(basis, surface_flux, volume_integral)
+
+coordinates_min = (-1.0, -1.0)
+coordinates_max = ( 1.0,  1.0)
+mesh = TreeMesh(coordinates_min, coordinates_max,
+                initial_refinement_level=5,
+                n_cells_max=100_000)
+
+
+semi = SemidiscretizationHyperbolic(mesh, equations, initial_condition, solver)
+
+
+###############################################################################
+# ODE solvers, callbacks etc.
+
[…]

### 2021-08-04 paper/paper.tex
add colliding flows example

@@
     \item Compressible Euler equations (optionally with self-gravity)
-    \item Magnetohydrodynamics equations
+    \item Magnetohydrodynamics (MHD) equations
+    \item Multicomponent compressible Euler and MHD equations
     \item Acoustic perturbation equations
@@
   \label{fig:kelvin_helmholtzs}
+\end{figure}
+
+Figure~\ref{fig:colliding_flows} demonstrates entropy-dissipative shock-capturing
+methods with positivity-preserving limiters and adaptive mesh refinement applied
+to supersonic colliding flows inspired by astrophysical setups.
+
+\begin{figure}[htbp]
+  \begin{subfigure}{0.53\linewidth}
+    \includegraphics[width=\textwidth]{../figures/colliding_flows_density_t15}
+    \caption{Density at time $t = 15$.}
+  \end{subfigure}%
+  \hspace*{\fill}
+  \begin{subfigure}{0.47\linewidth}
+    \includegraphics[width=\textwidth]{../figures/colliding_flows_mesh_t15}
+    \caption{Mesh at time $t = 15$.}
+  \end{subfigure}%
+  \\
+  \begin{subfigure}{0.53\linewidth}
+    \includegraphics[width=\textwidth]{../figures/colliding_flows_density_t25}
+    \caption{Density at time $t = 25$.}
+  \end{subfigure}%
+  \hspace*{\fill}
+  \begin{subfigure}{0.47\linewidth}
+    \includegraphics[width=\textwidth]{../figures/colliding_flows_mesh_t25}
+    \caption{Mesh at time $t = 25$.}
+  \end{subfigure}%
+  \caption{Numerical solutions of two colliding supersonic flows using
+           entropy-dissipative shock-capturing methods, positivity-preserving
+           limiters, and adaptive mesh refinement for the compressible Euler
+           equations.}
+  \label{fig:colliding_flows}
 \end{figure}

### 2021-08-04 figure_colliding_flows/elixir_euler_colliding_flow_amr.jl

@@ -1,0 +1,124 @@
+
+using OrdinaryDiffEq
+using Trixi
+
+###############################################################################
+# semidiscretization of the compressible Euler equations
+
+gamma = 1.001 # almost isothermal when gamma reaches 1
+equations = CompressibleEulerEquations2D(gamma)
+
+# This is a hand made colliding flow setup without reference. Features Mach=70 inflow from both
+# sides, with relative low temperature, such that pressure keeps relatively small
+# Computed with gamma close to 1, to simulate isothermal gas
+function initial_condition(x, t, equations::CompressibleEulerEquations2D)
+  # change discontinuity to tanh
+  # resolution 128^2 elements (refined close to the interface) and N=3 (total of 512^2 DOF)
+  # domain size is [-64,+64]^2
+  @unpack gamma = equations
+  rho = 0.0247
+  c = 0.2
+  p = c^2 / gamma * rho
+  vel = 13.907432274789372
+  slope = 1.0
+  v1 = -vel * tanh(slope * x[1])
+  # add small initial disturbance to the field, but only close to the interface
+  if abs(x[1]) < 10
+    v1 = v1 * (1 + 0.01 * sin(pi * x[2]))
+  end
+  v2 = 0.0
+  return prim2cons(SVector(rho, v1, v2, p), equations)
+end
+
+
+boundary_conditions = (x_neg=BoundaryConditionDirichlet(initial_condition),
+                       x_pos=BoundaryConditionDirichlet(initial_condition),
+                       y_neg=boundary_condition_periodic,
+                       y_pos=boundary_condition_periodic)
+
+
+
+surface_flux = flux_lax_friedrichs
+volume_flux  = flux_ranocha
+polydeg = 3
+basis = LobattoLegendreBasis(polydeg)
+
+# shock capturing necessary for this tough example, however alpha_max = 0.5 is fine
+indicator_sc = IndicatorHennemannGassner(equations, basis,
+                                         alpha_max=0.5,
+                                         alpha_min=0.0001,
+                                         alpha_smooth=true,
+                                         variable=density_pressure)
+volume_integral = VolumeIntegralShockCapturingHG(indicator_sc;
+                                                 volume_flux_dg=volume_flux,
+                                                 volume_flux_fv=surface_flux)
+solver = DGSEM(basis, surface_flux, volume_integral)
+
+coordinates_min = (-64.0, -64.0)
+coordinates_max = ( 64.0,  64.0)
+
[…]

### 2021-08-04 figure_kelvin_helmholtz/elixir_euler_kelvin_helmholtz_instability_amr.jl

@@ -106,5 +106,5 @@ plot(pd["rho"])
 plot!(title="", colorbar_title="", size=(650, 500); fontsizes...)
 savefig(joinpath(figdir, "kelvin_helmholtz_density_t" * string(round(Int, last(tspan))) * ".pdf"))
 
-plot(getmesh(pd), xlabel="\$x\$", ylabel="\$y\$", size=(580, 500), linewidth=2; fontsizes...)
+plot(getmesh(pd), xlabel="\$x\$", ylabel="\$y\$", size=(580, 500), linewidth=2, grid=false; fontsizes...)
 savefig(joinpath(figdir, "kelvin_helmholtz_mesh_t" * string(round(Int, last(tspan))) * ".pdf"))

### 2021-08-04 figure_colliding_flows/elixir_euler_colliding_flow_amr.jl
remove colliding flows setup since it does not look that impressive for a general audience

@@ -1,124 +1,0 @@
-
-using OrdinaryDiffEq
-using Trixi
-
-###############################################################################
-# semidiscretization of the compressible Euler equations
-
-gamma = 1.001 # almost isothermal when gamma reaches 1
-equations = CompressibleEulerEquations2D(gamma)
-
-# This is a hand made colliding flow setup without reference. Features Mach=70 inflow from both
-# sides, with relative low temperature, such that pressure keeps relatively small
-# Computed with gamma close to 1, to simulate isothermal gas
-function initial_condition(x, t, equations::CompressibleEulerEquations2D)
-  # change discontinuity to tanh
-  # resolution 128^2 elements (refined close to the interface) and N=3 (total of 512^2 DOF)
-  # domain size is [-64,+64]^2
-  @unpack gamma = equations
-  rho = 0.0247
-  c = 0.2
-  p = c^2 / gamma * rho
-  vel = 13.907432274789372
-  slope = 1.0
-  v1 = -vel * tanh(slope * x[1])
-  # add small initial disturbance to the field, but only close to the interface
-  if abs(x[1]) < 10
-    v1 = v1 * (1 + 0.01 * sin(pi * x[2]))
-  end
-  v2 = 0.0
-  return prim2cons(SVector(rho, v1, v2, p), equations)
-end
-
-
-boundary_conditions = (x_neg=BoundaryConditionDirichlet(initial_condition),
-                       x_pos=BoundaryConditionDirichlet(initial_condition),
-                       y_neg=boundary_condition_periodic,
-                       y_pos=boundary_condition_periodic)
-
-
-
-surface_flux = flux_lax_friedrichs
-volume_flux  = flux_ranocha
-polydeg = 3
-basis = LobattoLegendreBasis(polydeg)
-
-# shock capturing necessary for this tough example, however alpha_max = 0.5 is fine
-indicator_sc = IndicatorHennemannGassner(equations, basis,
-                                         alpha_max=0.5,
-                                         alpha_min=0.0001,
-                                         alpha_smooth=true,
-                                         variable=density_pressure)
-volume_integral = VolumeIntegralShockCapturingHG(indicator_sc;
-                                                 volume_flux_dg=volume_flux,
-                                                 volume_flux_fv=surface_flux)
-solver = DGSEM(basis, surface_flux, volume_integral)
-
-coordinates_min = (-64.0, -64.0)
-coordinates_max = ( 64.0,  64.0)
-
[…]

### 2021-08-04 paper/paper.tex
remove colliding flows setup since it does not look that impressive for a general audience

--- 2021-08-03 paper/paper.tex
@@
 We are broadly interested in simulation-focused scientific computing, in particular
-in hyperbolic partial differential equations (PDEs), computational fluid dynamics (CFD),
+numerical approximations for hyperbolic partial differential equations (PDEs), computational fluid dynamics (CFD),
 and related problems. The focus of our research ranges from specific applications
@@
 fast, even compared to established high-performance open source software,
-as described in details in Section~\ref{sec:performance-comparison}. In this
+as described in detail in Section~\ref{sec:performance-comparison}. In this
 section, we focus mostly on the first two goals.
@@
 The most important design decision was to create \trixi as a library. In the
-context of simulation-focused scientific computing, there are a lot of (open
+context of simulation-focused scientific computing, there are many (open
 or closed source) codes that are designed as monolithic applications that can
@@
 to implement anything for curvilinear coordinates or shock capturing approaches.
-In addition, it allows to extend \trixi step by step with new capabilities.
+In addition, it allows a user to extend \trixi step by step with new capabilities.
 For example, there are ongoing efforts to incorporate summation by parts finite
 difference methods via SummationByPartsOperators.jl \cite{ranocha2021sbp} and
-discontinuous Galerkin methods on simplices via StartUpDG.jl\footnote{\url{https://github.com/jlchan/StartUpDG.jl}} 
-in \trixi. This becomes feasible since not every feature combination needs to be considered, 
-enabling developers to introduce variants of existing solvers on only a subset of the available
-mesh types via multiple dispatch.
+discontinuous Galerkin methods on simplex elements via StartUpDG.jl\footnote{\url{https://github.com/jlchan/StartUpDG.jl}} in \trixi.
+This is feasible because the modular design of \trixi gives users the flexibility to select
+a subset of the available meshes/methods, selected via multiple dispatch,
+to test their newly implemented feature(s).
 
@@
 driven development and continuous integration (CI), which makes code restructuring
-and optimization feasible.
+and optimization straightforward.
 
 Julia itself and most packages are developed on GitHub. The community provides
-a lot of tools for using corresponding CI setups, preparing and publishing
+many tools to use corresponding CI setups to prepare and publish
 documentation\footnote{\url{https://github.com/JuliaDocs/Documenter.jl}},
@@
 in terms of human workload to be able to reuse existing implementations. This
-allows expert on time integration methods to develop specialized algorithms and
+allows experts on time integration methods to develop specialized algorithms and
 implement them in open source software while practitioners or researchers focusing
@@
 Such a difference between the programming languages used in the front- and
-back-ends lead to a well-known barrier between user and developers of a library.
+back-ends lead to a well-known barrier between a user and the developers of a library.
 Julia was designed from the beginning to make it possible to write both simple

### 2021-08-04 paper/paper.tex
update TODO notes

@@
 
-\todo{TODO: Be careful how to formulate this, be neutral and avoid making people angry} %TODO
+\todo{TODO: Gregor. Be careful how to formulate this, be neutral and avoid making people angry} %TODO: Gregor
 
@@
   \item Prepare repro repo including scripts for \trixi and FLUXO
-  \item Restructure non-conservative stuff and add MHD benchmarks
+  \item Rerun Euler benchmarks and add MHD benchmarks
   \item Shall we re-create the schematic overview in tikz?
-  \item Shall we compare \trixi (in terms of structure, not performance!)
-        to other open source frameworks such as MFEM,
-        NGSolve, DealII, DUNE, FENICS, Firedrake, ClawPack, ...? Definitely not
-        explicitly to not make people angry, but maybe implicitly without
-        mentioning their names?
-        \begin{itemize}
-          \item Gregor will think about it!
-        \end{itemize}
+  \item Gregor: Other aspects of other codes that we like more in \trixi?
   \item Are there other aspects we should discuss?
   \item Check affiliations
-  \item Check order of authors
-  \item Maybe include some cool pictures? Give a flavor of the kind of simulations
-        you can do with \trixi? Pictures trigger people! For example after the
-        list of main features?
-  \item A simple example where we really use multiple dispatch?
-        \begin{itemize}
-          \item Just pass functions around
-          \item Dispatch on parameters of DG struct
-        \end{itemize}
 \end{itemize}

### 2021-08-04 paper/paper.tex
do not introduce DG twice
reformulate CI stuff

@@
 
-  \item High-order matrix-free discontinuous Galerkin (DG) methods
+  \item High-order matrix-free discontinuous Galerkin methods
   \begin{itemize}
@@
 Julia itself and most packages are developed on GitHub. The community provides
-many tools to use corresponding CI setups to prepare and publish
+many tools one can use to run corresponding CI setups, prepare and publish
 documentation\footnote{\url{https://github.com/JuliaDocs/Documenter.jl}},
-and related tasks. This setup also encourages other good software development
+and other related tasks. This setup also encourages other good software development
 practices that we use in \trixi, such as mandatory code review before pull requests

### 2021-08-04 paper/paper.tex
update Trixi version

@@
 % TODO: Update Trixi version below
-As of version v0.3.50 (July 2021), \trixi concentrates mainly on discontinuous
+As of version v0.3.53 (August 2021), \trixi concentrates mainly on discontinuous
 Galerkin (DG) methods \cite{hesthaven2007nodal, kopriva2009implementing}.
@@
 just load the library and write their simulation setups in Julia. Scripts which describe
-a simulation setup for \trixi are referred to as ``elixirs''. As of version v0.3.50, \trixi
-comes with \ca 200 example elixirs. A basic elixir might look as follows:
+a simulation setup for \trixi are referred to as ``elixirs''. As of version
+v0.3.53 (August 2021), \trixi comes with \ca 200 example elixirs. A basic elixir
+might look as follows:
 % TODO: Update Trixi version above

### 2021-08-05 paper/paper.tex
fix typo
add Manifest and make some minor tweaks (when abbreviations are introduced etc.)

@@
 \trixi\footnote{\url{https://github.com/trixi-framework/Trixi.jl}}, providing adaptive
-high-order numerical simulations of hyperbolic partial differential equations (PDEs)
+high-order numerical simulations of hyperbolic PDEs
 in Julia \cite{bezanson2017julia}. Starting as an experiment, we have been able
@@
 and easy-to-use framework of high-order methods for hyperbolic PDEs. Next, we
-compare the serial performance with an established HPC Fortran code in
+compare the serial performance with an established high-performance computing (HPC) Fortran code in
 Section~\ref{sec:performance-comparison}, demonstrating that Julia is not generically
@@
 computing based on our experience with \trixi in Section~\ref{sec:assessment-of-julia}.
-Finally, we summary our findings and conclusions in Section~\ref{sec:summary}.
+Finally, we summarize our findings and conclusions in Section~\ref{sec:summary}.
 
@@
     \item Kinetic energy preserving and entropy-stable methods
-    \item Entropy-stable shock-capturing
+    \item Entropy-stable shock capturing
     \item Positivity-preserving limiting
@@
     \item In-situ visualization with Plots.jl
-    \item Postprocessing with ParaView/Visit via Trixi2Vtk.jl
+    \item Postprocessing with ParaView/Visit via Trixi2Vtk.jl\footnote{\url{https://github.com/trixi-framework/Trixi2Vtk.jl}}
   \end{itemize}
@@
 discretization method such as discontinuous Galerkin or finite difference
-methods, kinetic energy preserving or shock capturing approaches. Finally, the
+methods, kinetic energy preserving or shock-capturing approaches. Finally, the
 \lstinline{mesh} has a necessarily hybrid role including information about the
@@
 (in this particular order). Thus, it is likely possible to make (some parts) even 
-faster at the cost of simplicity and readibility. Nevertheless, \trixi is already quite
+faster at the cost of simplicity and readability. Nevertheless, \trixi is already quite
 fast, even compared to established high-performance open source software,
@@
 example, changing the volume terms of the DG discretization from the standard
-weak form to entropy stable and/or shock capturing methods can be achieved by
+weak form to entropy-stable and/or shock-capturing methods can be achieved by
 passing one additional parameter to the \lstinline{DGSEM} constructor in the
@@
 physics model only on Cartesian grids with smooth solutions, there is no need
-to implement anything for curvilinear coordinates or shock capturing approaches.
+to implement anything for curvilinear coordinates or shock-capturing approaches.
 In addition, it allows a user to extend \trixi step by step with new capabilities.
@@
 to achieve entropy conservation/dissipation or kinetic energy preservation.
-At the same time, both codes also have features the other one does not, \eg
+At the same time, both codes also have features the other one does not, \eg,
 parabolic equations and MPI in FLUXO or multiple mesh types and more physics setups
@@
 agreement 2020-03642 VR.
+Jesse Chan acknowledges support from the US National Science Foundation under
+awards DMS-1719818 and DMS-1943186.
+
+\todo{TODO: Gregor, add your funding sources that need to be here} %TODO Gregor
 
@@
 
-\todo{TODO: Add your funding sources that need to be here (if any)} %TODO Funding
-
-Jesse Chan acknowledges support from the US National Science Foundation under 
-awards DMS-1719818 and DMS-1943186. 
 % References
@@
   \item Rerun Euler benchmarks and add MHD benchmarks
-  \item Shall we re-create the schematic overview in tikz?
   \item Gregor: Other aspects of other codes that we like more in \trixi?

### 2021-08-05 figure_kelvin_helmholtz/elixir_euler_kelvin_helmholtz_instability_amr.jl

@@ -102,8 +102,7 @@ figdir = joinpath(dirname(@__DIR__), "figures")
 fontsizes = (xtickfontsize=18, ytickfontsize=18, xguidefontsize=20, yguidefontsize=20, legendfontsize=18)
 
 pd = PlotData2D(sol)
-plot(pd["rho"])
-plot!(title="", colorbar_title="", size=(650, 500); fontsizes...)
+plot(pd["rho"], title="", colorbar_title="", size=(650, 500), clim=(0.47, 2.3); fontsizes...)
 savefig(joinpath(figdir, "kelvin_helmholtz_density_t" * string(round(Int, last(tspan))) * ".pdf"))
 
 plot(getmesh(pd), xlabel="\$x\$", ylabel="\$y\$", size=(580, 500), linewidth=2, grid=false; fontsizes...)

### 2021-08-05 paper/paper.tex
plot relative performance of the TreeMEsh

@@
   \caption{Run time per right-hand side evaluation and degree of freedom for
-           different DG discretizations of the 3D compressible Euler equations.}
+           different DG discretizations of the 3D compressible Euler equations
+           in FLUXO (Fortran) and \trixi (Julia).}
   \label{fig:PID-Euler}
@@
 
-\todo{TODO: Plot of this comparison with the \lstinline{TreeMesh}?}
+\todo{TODO: Keep the plot of this comparison with the \lstinline{TreeMesh}?}
+
+\begin{figure}
+\centering
+  \begin{subfigure}{\linewidth}
+    \begin{tikzpicture}[
+      font=\footnotesize
+      ]
+      \begin{axis}[
+          xlabel={Polynomial degree},
+          % ylabel={Time per RHS call and DOF (relative)},
+          ylabel={Relative run time {[-]}},
+          width=\textwidth,
+          height=0.66\textwidth,
+          legend columns=-1,
+          legend pos=south west,
+          ymin=0.0,
+          ymax=1.1,
+          ytick={0, 0.2, 0.4, 0.5, 0.6, 0.8, 1},
+          grid=major,
+        ]
+        \addplot[orange, mark=*, solid] % TreeMesh, weak form
+          table [x index=0, y index=1]{../data/pids_euler_weak_cartesian.dat};
+          \addlegendentry{Weak form}
+
+        \addplot[orange, mark=*, dashed] % TreeMesh, flux differencing
+          table [x index=0, y index=1]{../data/pids_euler_ranocha_cartesian.dat};
+          \addlegendentry{Flux differencing}
+      \end{axis}
+    \end{tikzpicture}
+  \end{subfigure}%
+  \caption{Run time per right-hand side evaluation and degree of freedom for
+           different DG discretizations of the 3D compressible Euler equations
+           in FLUXO (Fortran) and \trixi (Julia).}
+  \label{fig:PID-TreeMesh}
+\end{figure}
 

### 2021-08-05 paper/paper.tex
remove the plot of the relative performance of the TreeMesh

--- 2021-08-05 paper/paper.tex
@@
 more efficient than the curvilinear meshes in \trixi.
-
-\todo{TODO: Plot of this comparison with the \lstinline{TreeMesh}?}
 

### 2021-08-05 paper/paper.tex
make some minor tweaks (references, when abbreviations are introduced etc.)

@@
 
+\todo{TODO: Add MHD results}
+
 Note that we do not want to state that Julia is generically faster than Fortran,
@@
 more efficient than the curvilinear meshes in \trixi.
-
-\todo{TODO: Add MHD results?}
 
@@
 testing framework is included in the standard library and default package
-generation setups prepare the file structure accordingly. This allows test
-driven development and continuous integration (CI), which makes code restructuring
+generation setups prepare the file structure accordingly. This allows test-driven
+development and continuous integration (CI), which makes code restructuring
 and optimization straightforward.
@@
 There is still work to be done how to improve the experience in combination
-with MPI on high performance computing (HPC) clusters, but some promising
+with MPI on HPC clusters, but some promising
 approaches exist \cite{byrne2021mpi}.
@@
 implement them in open source software while practitioners or researchers focusing
-on spatial semidiscretizations ca benefit immediately by changing only one line
+on spatial semidiscretizations can benefit immediately by changing only one line
 of code. For example, optimized time integration methods were developed in
@@
 source code, which makes it easy to set up new simulation approaches and analyze
-existing ones, \eg by fluctuation simulations \cite{ranocha2021preventing}.
+existing ones, \eg, by fluctuation simulations \cite{ranocha2021preventing}.
 
@@
 parallelism. The scalability of high-order methods for hyperbolic PDEs written
-in Julia to high performance computing applications still needs to be demonstrated.
+in Julia to high-performance computing applications still needs to be demonstrated.
 However, we do not expect this to remain an unsolvable problem since Julia is not 

### 2021-08-05 data/compute_relative_data.jl
add data/compute_relative_data.jl

@@ -1,0 +1,27 @@
+using DelimitedFiles
+
+"""
+    save_relative_data(filename_in)
+
+Convert the absolute PIDs contained in `filename_in` and save them in another
+file. The input filename `filename_in` should be of the form `some_name.dat`.
+Then, the output filename will be `some_name_relative.dat`.
+"""
+function save_relative_data(filename_in)
+  @assert endswith(filename_in, ".dat")
+  filename_out = replace(filename_in, ".dat" => "_relative.dat")
+
+  open(filename_out, "w") do io
+    data, header = readdlm(filename_in, header=true)
+    writedlm(io, header)
+    reference = data[:, end]
+    for j in 2:size(data, 2)
+        data[:, j] ./= reference
+    end
+    writedlm(io, data)
+  end
+end
+
+save_relative_data("pids_euler_kennedygruber.dat")
+save_relative_data("pids_euler_ranocha.dat")
+save_relative_data("pids_euler_weak.dat")

### 2021-08-05 paper/paper.tex
added gingerbread man example figures and reproducibility instructions

--- 2021-08-05 paper/paper.tex
@@
 methods with adaptive mesh refinement applied to a classical Kelvin-Helmholtz
-flow instability problem.
+flow instability problem. Figure~\ref{fig:pressure_waves} shows the approximation
+of acoustic perturbation equation~\cite{ewert2003acoustic} wave scattering on a
+curvilinear and unstructured domain. The 2D quadrilateral mesh in
+Figure~\ref{fig:pressure_waves}(d) was generated with
+HOHQMesh.jl\xspace\footnote{\url{https://github.com/trixi-framework/HOHQMesh.jl}}.
 
@@
 
-\todo{TODO: Andrew, add gingerbread man example for acoustics} %TODO: Andrew
+\begin{figure}[htbp]
+  \hspace*{\fill}
+  \begin{subfigure}{0.475\linewidth}
+    \includegraphics[width=\textwidth]{../figures/ginger_time_00}
+    \caption{Pressure at time $t = 0$.}
+  \end{subfigure}%
+  \begin{subfigure}{0.475\linewidth}
+    \includegraphics[width=\textwidth]{../figures/ginger_time_08}
+    \caption{Pressure at time $t = 8$.}
+  \end{subfigure}%
+  \\
+  \begin{subfigure}{0.475\linewidth}
+    \includegraphics[width=\textwidth]{../figures/ginger_time_16}
+    \caption{Pressure at time $t = 16$.}
+  \end{subfigure}%
+  \hspace*{\fill}
+  \begin{subfigure}{0.475\linewidth}
+    \includegraphics[width=\textwidth]{../figures/ginger_mesh}
+    \caption{Unstructured quad mesh.}
+  \end{subfigure}%
+  \caption{Numerical solutions of pressure wave scattering for the acoustic 
+                perturbation equations at three times as well as the unstructured, 
+                curvilinear quadrilateral mesh.}
+  \label{fig:pressure_waves}
+\end{figure}
 
@@
   \caption{Run time per right-hand side evaluation and degree of freedom for
-           different DG discretizations of the 3D compressible Euler equations
-           in FLUXO (Fortran) and \trixi (Julia).}
+           different DG discretizations of the 3D compressible Euler equations.}
   \label{fig:PID-Euler}
@@
 more efficient than the curvilinear meshes in \trixi.
+
+\todo{TODO: Plot of this comparison with the \lstinline{TreeMesh}?}
 

### 2021-08-05 figure_gingerbread_man/elixir_ape_gingerbreadman.jl

@@ -1,0 +1,91 @@
+
+###############################################################################
+# generate the gingerbread man mesh file using HOHQMesh
+using HOHQMesh
+
+control_file = joinpath(@__DIR__, "mesh_gingerbread_man.control")
+output = generate_mesh(control_file)
+
+###############################################################################
+# setup the Trixi simulation
+using OrdinaryDiffEq
+using Trixi
+
+###############################################################################
+# semidiscretization of the acoustic perturbation equations
+
+equations = AcousticPerturbationEquations2D(v_mean_global=(0.0, -0.5), c_mean_global=1.0,
+                                            rho_mean_global=1.0)
+
+# Create DG solver with polynomial degree = 6 and (local) Lax-Friedrichs/Rusanov flux
+solver = DGSEM(polydeg=6, surface_flux=flux_lax_friedrichs)
+
+# Create unstructured quadrilateral mesh from a file
+mesh_file = joinpath(@__DIR__, "out/mesh_gingerbread_man.mesh")
+mesh = UnstructuredMesh2D(mesh_file)
+
+# Setup the initial and boundary conditions
+function initial_condition_ginger_pulse(x, t, equations::AcousticPerturbationEquations2D)
+    v1_prime = 0.0
+    v2_prime = 0.0
+    nose          = exp(-log(2) * ((x[1] - 0.0)^2 + (x[2] - 43.5)^2) / 2.0)
+    button_top    = exp(-log(2) * ((x[1] - 0.0)^2 + (x[2] - 30.25)^2) )
+    button_bottom = exp(-log(2) * ((x[1] - 0.0)^2 + (x[2] - 22.5)^2) )
+
+    return SVector(v1_prime, v2_prime, nose + button_top + button_bottom,
+                   Trixi.global_mean_vars(equations)...)
+  end
+
+initial_condition = initial_condition_ginger_pulse
+
+boundary_condition_slip_wall = BoundaryConditionWall(boundary_state_slip_wall)
+boundary_conditions = Dict( :Body    => boundary_condition_slip_wall,
+                            :Button1 => boundary_condition_slip_wall,
+                            :Button2 => boundary_condition_slip_wall,
+                            :Eye1    => boundary_condition_slip_wall,
+                            :Eye2    => boundary_condition_slip_wall,
+                            :Smile   => boundary_condition_slip_wall,
+                            :Bowtie  => boundary_condition_slip_wall )
+
[…]

### 2021-08-05 paper/paper.tex
fixed figure spacing in gingerbread results
better reference to the mesh sub figure

@@
 curvilinear and unstructured domain. The 2D quadrilateral mesh in
-Figure~\ref{fig:pressure_waves}(d) was generated with
+Figure~\ref{fig:ginger_mesh} was generated with
 HOHQMesh.jl\xspace\footnote{\url{https://github.com/trixi-framework/HOHQMesh.jl}}.
@@
 \begin{figure}[htbp]
-  \hspace*{\fill}
   \begin{subfigure}{0.475\linewidth}
@@
   \end{subfigure}%
+  \hspace*{\fill}
   \begin{subfigure}{0.475\linewidth}
@@
     \caption{Unstructured quad mesh.}
+    \label{fig:ginger_mesh}
   \end{subfigure}%

### 2021-08-05 paper/paper.tex
Apply some suggestions from code review
Co-authored-by: Hendrik Ranocha <ranocha@users.noreply.github.com>

@@
 flow instability problem. Figure~\ref{fig:pressure_waves} shows the approximation
-of acoustic perturbation equation~\cite{ewert2003acoustic} wave scattering on a
+of acoustic perturbation equations~\cite{ewert2003acoustic} wave scattering on a
 curvilinear and unstructured domain. The 2D quadrilateral mesh in

### 2021-08-05 figure_gingerbread_man/elixir_ape_gingerbreadman.jl

@@ -21,7 +21,7 @@ equations = AcousticPerturbationEquations2D(v_mean_global=(0.0, -0.5), c_mean_gl
 solver = DGSEM(polydeg=6, surface_flux=flux_lax_friedrichs)
 
 # Create unstructured quadrilateral mesh from a file
-mesh_file = joinpath(@__DIR__, "out/mesh_gingerbread_man.mesh")
+mesh_file = joinpath(@__DIR__, "out", "mesh_gingerbread_man.mesh")
 mesh = UnstructuredMesh2D(mesh_file)
 
 # Setup the initial and boundary conditions

### 2021-08-05 paper/paper.tex
draft of the setup section

@@
 
-FLUXO\footnote{\url{https://gitlab.com/project-fluxo/fluxo}} is an open source
+FLUXO\footnote{\url{https://github.com/project-fluxo/fluxo}} is an open source
 MPI Fortran code implementing discontinuous Galerkin methods on unstructured hex
-meshes in three space dimensions for advection diffusion equations. It also
-implements the same kind of modern DG methods based on flux differencing
+meshes in three space dimensions for advection-diffusion equations. It also
+implements the same kind of modern, flux differencing DG methods 
 to achieve entropy conservation/dissipation or kinetic energy preservation.
-At the same time, both codes also have features the other one does not, \eg,
-parabolic equations and MPI in FLUXO or multiple mesh types and more physics setups
-in \trixi. Nevertheless, both \trixi and FLUXO share a common set of features,
-enabling a reasonable comparison of their performance.
+At the same time, both codes have features the other does not, \eg,
+approximation capabilities for parabolic equations and MPI in FLUXO 
+or multiple mesh types and more physics setups in \trixi. Nevertheless, 
+both \trixi and FLUXO share a common set of features, \eg, high-order DG 
+methods on 3D curvilinear meshes, that enables a reasonable comparison of their performance.
 
@@
 
-\todo{TODO: Andrew, could you please describe the setups that you used? %TODO Andrew
-\begin{itemize}
-  \item Physical setup
-  \item Numerical methods/parameters
-  \item Software versions, compiler \etc
-  \item Hardware
-\end{itemize}}
+For the problem setup we consider a periodic box of the domain $[-1,1]^3$ with four elements in each
+spatial direction. To make the mesh curvilinear, the interior of the box is heavily warped
+by a combination of sine and cosine functions~\cite{chan2019efficient}. To integrate up to the
+final time $t=1.0$, both codes use the five-stage, four-order low-storage explicit
+Runge-Kutta method of Carpenter and Kennedy~\cite{CarpenterKennedy1994}.
+A stable explicit time step is adaptively computed according to the local maximum wave
+speed, the relative grid size, and an adjustable Courant-Friedrichs-Lewy (CFL)
+coefficient $\text{CFL}\in(0,1]$~\cite{gassner2011}. For the performance computations presented herein
+we fix this coefficient to be $\text{CFL}=0.5$.
+
+We compare the performance of FLUXO and \trixi for two smooth nonlinear problems,
+a manufactured solution for the 3D compressible Euler equations and Alfv\'{e}n wave propagation for
+the 3D ideal MHD equations~\cite{altmann2012}. These initial conditions, available in FLUXO and
+\trixi, are typically used for the purpose of demonstrating the high-order accuracy and convergence properties
+of the frameworks.
+
+For these comparisons we examine the performance of FLUXO and \trixi for the weak
+form as well as flux differencing implementations of the DG solver. For the weak form simulations we compute
+the coupling between elements with a Harten-Lax-van Leer (HLL) numerical surface flux function~\cite{Toro2009,li2005hllc}.
+For the flux differencing solver, that introduces an additional numerical volume flux function, we use an entropy-conservative flux
+function in both the volume and at the surface. For the compressible Euler equations this flux is also kinetic energy preserving~\cite{ranocha2020icosahom}
+while for the ideal MHD equations it is kinetic and magnetic energy preserving~\cite{hindenlang2019}. We note that both ideal MHD implementations
+contain additional computational overhead due to nonconservative terms necessary for the numerical approximation 
+to remain entropy consistent~\cite{bohm2020}. Additionally, the weak form DG solver type is recovered if one uses a central flux 
+in the volume of the flux differencing DG solver, albeit with additional computational cost.
+
+All simulations for the performance comparison were run on the High Performance Computing (HPC) system Tetralith 
+provided by the Swedish National Infrastructure for Computing (SNIC). 
+Each node of Tetralith has two sockets, each with an Intel(R) Xeon(R) Gold 6130 @ 2.10GHz with 96GiB of memory. 
+FLUXO was compiled using the Intel compiler suite v18.0.3 and \trixi used Julia v1.6.1. 
+The serial performance results we run on a single core. 
+We ran each configuration of a particular physical setup, polynomial degree, and solver type
+for each code five times, taking the smallest value from each in the comparison results.
+
+To assess the performance of the two codes we vary the polynomial degree (polydeg) in the range $[3,15]$.
+This corresponds to considering DG approximations of (increasing) high-order spatial accuracy, from fourth up to sixteenth order.
+The metric we use to analyze and compare FLUXO and \trixi simulations is a performance index (PID) that
+measures the time required to advance a single degree of freedom (DOF) from one stage of the explicit Runge-Kutta
+time integration scheme to the next. It is computed as
+\begin{equation}
+\label{eqn:pid}
+\text{PID} = \frac{\text{wall-clock-time}}{\#\text{time steps}\cdot\#\text{elements}\cdot(\text{polydeg}+1)^3}.
+\end{equation}
 

### 2021-08-05 paper/paper.tex
updated Euler figures with new PID numbers

@@
 To assess the performance of the two codes we vary the polynomial degree (polydeg) in the range $[3,15]$.
-This corresponds to considering DG approximations of (increasing) high-order spatial accuracy, from fourth up to sixteenth order.
+This corresponds to considering DG approximations of (increasing) spatial accuracy, from fourth up to sixteenth order.
 The metric we use to analyze and compare FLUXO and \trixi simulations is a performance index (PID) that
@@
 \end{equation}
+In other words, the PID acts as a metric for the run time required for the right hand side (RHS) evaluation of each DOF.
 
@@
 
-\begin{figure}
+\begin{figure}[!ht]
 \centering
@@
           ymin=0.0,
-          ymax=2.2e-6,
+          ymax=2.0e-6,
           grid=major,
@@
         \addplot[blue, mark=*, solid] % FLUXO, weak form
-          table [x index=0, y index=3]{../data/pids_euler_weak.dat};
+          table [x index=0, y index=2]{../new_pid_data/pids_euler_weak.dat};
           \addlegendentry{FLUXO}
@@
         \addplot[red, mark=diamond*, solid] % StructuredMesh, weak form
-          table [x index=0, y index=2]{../data/pids_euler_weak.dat};
+          table [x index=0, y index=1]{../new_pid_data/pids_euler_weak.dat};
           % \addlegendentry{\texttt{StructuredMesh} (\trixi)}
@@
         \addplot[blue, mark=*, dashed] % FLUXO, flux differencing
-          table [x index=0, y index=3]{../data/pids_euler_ranocha.dat};
+          table [x index=0, y index=2]{../new_pid_data/pids_euler_ranocha.dat};
 %         \addplot[orange, mark=square*, dashed] % TreeMesh, flux differencing
@@
         \addplot[red, mark=diamond*, dashed] % StructuredMesh, flux differencing
-          table [x index=0, y index=2]{../data/pids_euler_ranocha.dat};
-        \node [rotate=23] at (axis cs:9,8.2e-7) {Flux differencing};
-        \draw (axis cs:9,9.2e-7) -- (axis cs:7.5,10.2e-7);
+          table [x index=0, y index=1]{../new_pid_data/pids_euler_ranocha.dat};
+        \node [rotate=15] at (axis cs:9,8.2e-7) {Flux differencing};
+        \draw (axis cs:9,9.2e-7) -- (axis cs:8.5,10.2e-7);
         \draw (axis cs:9,7.2e-7) -- (axis cs:8.5,5.2e-7);
@@
         \addplot[blue, mark=*, solid, forget plot] % FLUXO, weak form
-          table [x index=0, y index=3]{../data/pids_euler_weak_relative.dat};
+          table [x index=0, y index=2]{../new_pid_data/pids_euler_weak_relative.dat};
 %         \addplot[orange, mark=square*, solid, forget plot] % TreeMesh, weak form
@@
         \addplot[red, mark=diamond*, solid] % StructuredMesh, weak form
-          table [x index=0, y index=2]{../data/pids_euler_weak_relative.dat};
+          table [x index=0, y index=1]{../new_pid_data/pids_euler_weak_relative.dat};
           \addlegendentry{Weak form}
@@
         \addplot[blue, mark=*, dashed, forget plot] % FLUXO, flux differencing
-          table [x index=0, y index=3]{../data/pids_euler_ranocha_relative.dat};
+          table [x index=0, y index=2]{../new_pid_data/pids_euler_ranocha_relative.dat};
 %         \addplot[orange, mark=square*, dashed, forget plot] % TreeMesh, flux differencing
@@
         \addplot[red, mark=diamond*, dashed] % StructuredMesh, flux differencing
-          table [x index=0, y index=2]{../data/pids_euler_ranocha_relative.dat};
+          table [x index=0, y index=1]{../new_pid_data/pids_euler_ranocha_relative.dat};
           \addlegendentry{Flux differencing}
@@
 
-\todo{TODO: Plot of this comparison with the \lstinline{TreeMesh}?}
+\todo{add Intel discussion}
 
-\todo{TODO: Add MHD results?}
+\todo{alter slightly with MHD results}
 

### 2021-08-05 paper/paper.tex
ARW: PID and Gingerbread results (#7)
* added gingerbread man example figures and reproducibility instructions

* fixed figure spacing in gingerbread results

* better reference to the mesh sub figure

* forgot to update the README now that HOHQMesh is used directly using a control file

* ignore output files of HOHQMesh

* Apply some suggestions from code review

Co-authored-by: Hendrik Ranocha <ranocha@users.noreply.github.com>

* adjust file paths in control file for consistency

* added Manifest for the gingerbread man runs

Co-authored-by: Hendrik Ranocha <mail@ranocha.de>
Co-authored-by:

--- 2021-08-05 paper/paper.tex
@@
   \caption{Run time per right-hand side evaluation and degree of freedom for
-           different DG discretizations of the 3D compressible Euler equations.}
+           different DG discretizations of the 3D compressible Euler equations
+           in FLUXO (Fortran) and \trixi (Julia).}
   \label{fig:PID-Euler}
 \end{figure}
+
+\todo{TODO: Add MHD results}
 
@@
 more efficient than the curvilinear meshes in \trixi.
-
-\todo{TODO: Plot of this comparison with the \lstinline{TreeMesh}?}
-
-\todo{TODO: Add MHD results?}
 

### 2021-08-05 figure_gingerbread_man/elixir_ape_gingerbreadman.jl

@@ -1,0 +1,91 @@
+
+###############################################################################
+# generate the gingerbread man mesh file using HOHQMesh
+using HOHQMesh
+
+control_file = joinpath(@__DIR__, "mesh_gingerbread_man.control")
+output = generate_mesh(control_file)
+
+###############################################################################
+# setup the Trixi simulation
+using OrdinaryDiffEq
+using Trixi
+
+###############################################################################
+# semidiscretization of the acoustic perturbation equations
+
+equations = AcousticPerturbationEquations2D(v_mean_global=(0.0, -0.5), c_mean_global=1.0,
+                                            rho_mean_global=1.0)
+
+# Create DG solver with polynomial degree = 6 and (local) Lax-Friedrichs/Rusanov flux
+solver = DGSEM(polydeg=6, surface_flux=flux_lax_friedrichs)
+
+# Create unstructured quadrilateral mesh from a file
+mesh_file = joinpath(@__DIR__, "out", "mesh_gingerbread_man.mesh")
+mesh = UnstructuredMesh2D(mesh_file)
+
+# Setup the initial and boundary conditions
+function initial_condition_ginger_pulse(x, t, equations::AcousticPerturbationEquations2D)
+    v1_prime = 0.0
+    v2_prime = 0.0
+    nose          = exp(-log(2) * ((x[1] - 0.0)^2 + (x[2] - 43.5)^2) / 2.0)
+    button_top    = exp(-log(2) * ((x[1] - 0.0)^2 + (x[2] - 30.25)^2) )
+    button_bottom = exp(-log(2) * ((x[1] - 0.0)^2 + (x[2] - 22.5)^2) )
+
+    return SVector(v1_prime, v2_prime, nose + button_top + button_bottom,
+                   Trixi.global_mean_vars(equations)...)
+  end
+
+initial_condition = initial_condition_ginger_pulse
+
+boundary_condition_slip_wall = BoundaryConditionWall(boundary_state_slip_wall)
+boundary_conditions = Dict( :Body    => boundary_condition_slip_wall,
+                            :Button1 => boundary_condition_slip_wall,
+                            :Button2 => boundary_condition_slip_wall,
+                            :Eye1    => boundary_condition_slip_wall,
+                            :Eye2    => boundary_condition_slip_wall,
+                            :Smile   => boundary_condition_slip_wall,
+                            :Bowtie  => boundary_condition_slip_wall )
+
[…]

### 2021-08-05 paper/paper.tex
partially finished PID section

--- 2021-08-05 paper/paper.tex
@@
 
+\begin{figure}[!ht]
+\centering
+  \begin{subfigure}{\linewidth}
+    \begin{tikzpicture}[
+      font=\footnotesize
+      ]
+      \begin{axis}[
+          xlabel={Polynomial degree},
+          ylabel={Time/RHS/DOF {[sec]}},
+          width=\textwidth,
+          height=0.66\textwidth,
+          legend columns=-1,
+          % legend to name=place_legend_here,
+          legend pos=north west,
+          ymin=0.0,
+          ymax=2.0e-6,
+          grid=major,
+        ]
+        \addplot[blue, mark=*, solid] % FLUXO, weak form
+          table [x index=0, y index=2]{../new_pid_data/pids_mhd_weak.dat};
+          \addlegendentry{FLUXO}
+        \addplot[red, mark=diamond*, solid] % StructuredMesh, weak form
+          table [x index=0, y index=1]{../new_pid_data/pids_mhd_weak.dat};
+          \addlegendentry{\trixi}
+        \node at (axis cs:13,3.5e-7) {Weak form};
+
+        \addplot[blue, mark=*, dashed] % FLUXO, flux differencing
+          table [x index=0, y index=2]{../new_pid_data/pids_mhd_flogor.dat};
+        \addplot[red, mark=diamond*, dashed] % StructuredMesh, flux differencing
+          table [x index=0, y index=1]{../new_pid_data/pids_mhd_flogor.dat};
+        \node [rotate=15] at (axis cs:9,8.2e-7) {Flux differencing};
+        \draw (axis cs:9,9.2e-7) -- (axis cs:8.5,10.2e-7);
+        \draw (axis cs:9,7.2e-7) -- (axis cs:8.5,5.2e-7);
+      \end{axis}
+    \end{tikzpicture}%
+    \caption{Absolute run times.}
+  \end{subfigure}%
+  \\
+  \begin{subfigure}{\linewidth}
+    \begin{tikzpicture}[
+      font=\footnotesize
+      ]
+      \begin{axis}[
+          xlabel={Polynomial degree},
+          ylabel={Time relative to FLUXO {[-]}},
+          width=\textwidth,
+          height=0.66\textwidth,
+          legend columns=-1,
+          legend pos=south west,
+          ymin=0.0,
+          ymax=1.1,
+          ytick={0, 0.2, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1},
+          grid=major,
+        ]
+        \addplot[blue, mark=*, solid, forget plot] % FLUXO, weak form
+          table [x index=0, y index=2]{../new_pid_data/pids_mhd_weak_relative.dat};
+        \addplot[red, mark=diamond*, solid] % StructuredMesh, weak form
+          table [x index=0, y index=1]{../new_pid_data/pids_mhd_weak_relative.dat};
+          \addlegendentry{Weak form}
+
+        \addplot[blue, mark=*, dashed, forget plot] % FLUXO, flux differencing
+          table [x index=0, y index=2]{../new_pid_data/pids_mhd_flogor_relative.dat};
+        \addplot[red, mark=diamond*, dashed] % StructuredMesh, flux differencing
+          table [x index=0, y index=1]{../new_pid_data/pids_mhd_flogor_relative.dat};
+          \addlegendentry{Flux differencing}
+      \end{axis}
+    \end{tikzpicture}
+    \caption{Run time relative to FLUXO.}
+  \end{subfigure}%
+  \caption{Run time per right-hand side evaluation and degree of freedom for
+           different DG discretizations of the 3D ideal MHD equations.}
+  \label{fig:PID-MHD}
+\end{figure}
+
 Note that we do not want to state that Julia is generically faster than Fortran,
@@
 \todo{alter slightly with MHD results}
+
+\todo{Trixi runs need redone due to further improvements}
 

### 2021-08-05 paper/paper.tex

@@
 
-\todo{add Intel discussion}
+\todo{TODO: add Intel discussion} % TODO: Andrew
 
-\todo{alter slightly with MHD results}
+\todo{TODO: alter slightly with MHD results} % TODO: Andrew
 
-\todo{Trixi runs need redone due to further improvements}
+\todo{TODO: \trixi runs need redone due to further improvements} % TODO: Andrew
 
@@
 parallelism. The scalability of high-order methods for hyperbolic PDEs written
-in Julia to high performance computing applications still needs to be demonstrated.
+in Julia to high-performance computing applications still needs to be demonstrated.
 However, we do not expect this to remain an unsolvable problem since Julia is not

### 2021-08-06 paper/paper.tex
Update paper/paper.tex
updated PID figures and added references

@@
 \trixi do not need to compile anything or learn parameter file syntax. Instead, they can
-just load the library and write their simulation setups in Julia. Scripts which describe
+just load the library and write their simulation setups in Julia. Scripts that describe
 a simulation setup for \trixi are referred to as ``elixirs''. As of version
@@
 a manufactured solution for the 3D compressible Euler equations and Alfv\'{e}n wave propagation for
-the 3D ideal MHD equations~\cite{altmann2012}. These initial conditions, available in FLUXO and
+the 3D ideal MHD equations~\cite{gassner2009,altmann2012}. These initial conditions, available in FLUXO and
 \trixi, are typically used for the purpose of demonstrating the high-order accuracy and convergence properties
@@
 the coupling between elements with a Harten-Lax-van Leer (HLL) numerical surface flux function~\cite{Toro2009,li2005hllc}.
-For the flux differencing solver, that introduces an additional numerical volume flux function, we use an entropy-conservative flux
+For the flux differencing solver, which introduces an additional numerical volume flux function, we use an entropy-conservative flux
 function in both the volume and at the surface. For the compressible Euler equations this flux is also kinetic energy preserving~\cite{ranocha2020icosahom}
@@
 to remain entropy consistent~\cite{bohm2020}. Additionally, the weak form DG solver type is recovered if one uses a central flux
-in the volume of the flux differencing DG solver, albeit with additional computational cost.
+in the volume of the flux differencing DG solver, albeit with additional computational cost~\cite{gassner2016split}.
 
@@
 FLUXO was compiled using the Intel compiler suite v18.0.3 and \trixi used Julia v1.6.1.
-The serial performance results we run on a single core.
+The serial performance results were run on a single core.
 We ran each configuration of a particular physical setup, polynomial degree, and solver type
@@
           \addlegendentry{\trixi}
-        \node at (axis cs:13,3.5e-7) {Weak form};
+        \node [rotate=12] at (axis cs:13.5,5.75e-7) {Central flux};
+        \draw (axis cs:13,6.0e-7) -- (axis cs:12.4,7.1e-7);
+        \draw (axis cs:13.4,6.3e-7) -- (axis cs:13.7,1.075e-6);
 
@@
           table [x index=0, y index=1]{../new_pid_data/pids_mhd_flogor.dat};
-        \node [rotate=15] at (axis cs:9,8.2e-7) {Flux differencing};
-        \draw (axis cs:9,9.2e-7) -- (axis cs:8.5,10.2e-7);
-        \draw (axis cs:9,7.2e-7) -- (axis cs:8.5,5.2e-7);
+        \node [rotate=20] at (axis cs:7,1.2e-6) {EC flux};
+        \draw (axis cs:8.4,1.1e-6) -- (axis cs:7.75,1.15e-6);
+        \draw (axis cs:7,1.1e-6) -- (axis cs:7.5,6.5e-7);
       \end{axis}
@@
           ymax=1.1,
-          ytick={0, 0.2, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1},
+          ytick={0, 0.2, 0.5, 0.6, 0.7, 0.8, 0.9, 1},
           grid=major,
@@
           table [x index=0, y index=1]{../new_pid_data/pids_mhd_weak_relative.dat};
-          \addlegendentry{Weak form}
+          \addlegendentry{Central flux}
 
@@
           table [x index=0, y index=1]{../new_pid_data/pids_mhd_flogor_relative.dat};
-          \addlegendentry{Flux differencing}
+          \addlegendentry{EC flux}
       \end{axis}
@@
   \caption{Run time per right-hand side evaluation and degree of freedom for
-           different DG discretizations of the 3D ideal MHD equations.}
+           the flux differencing DG discretization of the 3D ideal MHD equations.
+           Two configurations are compared using either the central volume flux
+           (algebraically equivalent to the weak form DG solver) and the entropy
+           conservative (EC) volume flux.}
   \label{fig:PID-MHD}

### 2021-08-06 paper/paper.tex
updated PID discussion and removed those TODOs

@@
 
-The results of this comparison are visualized in Figure~\ref{fig:PID-Euler}.
-Clearly, \trixi is more than 2x faster than the Fortran code FLUXO
-while both implement the same numerical methods. This demonstrates the
-suitability of Julia for this kind of simulation-focused scientific computing.
+The results of these performance comparisons are visualized for the 3D compressible Euler
+in Figure~\ref{fig:PID-Euler} and for the 3D ideal MHD equations in Figure~\ref{fig:PID-MHD}.
+We present the PID timings for both codes in the top portion of these two Figures as well as
+a relative comparison using FLUXO as reference. Therefore, smaller values in the bottom portion
+the Figures~\ref{fig:PID-Euler} and \ref{fig:PID-MHD} are better.
 
@@
 
+From these results, it is clear that \trixi is more than 2x faster for the compressible Euler equations
+and more than 1.5x faster for the ideal MHD equations than the Fortran code FLUXO.
+We ran the same PID tests in FLUXO compiled with GCC v6.4.0 but found that
+the Intel was between $5\%$ to $23\%$ faster.
+Thus, only results from the Intel compiler are included.
+It is also possible to see the additional computational overhead for the entropy conservative, flux
+differencing DG methods for both physical models. This is because the numerical fluxes are more complex, in
+terms of computational effort, and require special care to optimize their performance~\cite{hendrik_blog}.
+We reiterate that both codes implement the same numerical nodal DG methods on curvilinear meshes used in these tests.
+This demonstrates the suitability of Julia for this kind of simulation-focused scientific computing.
+
 Note that we do not want to state that Julia is generically faster than Fortran,
@@
 more efficient than the curvilinear meshes in \trixi.
-
-\todo{TODO: add Intel discussion} % TODO: Andrew
-
-\todo{TODO: alter slightly with MHD results} % TODO: Andrew
-
-\todo{TODO: \trixi runs need redone due to further improvements} % TODO: Andrew
 

### 2021-08-06 paper/paper.tex
minor revisions

@@
 For the flux differencing solver, which introduces an additional numerical volume flux function, we use an entropy-conservative flux
-function in both the volume and at the surface. For the compressible Euler equations this flux is also kinetic energy preserving~\cite{ranocha2020icosahom}
+function in both the volume and at the surface. For the compressible Euler equations
+this flux is also kinetic energy preserving~\cite{ranocha2020icosahom, ranocha2018thesis}
 while for the ideal MHD equations it is kinetic and magnetic energy preserving~\cite{hindenlang2019}. We note that both ideal MHD implementations
@@
 
-All simulations for the performance comparison were run on the High Performance Computing (HPC) system Tetralith
+All simulations for the performance comparison were run on the HPC system Tetralith
 provided by the Swedish National Infrastructure for Computing (SNIC).
@@
 
-To assess the performance of the two codes we vary the polynomial degree (polydeg) in the range $[3,15]$.
+To assess the performance of the two codes we vary the polynomial degree
+(\lstinline{polydeg}) in the range $[3,15]$.
 This corresponds to considering DG approximations of (increasing) spatial accuracy, from fourth up to sixteenth order.
@@
 \label{eqn:pid}
-\text{PID} = \frac{\text{wall-clock-time}}{\#\text{time steps}\cdot\#\text{elements}\cdot(\text{polydeg}+1)^3}.
+\text{PID} = \frac{\text{wall-clock-time}}{\#\text{time steps}\cdot\#\text{elements}\cdot(\text{\lstinline{polydeg}}+1)^3}.
 \end{equation}
@@
 in Figure~\ref{fig:PID-Euler} and for the 3D ideal MHD equations in Figure~\ref{fig:PID-MHD}.
-We present the PID timings for both codes in the top portion of these two Figures as well as
+We present the PID timings for both codes in the top portion of these two figures as well as
 a relative comparison using FLUXO as reference. Therefore, smaller values in the bottom portion
@@
   \item Prepare repro repo including scripts for \trixi and FLUXO
-  \item Rerun Euler benchmarks and add MHD benchmarks
   \item Gregor: Other aspects of other codes that we like more in \trixi?

### 2021-08-06 paper/paper.tex
fix typo
Co-authored-by: Erik Faulhaber <44124897+efaulhaber@users.noreply.github.com>

@@
 
-# Create a uniformely refined mesh with periodic
+# Create a uniformly refined mesh with periodic
 # boundary conditions in a square domain

### 2021-08-06 paper/paper.tex
fix Trixi.jl version

@@
 
-% TODO: Update Trixi version below
-As of version v0.3.53 (August 2021), \trixi concentrates mainly on discontinuous
+As of version v0.3.55 (August 2021), \trixi concentrates mainly on discontinuous
 Galerkin (DG) methods \cite{hesthaven2007nodal, kopriva2009implementing}.
@@
 a simulation setup for \trixi are referred to as ``elixirs''. As of version
-v0.3.53 (August 2021), \trixi comes with \ca 200 example elixirs. A basic elixir
+v0.3.55 (August 2021), \trixi comes with \ca 200 example elixirs. A basic elixir
 might look as follows:
-% TODO: Update Trixi version above
 

### 2021-08-06 pid_runs/elixir_euler_pid.jl
added elixirs and instructions for PID reproducilibity of Trixi

@@ -1,0 +1,80 @@
+
+using OrdinaryDiffEq
+using Trixi
+
+###############################################################################
+# semidiscretization of the compressible Euler equations
+
+equations = CompressibleEulerEquations3D(1.4)
+
+initial_condition = initial_condition_convergence_test
+source_term=source_terms_convergence_test
+
+
+###############################################################################
+# Get the DG approximation space
+
+# For ECKEP flux split form
+#volume_flux = flux_ranocha
+#solver = DGSEM(polydeg=3, surface_flux=flux_ranocha,
+#               volume_integral=VolumeIntegralFluxDifferencing(volume_flux))
+
+# For standard DG weak form + HLL
+solver = DGSEM(polydeg=3, surface_flux=flux_hll,
+               volume_integral=VolumeIntegralWeakForm())
+
+###############################################################################
+# Get the curved quad mesh from a file
+
+# Mapping as described in https://arxiv.org/abs/2012.12040
+function mapping(xi, eta, zeta)
+  y = eta + 0.15 * (cos(1.5 * pi * xi) * cos(0.5 * pi * eta) * cos(0.5 * pi * zeta))
+
+  x = xi + 0.15 * (cos(0.5 * pi * xi) * cos(2 * pi * y) * cos(0.5 * pi * zeta))
+
+  z = zeta + 0.15 * (cos(0.5 * pi * x) * cos(pi * y) * cos(0.5 * pi * zeta))
+
+  return SVector(x, y, z)
+end
+
+cells_per_dimension = (4, 4, 4)
+mesh = StructuredMesh(cells_per_dimension, mapping)
+
+###############################################################################
+# create the semi discretization object
+
+semi = SemidiscretizationHyperbolic(mesh, equations, initial_condition, solver,
+                                    source_terms=source_term)
+
+###############################################################################
+# ODE solvers, callbacks etc.
+
+tspan = (0.0, 1.0)
+ode = semidiscretize(semi, tspan)
+
+summary_callback = SummaryCallback()
+
+analysis_interval = 100
+analysis_callback = AnalysisCallback(semi, interval=analysis_interval)
+
+alive_callback = AliveCallback(analysis_interval=analysis_interval)
+
+save_solution = SaveSolutionCallback(interval=50,
+                                     save_initial_solution=true,
+                                     save_final_solution=true)
+
+stepsize_callback = StepsizeCallback(cfl=0.5)
+
+callbacks = CallbackSet(summary_callback,
+                        analysis_callback,
[…]

### 2021-08-06 pid_runs/elixir_mhd_pid.jl

@@ -1,0 +1,80 @@
+
+using OrdinaryDiffEq
+using Trixi
+
+###############################################################################
+# semidiscretization of the compressible ideal GLM-MHD equations
+
+equations = IdealGlmMhdEquations3D(5/3)
+
+initial_condition = initial_condition_convergence_test
+
+
+# For ECKEP flux split form
+# volume_flux = (flux_hindenlang_gassner, flux_nonconservative_powell)
+# solver = DGSEM(polydeg=3, surface_flux=(flux_hindenlang_gassner, flux_nonconservative_powell),
+#               volume_integral=VolumeIntegralFluxDifferencing(volume_flux))
+
+# For standard DG weak form + HLL
+volume_flux = (flux_central, flux_nonconservative_powell)
+solver = DGSEM(polydeg=3, surface_flux=(flux_hll, flux_nonconservative_powell),
+               volume_integral=VolumeIntegralFluxDifferencing(volume_flux))
+
+# Create the mesh
+# Note, we use the domain [-1, 1]^3 for the Alfvén wave convergence test case so the
+# warped mapping simplifies (quite a bit)
+
+# Mapping as described in https://arxiv.org/abs/2012.12040
+function mapping(xi, eta, zeta)
+  y = eta + 0.15 * (cos(1.5 * pi * xi) * cos(0.5 * pi * eta) * cos(0.5 * pi * zeta))
+
+  x = xi + 0.15 * (cos(0.5 * pi * xi) * cos(2 * pi * y) * cos(0.5 * pi * zeta))
+
+  z = zeta + 0.15 * (cos(0.5 * pi * x) * cos(pi * y) * cos(0.5 * pi * zeta))
+
+  return SVector(x, y, z)
+end
+
+cells_per_dimension = (4, 4, 4)
+mesh = StructuredMesh(cells_per_dimension, mapping)
+
+# create the semi discretization object
+semi = SemidiscretizationHyperbolic(mesh, equations, initial_condition, solver)
+
+###############################################################################
+# ODE solvers, callbacks etc.
+
+tspan = (0.0, 1.0)
+ode = semidiscretize(semi, tspan)
+
+summary_callback = SummaryCallback()
+
+analysis_interval = 100
+analysis_callback = AnalysisCallback(semi, interval=analysis_interval)
+alive_callback = AliveCallback(analysis_interval=analysis_interval)
+
+save_solution = SaveSolutionCallback(interval=100,
+                                     save_initial_solution=true,
[…]

### 2021-08-06 pid_runs/elixir_euler_pid.jl
remove unneeded headers and unused callbacks from PID reproducibility files

@@ -57,18 +57,10 @@ summary_callback = SummaryCallback()
 analysis_interval = 100
 analysis_callback = AnalysisCallback(semi, interval=analysis_interval)
 
-alive_callback = AliveCallback(analysis_interval=analysis_interval)
-
-save_solution = SaveSolutionCallback(interval=50,
-                                     save_initial_solution=true,
-                                     save_final_solution=true)
-
 stepsize_callback = StepsizeCallback(cfl=0.5)
 
 callbacks = CallbackSet(summary_callback,
                         analysis_callback,
-#                        alive_callback,
-#                        save_solution,
                         stepsize_callback)
 
 ###############################################################################

### 2021-08-06 pid_runs/elixir_mhd_pid.jl

@@ -51,22 +51,13 @@ summary_callback = SummaryCallback()
 
 analysis_interval = 100
 analysis_callback = AnalysisCallback(semi, interval=analysis_interval)
-alive_callback = AliveCallback(analysis_interval=analysis_interval)
-
-save_solution = SaveSolutionCallback(interval=100,
-                                     save_initial_solution=true,
-                                     save_final_solution=true,
-                                     solution_variables=cons2prim)
 
 cfl = 0.5
 stepsize_callback = StepsizeCallback(cfl=cfl)
-
 glm_speed_callback = GlmSpeedCallback(glm_scale=0.5, cfl=cfl)
 
 callbacks = CallbackSet(summary_callback,
                         analysis_callback,
-#                        alive_callback,
-#                        save_solution,
                         stepsize_callback,
                         glm_speed_callback)
 

### 2021-08-06 paper/paper.tex
describe reproducibility material in the main README.md and clean up folders

@@
         \addplot[blue, mark=*, solid] % FLUXO, weak form
-          table [x index=0, y index=2]{../new_pid_data/pids_euler_weak.dat};
+          table [x index=0, y index=2]{../pid_data/pids_euler_weak.dat};
           \addlegendentry{FLUXO}
-%         \addplot[orange, mark=square*, solid] % TreeMesh, weak form
-%           table [x index=0, y index=1]{../data/pids_euler_weak.dat};
-%           \addlegendentry{\texttt{TreeMesh} (\trixi)}
         \addplot[red, mark=diamond*, solid] % StructuredMesh, weak form
-          table [x index=0, y index=1]{../new_pid_data/pids_euler_weak.dat};
-          % \addlegendentry{\texttt{StructuredMesh} (\trixi)}
+          table [x index=0, y index=1]{../pid_data/pids_euler_weak.dat};
           \addlegendentry{\trixi}
@@
         \addplot[blue, mark=*, dashed] % FLUXO, flux differencing
-          table [x index=0, y index=2]{../new_pid_data/pids_euler_ranocha.dat};
-%         \addplot[orange, mark=square*, dashed] % TreeMesh, flux differencing
-%           table [x index=0, y index=1]{../data/pids_euler_ranocha.dat};
+          table [x index=0, y index=2]{../pid_data/pids_euler_ranocha.dat};
         \addplot[red, mark=diamond*, dashed] % StructuredMesh, flux differencing
-          table [x index=0, y index=1]{../new_pid_data/pids_euler_ranocha.dat};
+          table [x index=0, y index=1]{../pid_data/pids_euler_ranocha.dat};
         \node [rotate=15] at (axis cs:9,8.2e-7) {Flux differencing};
@@
         \addplot[blue, mark=*, solid, forget plot] % FLUXO, weak form
-          table [x index=0, y index=2]{../new_pid_data/pids_euler_weak_relative.dat};
-%         \addplot[orange, mark=square*, solid, forget plot] % TreeMesh, weak form
-%           table [x index=0, y index=1]{../data/pids_euler_weak_relative.dat};
+          table [x index=0, y index=2]{../pid_data/pids_euler_weak_relative.dat};
         \addplot[red, mark=diamond*, solid] % StructuredMesh, weak form
-          table [x index=0, y index=1]{../new_pid_data/pids_euler_weak_relative.dat};
+          table [x index=0, y index=1]{../pid_data/pids_euler_weak_relative.dat};
           \addlegendentry{Weak form}
@@
         \addplot[blue, mark=*, dashed, forget plot] % FLUXO, flux differencing
-          table [x index=0, y index=2]{../new_pid_data/pids_euler_ranocha_relative.dat};
-%         \addplot[orange, mark=square*, dashed, forget plot] % TreeMesh, flux differencing
-%           table [x index=0, y index=1]{../data/pids_euler_ranocha_relative.dat};
+          table [x index=0, y index=2]{../pid_data/pids_euler_ranocha_relative.dat};
         \addplot[red, mark=diamond*, dashed] % StructuredMesh, flux differencing
-          table [x index=0, y index=1]{../new_pid_data/pids_euler_ranocha_relative.dat};
+          table [x index=0, y index=1]{../pid_data/pids_euler_ranocha_relative.dat};
           \addlegendentry{Flux differencing}
@@
         \addplot[blue, mark=*, solid] % FLUXO, weak form
-          table [x index=0, y index=2]{../new_pid_data/pids_mhd_weak.dat};
+          table [x index=0, y index=2]{../pid_data/pids_mhd_weak.dat};
           \addlegendentry{FLUXO}
         \addplot[red, mark=diamond*, solid] % StructuredMesh, weak form
-          table [x index=0, y index=1]{../new_pid_data/pids_mhd_weak.dat};
+          table [x index=0, y index=1]{../pid_data/pids_mhd_weak.dat};
           \addlegendentry{\trixi}
@@
         \addplot[blue, mark=*, dashed] % FLUXO, flux differencing
-          table [x index=0, y index=2]{../new_pid_data/pids_mhd_flogor.dat};
+          table [x index=0, y index=2]{../pid_data/pids_mhd_flogor.dat};
         \addplot[red, mark=diamond*, dashed] % StructuredMesh, flux differencing
-          table [x index=0, y index=1]{../new_pid_data/pids_mhd_flogor.dat};
+          table [x index=0, y index=1]{../pid_data/pids_mhd_flogor.dat};
         \node [rotate=20] at (axis cs:7,1.2e-6) {EC flux};
@@
         \addplot[blue, mark=*, solid, forget plot] % FLUXO, weak form
-          table [x index=0, y index=2]{../new_pid_data/pids_mhd_weak_relative.dat};
+          table [x index=0, y index=2]{../pid_data/pids_mhd_weak_relative.dat};
         \addplot[red, mark=diamond*, solid] % StructuredMesh, weak form
-          table [x index=0, y index=1]{../new_pid_data/pids_mhd_weak_relative.dat};
+          table [x index=0, y index=1]{../pid_data/pids_mhd_weak_relative.dat};
           \addlegendentry{Central flux}
@@
         \addplot[blue, mark=*, dashed, forget plot] % FLUXO, flux differencing
-          table [x index=0, y index=2]{../new_pid_data/pids_mhd_flogor_relative.dat};
+          table [x index=0, y index=2]{../pid_data/pids_mhd_flogor_relative.dat};
         \addplot[red, mark=diamond*, dashed] % StructuredMesh, flux differencing
-          table [x index=0, y index=1]{../new_pid_data/pids_mhd_flogor_relative.dat};
+          table [x index=0, y index=1]{../pid_data/pids_mhd_flogor_relative.dat};
           \addlegendentry{EC flux}
@@
 We ran the same PID tests in FLUXO compiled with GCC v6.4.0 but found that
-the Intel was between $5\%$ to $23\%$ faster.
+the Intel was between \SI{5}{\percent} to \SI{23}{\percent} faster.
 Thus, only results from the Intel compiler are included.
@@
 as fast as code written in these traditional scientific computing languages.
-%We believe that it will be possible to get the same efficiency by rewriting
-%FLUXO appropriately.
 Together, the members of our team have strong backgrounds in C, C++, and
@@
 (such as the polynomial degree and the choice of the volume terms), the
-Cartesian \lstinline{TreeMesh} can be \ca \SI{10}{\percent}--\SI{25}{\percent}
+Cartesian \lstinline{TreeMesh} can be \ca \SI{10}{\percent} to \SI{25}{\percent}
 more efficient than the curvilinear meshes in \trixi.

### 2021-08-06 data/compute_relative_data.jl

@@ -1,27 +1,0 @@
-using DelimitedFiles
-
-"""
-    save_relative_data(filename_in)
-
-Convert the absolute PIDs contained in `filename_in` and save them in another
-file. The input filename `filename_in` should be of the form `some_name.dat`.
-Then, the output filename will be `some_name_relative.dat`.
-"""
-function save_relative_data(filename_in)
-  @assert endswith(filename_in, ".dat")
-  filename_out = replace(filename_in, ".dat" => "_relative.dat")
-
-  open(filename_out, "w") do io
-    data, header = readdlm(filename_in, header=true)
-    writedlm(io, header)
-    reference = data[:, end]
-    for j in 2:size(data, 2)
-        data[:, j] ./= reference
-    end
-    writedlm(io, data)
-  end
-end
-
-save_relative_data("pids_euler_kennedygruber.dat")
-save_relative_data("pids_euler_ranocha.dat")
-save_relative_data("pids_euler_weak.dat")

### 2021-08-06 paper/paper.tex
added some changes directly in the text, but also added a TODO comment regarding the examples shown

--- 2021-08-06 paper/paper.tex
@@
 and related problems. The focus of our research ranges from specific applications
-in CFD to general multi-physics coupling strategies to the high-order numerical methods
-such simulations are built upon. In addition, we are academics and involved in teaching
+in CFD to general multi-physics coupling strategies to the development and analysis of high-order numerical methods
+our simulations are built upon. In addition, we are academics and involved in teaching
 students in these areas of science. Thus, we would like to have a
@@
 In addition, suitable initial and boundary conditions (ICs, BCs) are required.
-\trixi also handles non-conservative source terms containing derivatives of the
-conserved quantities as in the shallow water equations or magnetohydrodynamics
+\trixi also handles non-conservative PDE terms as in the shallow water equations or magnetohydrodynamics equations
 with divergence cleaning.
@@
 Some of the main features of \trixi are demonstrated in the following figures.
+
+\todo{I am not super happy about using KHI for a shock capturing demonstration.
+There are no shocks in a KHI - so it is a bit weird.
+I would reformulate this to "demonstrates non-linear stability via positivity and entropy-stability with adaptive mesh refinment....}
+
 Figure~\ref{fig:kelvin_helmholtzs} demonstrates entropy-dissipative shock-capturing
 methods with adaptive mesh refinement applied to a classical Kelvin-Helmholtz
-flow instability problem. Figure~\ref{fig:pressure_waves} shows the approximation
+flow instability problem.
+
+\todo{Maybe we can add a particular shock capturing test case as well, like Sedov with AMR, or the new Mach 2000 jet :-)}
+
+Figure~\ref{fig:pressure_waves} shows the approximation
 of acoustic perturbation equations~\cite{ewert2003acoustic} wave scattering on a
@@
 
-\todo{TODO: Gregor. Be careful how to formulate this, be neutral and avoid making people angry} %TODO: Gregor
+% GG: ok with me now!
 
@@
 \label{eqn:pid}
-\text{PID} = \frac{\text{wall-clock-time}}{\#\text{time steps}\cdot\#\text{elements}\cdot(\text{\lstinline{polydeg}}+1)^3}.
+\text{PID} = \frac{\text{wall-clock-time}}{\#\text{time steps}\cdot 5\cdot\#\text{elements}\cdot(\text{\lstinline{polydeg}}+1)^3},
 \end{equation}
-In other words, the PID acts as a metric for the run time required for the right hand side (RHS) evaluation of each DOF.
+where $5$ is the number of RK stages per time step. In other words, the PID acts as a metric for the run time required for the right hand side (RHS) evaluation of each DOF.
 
@@
 awards DMS-1719818 and DMS-1943186.
-
-\todo{TODO: Gregor, add your funding sources that need to be here} %TODO Gregor
-
+Gregor J.
+Gassner and Michael Schlottke-Lakemper received funding from the European Research Council through the ERC Starting Grant “An Exascale aware and Un-crashable Space-Time-Adaptive Discontinuous Spectral Element Solver for Non-Linear Conservation Laws” (Extreme), ERC grant agreement no.
+714487.
 The performance computations were enabled by resources provided by the Swedish National

### 2021-08-06 paper/paper.tex
Update paper/paper.tex
describe reproducibility material in the main README.md and clean up folders (#11)

--- 2021-08-06 paper/paper.tex
@@
 
-\todo{TODO: Gregor. Be careful how to formulate this, be neutral and avoid making people angry} %TODO: Gregor
-
 \subsection{Description of the setup}

### 2021-08-06 paper/paper.tex
use Trixi.jl v0.2.54 consistently

@@
 
-As of version v0.3.55 (August 2021), \trixi concentrates mainly on discontinuous
+As of version v0.3.54 (August 2021), \trixi concentrates mainly on discontinuous
 Galerkin (DG) methods \cite{hesthaven2007nodal, kopriva2009implementing}.
@@
 a simulation setup for \trixi are referred to as ``elixirs''. As of version
-v0.3.55 (August 2021), \trixi comes with \ca 200 example elixirs. A basic elixir
+v0.3.54 (August 2021), \trixi comes with \ca 200 example elixirs. A basic elixir
 might look as follows:

### 2021-08-06 paper/paper.tex
add jet

@@
 Some of the main features of \trixi are demonstrated in the following figures.
+The detailed physical setups and numerical parameters as well as all code
+necessary to reproduce these figures are available in the reproducibility
+repository for this article \cite{ranocha2021adaptiveRepro}.
 
-\todo{I am not super happy about using KHI for a shock capturing demonstration.
-There are no shocks in a KHI - so it is a bit weird.
-I would reformulate this to "demonstrates non-linear stability via positivity and entropy-stability with adaptive mesh refinment....}
+Figure~\ref{fig:jet} demonstrates demonstrates entropy-dissipative shock-capturing
+methods with positivity-preserving limiters and adaptive mesh refinement applied
+to an astrophysical supersonic jet with Mach number 2000.
+\begin{figure}[!h]
+  \begin{subfigure}{0.53\linewidth}
+    \includegraphics[width=\textwidth]{../figures/jet_density}
+    \caption{Density at time $t = 10^{-3}$.}
+  \end{subfigure}%
+  \hspace*{\fill}
+  \begin{subfigure}{0.47\linewidth}
+    \includegraphics[width=\textwidth]{../figures/jet_mesh}
+    \caption{Mesh at time $t = 10^{-3}$.}
+  \end{subfigure}%
+  \caption{Numerical solutions of a supersonic jet with Mach number 2000 using
+           entropy-dissipative shock-capturing methods with positivity-preserving
+           limiters and adaptive mesh refinement for the compressible Euler equations.}
+  \label{fig:jet}
+\end{figure}
 
-Figure~\ref{fig:kelvin_helmholtzs} demonstrates entropy-dissipative shock-capturing
-methods with adaptive mesh refinement applied to a classical Kelvin-Helmholtz
-flow instability problem.
-
-\todo{Maybe we can add a particular shock capturing test case as well, like Sedov with AMR, or the new Mach 2000 jet :-)}
-
-Figure~\ref{fig:pressure_waves} shows the approximation
-of acoustic perturbation equations~\cite{ewert2003acoustic} wave scattering on a
-curvilinear and unstructured domain. The 2D quadrilateral mesh in
-Figure~\ref{fig:ginger_mesh} was generated with
-HOHQMesh.jl\xspace\footnote{\url{https://github.com/trixi-framework/HOHQMesh.jl}}.
-
-\begin{figure}[htbp]
+Figure~\ref{fig:kelvin_helmholtzs} demonstrates demonstrates nonlinear stability
+obtained with entropy-stable methods and adaptive mesh refinement applied to a
+classical Kelvin-Helmholtz flow instability problem.
+\begin{figure}[!h]
   \begin{subfigure}{0.53\linewidth}
@@
   \caption{Numerical solutions of a Kelvin-Helmholtz instability using
-           entropy-dissipative shock-capturing methods and adaptive mesh
-           refinement for the compressible Euler equations.}
+           entropy-stable methods and adaptive mesh refinement for the
+           compressible Euler equations.}
   \label{fig:kelvin_helmholtzs}
@@
 
-\begin{figure}[htbp]
+Figure~\ref{fig:pressure_waves} shows the approximation
+of acoustic perturbation equations~\cite{ewert2003acoustic} wave scattering on a
+curvilinear and unstructured domain. The 2D quadrilateral mesh in
+Figure~\ref{fig:ginger_mesh} was generated with
+HOHQMesh.jl\xspace\footnote{\url{https://github.com/trixi-framework/HOHQMesh.jl}}.
+\begin{figure}[!h]
   \begin{subfigure}{0.475\linewidth}
@@
 \end{figure}
-
-The detailed physical setups and numerical parameters as well as all code
-necessary to reproduce these figures are available in the reproducibility
-repository for this article \cite{ranocha2021adaptiveRepro}.
 

### 2021-08-06 figure_jet/elixir_euler_astro_jet.jl

@@ -1,0 +1,129 @@
+
+using OrdinaryDiffEq
+using Trixi
+
+###############################################################################
+# semidiscretization of the compressible Euler equations
+gamma = 5/3
+equations = CompressibleEulerEquations2D(gamma)
+
+# Initial condition adopted from
+# - Yong Liu, Jianfang Lu, and Chi-Wang Shu
+#   An oscillation free discontinuous Galerkin method for hyperbolic systems
+#   https://tinyurl.com/c76fjtx4
+# Mach = 2000 jet
+function initial_condition_astro_jet(x, t, equations::CompressibleEulerEquations2D)
+  @unpack gamma = equations
+  rho = 0.5
+  v1 = 0.0
+  v2 = 0.0
+  p =  0.4127
+  # add inflow for t>0 at x=-0.5
+  # domain size is [-0.5,+0.5]^2
+  if (t > 0) && (x[1] ≈ -0.5) && (abs(x[2]) < 0.05)
+    rho = 5.0
+    v1 = 800 # about Mach number Ma = 2000
+    v2 = 0.0
+    p = 0.4127
+  end
+  return prim2cons(SVector(rho, v1, v2, p), equations)
+end
+initial_condition = initial_condition_astro_jet
+
+boundary_conditions = (x_neg=BoundaryConditionDirichlet(initial_condition_astro_jet),
+                       x_pos=BoundaryConditionDirichlet(initial_condition_astro_jet),
+                       y_neg=boundary_condition_periodic,
+                       y_pos=boundary_condition_periodic)
+
+surface_flux = flux_lax_friedrichs # HLLC needs more shock capturing (alpha_max)
+volume_flux  = flux_ranocha # works with Chandrashekar flux as well
+polydeg = 3
+basis = LobattoLegendreBasis(polydeg)
+
+# shock capturing necessary for this tough example
+indicator_sc = IndicatorHennemannGassner(equations, basis,
+                                         alpha_max=0.3,
+                                         alpha_min=0.0001,
+                                         alpha_smooth=true,
+                                         variable=density_pressure)
+volume_integral = VolumeIntegralShockCapturingHG(indicator_sc;
+                                                 volume_flux_dg=volume_flux,
+                                                 volume_flux_fv=surface_flux)
+
+solver = DGSEM(basis, surface_flux, volume_integral)
+
+coordinates_min = (-0.5, -0.5)
+coordinates_max = ( 0.5,  0.5)
+
[…]

### 2021-08-07 paper/paper.tex
cite Liu, Lu, Shu (2021) for the jet problem

@@
 methods with positivity-preserving limiters and adaptive mesh refinement applied
-to an astrophysical supersonic jet with Mach number 2000.
+to an astrophysical supersonic jet with Mach number 2000 \cite{liu2021oscillation}.
 \begin{figure}[!h]

### 2021-08-07 figure_jet/elixir_euler_astro_jet.jl

@@ -10,7 +10,7 @@ equations = CompressibleEulerEquations2D(gamma)
 # Initial condition adopted from
 # - Yong Liu, Jianfang Lu, and Chi-Wang Shu
 #   An oscillation free discontinuous Galerkin method for hyperbolic systems
-#   https://tinyurl.com/c76fjtx4
+#   Preprint available at https://tinyurl.com/c76fjtx4
 # Mach = 2000 jet
 function initial_condition_astro_jet(x, t, equations::CompressibleEulerEquations2D)
   @unpack gamma = equations

### 2021-08-08 paper/paper.tex
Trixi.jl v0.3.55

@@
 
-As of version v0.3.54 (August 2021), \trixi concentrates mainly on discontinuous
+As of version v0.3.55 (August 2021), \trixi concentrates mainly on discontinuous
 Galerkin (DG) methods \cite{hesthaven2007nodal, kopriva2009implementing}.
@@
 a simulation setup for \trixi are referred to as ``elixirs''. As of version
-v0.3.54 (August 2021), \trixi comes with \ca 200 example elixirs. A basic elixir
+v0.3.55 (August 2021), \trixi comes with \ca 200 example elixirs. A basic elixir
 might look as follows:

### 2021-08-09 paper/paper.tex
add template bibtex entries
minor edits
- added "sub-cell"
- fixed error with "demonstrates demonstrates"
- clarified paragraph on Intel compiler

@@
     \item Kinetic energy preserving and entropy-stable methods
-    \item Entropy-stable shock capturing
-    \item Positivity-preserving limiting
+    \item Entropy-stable sub-cell shock capturing
+    \item Sub-cell positivity-preserving limiting
   \end{itemize}
@@
 
-Figure~\ref{fig:jet} demonstrates demonstrates entropy-dissipative shock-capturing
-methods with positivity-preserving limiters and adaptive mesh refinement applied
+Figure~\ref{fig:jet} demonstrates sub-cell entropy-dissipative shock-capturing
+methods with sub-cell positivity-preserving limiters and adaptive mesh refinement applied
 to an astrophysical supersonic jet with Mach number 2000 \cite{liu2021oscillation}.
@@
   \caption{Numerical solutions of a supersonic jet with Mach number 2000 using
-           entropy-dissipative shock-capturing methods with positivity-preserving
+           sub-cell entropy-dissipative shock-capturing methods with sub-cell positivity-preserving
            limiters and adaptive mesh refinement for the compressible Euler equations.}
@@
 \end{equation}
-where $5$ is the number of RK stages per time step. In other words, the PID acts as a metric for the run time required for the right hand side (RHS) evaluation of each DOF.
+where $5$ is the number of RK stages per time step for the selected time-stepping method. In other words, the PID acts 
+as a metric for the run time required for the right hand side (RHS) evaluation of each DOF.
 
@@
 We ran the same PID tests in FLUXO compiled with GCC v6.4.0 but found that
-the Intel was between \SI{5}{\percent} to \SI{23}{\percent} faster.
+the version compiled using Intel was between \SI{5}{\percent} to \SI{23}{\percent} faster.
 Thus, only results from the Intel compiler are included.
@@
 \todo{TODO: Michael, could you please write something here? You are much more
-experienced with MPI and HPC. Be aware of \cite{omlin2020solving} to not step
-on anybodies toes...} %TODO Michael
+experienced with MPI and HPC. Be aware of \cite{omlin2020solving, byrne2021mpi}
+to not step on anybodies toes...} %TODO Michael
 

### 2021-08-10 paper/paper.tex
Remove color boxes around links
Shorten list of goals
Various fixes here and there

@@
 % JuliaCon proceedings template
-\documentclass{juliacon}
+% \documentclass{juliacon}
+% FIXME sloede: I personally hate these color boxes around links, but maybe we should not mess with
+% the official style?
+\documentclass[hidelinks]{juliacon}
 \setcounter{page}{1}
@@
 \begin{enumerate}
-  \item extendable to be useful for research and development
-  \item easy to apply/understand to be useful for students/collaborators
-  \item fast enough to be useful beyond small toy problems
+  \item extendable for research and development,
+  \item easy to apply/understand for students/collaborators,
+  \item fast enough for applied 3D problems.
 \end{enumerate}
@@
 Euler equations of an ideal gas.
-The physical model is specified by the fluxes $f^j$ and the source term $s$.
+The physical system is specified by the fluxes $f^j$ and the source term $s$.
 In addition, suitable initial and boundary conditions (ICs, BCs) are required.
@@
 \cite{tadmor1987numerical, lefloch2002fully, fisher2013discretely,
-ranocha2018comparison, chen2017entropy}. Currently, \trixi comes with the
-following features.
+ranocha2018comparison, chen2017entropy}. Currently, \trixi offers the
+following features:
 \begin{itemize}
@@
 
-Figure~\ref{fig:kelvin_helmholtzs} demonstrates demonstrates nonlinear stability
+Figure~\ref{fig:kelvin_helmholtzs} demonstrates nonlinear stability
 obtained with entropy-stable methods and adaptive mesh refinement applied to a
@@
 of acoustic perturbation equations~\cite{ewert2003acoustic} wave scattering on a
-curvilinear and unstructured domain. The 2D quadrilateral mesh in
-Figure~\ref{fig:ginger_mesh} was generated with
+curvilinear and unstructured domain. The 2D quadrilateral mesh used in the simulation
+(Figure~\ref{fig:ginger_mesh}) was generated with
 HOHQMesh.jl\xspace\footnote{\url{https://github.com/trixi-framework/HOHQMesh.jl}}.
@@
   \caption{Numerical solutions of pressure wave scattering for the acoustic
-                perturbation equations at three times as well as the unstructured,
+                perturbation equations at three points in time as well as the unstructured,
                 curvilinear quadrilateral mesh.}

### 2021-08-10 paper/paper.tex
Replace "ca." by approximately and remove it where already a range is given

@@
 a simulation setup for \trixi are referred to as ``elixirs''. As of version
-v0.3.55 (August 2021), \trixi comes with \ca 200 example elixirs. A basic elixir
+v0.3.55 (August 2021), \trixi comes with approximately 200 example elixirs. A basic elixir
 might look as follows:
@@
 (such as the polynomial degree and the choice of the volume terms), the
-Cartesian \lstinline{TreeMesh} can be \ca \SI{10}{\percent} to \SI{25}{\percent}
+Cartesian \lstinline{TreeMesh} can be \SI{10}{\percent} to \SI{25}{\percent}
 more efficient than the curvilinear meshes in \trixi.

### 2021-08-10 paper/paper.tex
Use four spaces for line continuation indentation
Various small changes

@@
 % \documentclass{juliacon}
-% FIXME sloede: I personally hate these color boxes around links, but maybe we should not mess with
-% the official style?
-\documentclass[hidelinks]{juliacon}
+\documentclass[hidelinks]{juliacon} % To disable ugly color boxes around clickable links
 \setcounter{page}{1}
@@
 % nor formula + references etc.
-We present Trixi.jl, a Julia library for adaptive high-order numerical simulations
+We present Trixi.jl, a Julia package for adaptive high-order numerical simulations
 of hyperbolic partial differential equations (PDEs). Utilizing Julia's strengths,
@@
 integration method. Currently, \trixi focuses on the spatial semidiscretization
-and uses mostly Runge-Kutta methods implemented in OrdinaryDiffEq.jl, a part
+and uses mostly Runge-Kutta methods implemented in OrdinaryDiffEq.jl, which is part
 of DifferentialEquations.jl \cite{rackauckas2017differentialequations}.
@@
 allows us to provide extended functionality for \trixi without modifying any
-main loop. In particular, diverse tasks such as input/output operations, adaptive
+main loop. In particular, various tasks such as input/output operations, adaptive
 mesh refinement (AMR), and positivity preserving limiting are implemented using
@@
 
-We have presented \trixi, a Julia library for adaptive high-order numerical
+We have presented \trixi, a Julia package for adaptive high-order numerical
 simulations of hyperbolic PDEs. As researchers in numerical analysis and

### 2021-08-10 paper/paper.tex
Adapt section titles

@@
 
-\section{Design of \trixi and overview of its features}
+\section{Capabilities and design of \trixi}
 \label{sec:design-of-trixi}
 
-\trixi is designed as a library of high-order methods for hyperbolic PDEs of
-the form
+\trixi is designed as a simulation framework and library of high-order methods for conservation laws
+of the form
 \begin{equation}
@@
 
-\subsection{High-level overview of the structure}
+\subsection{High-level overview of the code structure}
 
@@
 
-\section{Performance comparison with FLUXO}
+\section{Performance comparison with Fortran}
 \label{sec:performance-comparison}

### 2021-08-10 paper/paper.tex
Small change

@@
 Instead of providing only the very basic discretization ingredients, as with
-other open source frameworks for PDEs, \trixi also includes some widely used
-physical systems and analysis routines, such as computation of the integrated kinetic energy
+other open source libraries for PDEs, \trixi also includes some widely used
+physical systems as well as analysis routines, such as computation of the integrated kinetic energy
 for the compressible Euler equations,
 to make it easier to use. Nevertheless, users are not restricted
-to the physics models bundled in \trixi. To setup a new type of \lstinline{equations},
-it is only necessary to create an appropriate \lstinline{struct} bundling all
-parameters and implementing pointwise operations such as the calculation of
-the fluxes $f^j$ in \eqref{eq:hcl} or two-point numerical fluxes. Due to Julia's
-design and just-ahead-of-time compiler, these fluxes can be inlined into the
+to the physics models bundled in \trixi. To set up a new type of \lstinline{equations},
+it is only necessary to create an appropriate \lstinline{struct} containing all
+parameters and to implement pointwise operations such as the calculation of
+the fluxes $f^j$ in \eqref{eq:hcl} or two-point numerical fluxes. Due to the Julia language
+design with just-ahead-of-time compilation, these fluxes can be inlined into the
 library functions of \trixi. Thus, the physics can be separated completely from

### 2021-08-10 paper/paper.tex
Rewrite first paragraph of section 3

@@
 FLUXO\footnote{\url{https://github.com/project-fluxo/fluxo}} is an open source
-MPI Fortran code implementing discontinuous Galerkin methods on unstructured hex
-meshes in three space dimensions for advection-diffusion equations. It also
-implements the same kind of modern, flux differencing DG methods
-to achieve entropy conservation/dissipation or kinetic energy preservation.
+Fortran code implementing discontinuous Galerkin methods on unstructured hexahedral
+meshes in 3D for advection-diffusion equations. It provides
+the same kind of modern, flux differencing DG methods
+to achieve entropy conservation/dissipation or kinetic energy preservation that are used in \trixi.
 At the same time, both codes have features the other does not, \eg,
-approximation capabilities for parabolic equations and MPI in FLUXO
+support for parabolic equations and multi-node parallelism in FLUXO,
 or multiple mesh types and more physics setups in \trixi. Nevertheless,
-both \trixi and FLUXO share a common set of features, \eg, high-order DG
-methods on 3D curvilinear meshes, that enables a reasonable comparison of their performance.
+both \trixi and FLUXO share a common set of features, \ie, high-order DG
+methods on 3D curvilinear meshes, which allows a reasonable comparison of their performance.
 
@@
 Here, we compare the serial performance of \trixi and FLUXO when solving a
-hyperbolic PDE in three space dimensions on curvilinear hex meshes.
+hyperbolic PDE in three space dimensions on curvilinear hexahedral meshes.
 

### 2021-08-10 paper/paper.tex
More changes to section 3

@@
 by a combination of sine and cosine functions~\cite{chan2019efficient}. To integrate up to the
-final time $t=1.0$, both codes use the five-stage, four-order low-storage explicit
+final time $t=1.0$, both codes use the five-stage, fourth-order low-storage explicit
 Runge-Kutta method of Carpenter and Kennedy~\cite{CarpenterKennedy1994}.
@@
 while for the ideal MHD equations it is kinetic and magnetic energy preserving~\cite{hindenlang2019}. We note that both ideal MHD implementations
-contain additional computational overhead due to nonconservative terms necessary for the numerical approximation
+require additional computational effort due to nonconservative terms necessary for the numerical approximation
 to remain entropy consistent~\cite{bohm2020}. Additionally, the weak form DG solver type is recovered if one uses a central flux
@@
 We ran each configuration of a particular physical setup, polynomial degree, and solver type
-for each code five times, taking the smallest value from each in the comparison results.
+for each code five times, taking the smallest value from each for the comparison results.
 
-To assess the performance of the two codes we vary the polynomial degree
+To assess the performance of the two codes, we vary the polynomial degree
 (\lstinline{polydeg}) in the range $[3,15]$.
@@
 \label{eqn:pid}
-\text{PID} = \frac{\text{wall-clock-time}}{\#\text{time steps}\cdot 5\cdot\#\text{elements}\cdot(\text{\lstinline{polydeg}}+1)^3},
+\text{PID} = \frac{\text{wall-clock time}}{\#\text{time steps}\cdot 5\cdot\#\text{elements}\cdot(\text{\lstinline{polydeg}}+1)^3},
 \end{equation}
-where $5$ is the number of RK stages per time step for the selected time-stepping method. In other words, the PID acts 
+where $5$ is the number of Runge-Kutta stages per time step for the selected time-stepping method. In other words, the PID acts
 as a metric for the run time required for the right hand side (RHS) evaluation of each DOF.
@@
 The code and detailed information necessary to reproduce these numerical
-experiments is available in the accompanying repository \cite{ranocha2021adaptiveRepro}.
+experiments are available in the accompanying repository \cite{ranocha2021adaptiveRepro}.
 
@@
 
-The results of these performance comparisons are visualized for the 3D compressible Euler
+The results of these performance measurements are visualized for the 3D compressible Euler equations
 in Figure~\ref{fig:PID-Euler} and for the 3D ideal MHD equations in Figure~\ref{fig:PID-MHD}.
-We present the PID timings for both codes in the top portion of these two figures as well as
-a relative comparison using FLUXO as reference. Therefore, smaller values in the bottom portion
-the Figures~\ref{fig:PID-Euler} and \ref{fig:PID-MHD} are better.
+We present the absolute PID timings for both codes in the top portion of these two figures and 
+a relative comparison using FLUXO as reference in the bottom portion; smaller values thus are always
+better.
 
@@
 Thus, only results from the Intel compiler are included.
-It is also possible to see the additional computational overhead for the entropy conservative, flux
-differencing DG methods for both physical models. This is because the numerical fluxes are more complex, in
-terms of computational effort, and require special care to optimize their performance~\cite{hendrik_blog}.
+It is also possible to see the additional computational effort for the entropy-conservative flux
+differencing DG methods for both physical systems. This is because the numerical fluxes are more
+expensive in
+terms of computational cost, and require special care to optimize their performance~\cite{hendrik_blog}.
 We reiterate that both codes implement the same numerical nodal DG methods on curvilinear meshes used in these tests.

### 2021-08-10 paper/paper.tex
Small changes to Section 4

@@
 
-Julia was designed from scratch to be useful for numerical computing
+Julia was designed from scratch to be suitable for numerical computing
 \cite{bezanson2017julia}. Its usefulness has been demonstrated for example for
@@
 
-As demonstrated in Section~\ref{sec:performance-comparison}, there is no reason
-why Julia should be generically slower than traditional high-performance programming
-languages such as C, C++, and Fortran. A minor exception of this rule is that
-Julia is built on LLVM, which does not perform the same optimizations as most
-optimizing C/C++/Fortran compilers in use nowadays. In particular, LLVM does
-not use fused multiply add (FMA) instructions by default while GCC and Intel
+As demonstrated in Section~\ref{sec:performance-comparison},
+Julia is not generically slower than traditional high-performance programming
+languages such as C, C++, and Fortran. A minor exception is due to
+Julia being built on top of LLVM, which does not perform the same code optimizations as other
+C/C++/Fortran compilers. In particular, LLVM does
+not use fused multiply-add (FMA) instructions by default, while GCC and Intel
 compilers do. Thus, one needs to opt-in to these optimizations explicitly in
-Julia. Luckily, that is relatively easy due to Julia's code manipulation
-techniques which enable the macro \lstinline{@muladd}\footnote{\url{https://github.com/SciML/MuladdMacro.jl}}.
+Julia, e.g., by using the macro
+\lstinline{@muladd}\footnote{\url{https://github.com/SciML/MuladdMacro.jl}}.
 
@@
 
-Many scientists implementing numerical methods often have no real experience or
-education in software development. Julia and its ecosystem support these
+Many scientists implementing numerical methods often have no formal training
+in software development. Julia and its ecosystem support these
 researchers by making it easy to set up unit and regression tests, since a
@@
 
-Julia itself and most packages are developed on GitHub. The community provides
-many tools one can use to run corresponding CI setups, prepare and publish
+Julia itself and most packages are developed on GitHub. The Julia community provides
+several tools to run corresponding CI setups, prepare and publish
 documentation\footnote{\url{https://github.com/JuliaDocs/Documenter.jl}},
 and other related tasks. This setup also encourages other good software development
-practices that we use in \trixi, such as mandatory code review before pull requests
-are merged.
+practices that we use in \trixi, such as mandatory code reviews.
 
-In addition, the package manager encourages a unified form of semantic versioning
+In addition, the package manager fosters a unified form of semantic versioning
 across the Julia ecosystem, which lays the basis for the following three
@@
 
-The package manager makes it easy to reproduce the exact run environment, including
+The package manager makes it easy to reproduce the exact runtime environment, including
 binary dependencies, used to generate numerical results. We have used this feature for all of our papers based on
@@
 To incorporate some of the adaptive mesh capabilities into \trixi
-we have written a Julia wrapper\footnote{\url{https://github.com/trixi-framework/P4est.jl}}
-of the C library \texttt{p4est} \cite{burstedde2011p4est}. In our experience,
-it is relatively easy to do so in Julia and the package manager in combination
+we have created a Julia wrapper\footnote{\url{https://github.com/trixi-framework/P4est.jl}}
+for the C library \texttt{p4est} \cite{burstedde2011p4est}. In our experience,
+it is relatively easy to do so in Julia, and the package manager in combination
 with BinaryBuilder.jl makes it convenient to distribute the necessary binaries.
@@
 Due to Julia's package manager, the cost of using external dependencies is low
-enough to not be a problem. Traditionally, many scientific simulation codes tend
+enough to not become an issue. Traditionally, many scientific simulation codes tend
 to reduce the number of external dependencies as much as possible due to the
 complexity of handling different versions and making everything work together.
-Such an approach typically leads to significant code duplication. For example, many
+Such an approach often leads to significant code duplication. For example, many
 CFD codes implement their own time integration methods, arguing that the spatial
 part is much more complex. While this is often true, it is still more efficient
-in terms of human workload to be able to reuse existing implementations. This
+in terms of developer time to reuse existing implementations. This
 allows experts on time integration methods to develop specialized algorithms and
-implement them in open source software while practitioners or researchers focusing
-on spatial semidiscretizations can benefit immediately by changing only one line
-of code. For example, optimized time integration methods were developed in
+implement them in open source software, while practitioners or researchers focusing
+on spatial semidiscretizations can benefit with near-zero effort.
+In our case, optimized time integration methods were developed in
 \cite{ranocha2021optimized} and implemented in OrdinaryDiffEq.jl. These methods
 can be used with \trixi by changing a single line of code. In contrast, researchers
-working without external dependencies would need to spend a day or two to digest
-details before even beginning implementation. This additional overhead in developer
-time makes it less likely that researchers would implement these methods in their own
+working without external dependencies would need extra time to digest a new method's
+details before even beginning an implementation. This additional overhead
+makes it less likely that researchers would use these novel schemes in their own
 codes and benefit from recent algorithmic developments.
 
-An additional example is given by automatic differentiation. Due to dispatch,
-it is conveniently possible to use forward mode automatic differentiation
+An additional example is given by automatic differentiation. Due to dynamic multiple dispatch,
+it is possible to conveniently use forward mode automatic differentiation
 \cite{revels2016forward} to compute Jacobians of semidiscretizations of nonlinear
-conservation laws with \trixi or to differentiate through a complete simulation
+conservation laws with \trixi or to differentiate through a complete simulation,
 including time integration methods from OrdinaryDiffEq.jl.
@@
 Such a difference between the programming languages used in the front- and
-back-ends lead to a well-known barrier between a user and the developers of a library.
+backend leads to a well-known barrier between a user and the developers of a library.
 Julia was designed from the beginning to make it possible to write both simple
@@
 two-language problem --- at least to a significant extent. Highly efficient low-level
-code will of course still look different from simple high-level code, but the
+code will still look different from simple high-level code, but the
 barrier between them is much smaller. In particular, this makes it easier for
@@
 code. In addition, Julia's expressiveness and the ability to work with a restricted
-subset of all possible features have enabled more than 18 students to use \trixi
-for their coursework or theses. In our experience, the time required to get
-started with traditional monolithic code bases is so long that students either
-choose to build their own implementations specialized on their tasks or can only
-work on toy problems since the remaining time is rather restricted.
+subset of all possible features have enabled already more than 18 students to use \trixi
+for their coursework or theses. In our experience, the effort to get
+started with traditional monolithic code bases is often so large that due to time restrictions
+students either choose to build their own implementations specialized on their tasks or only work on
+projects with a limited scope.
 
-Another anecdotal example of ease of use is the preprint \cite{singh2021linear}.
+Another anecdotal example for the ease of use is the preprint \cite{singh2021linear}.
 While the authors utilized \trixi for some numerical experiments, they did so
 independently of the development team (who learned that the authors were
-using \trixi only \textit{after} reading the preprint!).
+using \trixi only \textit{after} reading the preprint).
 

### 2021-08-10 paper/paper.tex
Small change to section 5

@@
 
-Learning a new programming language requires of course some effort and needs some
-time. Nevertheless, it has paid off in our experience and we do not want to miss the
-nice features of \trixi enabled by Julia and its ecosystem. Thus, while there is
-of course no need to switch to Julia if people are satisfied with their existing
+Learning a new programming language naturally requires some time and effort.
+Nevertheless, it has paid off in our experience and we do not want to miss the
+many useful features of \trixi enabled by Julia and its ecosystem. Thus, while there is
+no need to switch to Julia if people are satisfied with their existing
 tools, we encourage researchers to try out Julia and stay for a while. The Julia

### 2021-08-10 paper/paper.tex
Some fixes due to comments

@@
 languages such as C, C++, and Fortran. A minor exception is due to
-Julia being built on top of LLVM, which does not perform the same code optimizations as other
+Julia being built on top of LLVM, which does not perform the same optimizations as other
 C/C++/Fortran compilers. In particular, LLVM does
 not use fused multiply-add (FMA) instructions by default, while GCC and Intel
-compilers do. Thus, one needs to opt-in to these optimizations explicitly in
-Julia, e.g., by using the macro
+compilers do, and thus one needs to opt-in to these optimizations explicitly.
+Relying on Julia's code manipulation techniques, this can be achieved by using the macro
 \lstinline{@muladd}\footnote{\url{https://github.com/SciML/MuladdMacro.jl}}.
@@
 
-An additional example is given by automatic differentiation. Due to dynamic multiple dispatch,
+An additional example is given by automatic differentiation. Due to multiple dispatch,
 it is possible to conveniently use forward mode automatic differentiation

### 2021-08-10 paper/paper.tex
Use GitLab.com as the repo URL for FLUXO

@@
 
-FLUXO\footnote{\url{https://github.com/project-fluxo/fluxo}} is an open source
+FLUXO\footnote{\url{https://gitlab.com/project-fluxo/fluxo}} is an open source
 Fortran code implementing discontinuous Galerkin methods on unstructured hexahedral

### 2021-08-10 paper/paper.tex
Use "acoustic pressure" instead of "pressure" to be clearer
Small fixes (#16)
* Use GitLab.com as the repo URL for FLUXO

* Use "acoustic pressure" instead of "pressure" to be clearer

* Shorten title

* Change "extendable" -> "extensible"

@@
 of hyperbolic partial differential equations (PDEs). Utilizing Julia's strengths,
-Trixi.jl is extendable, easy to use, and fast. We describe the main design choices
+Trixi.jl is extensible, easy to use, and fast. We describe the main design choices
 that enable these features and compare Trixi.jl with an established open
@@
 \begin{enumerate}
-  \item extendable for research and development,
+  \item extensible for research and development,
   \item easy to apply/understand for students/collaborators,
@@
 In this article, we present an overview of the main features and design decisions
-of \trixi in Section~\ref{sec:design-of-trixi}, laying the ground for an extendable
+of \trixi in Section~\ref{sec:design-of-trixi}, laying the ground for an extensible
 and easy-to-use framework of high-order methods for hyperbolic PDEs. Next, we
@@
     \includegraphics[width=\textwidth]{../figures/ginger_time_00}
-    \caption{Pressure at time $t = 0$.}
+    \caption{Acoustic pressure at time $t = 0$.}
   \end{subfigure}%
@@
     \includegraphics[width=\textwidth]{../figures/ginger_time_08}
-    \caption{Pressure at time $t = 8$.}
+    \caption{Acoustic pressure at time $t = 8$.}
   \end{subfigure}%
@@
     \includegraphics[width=\textwidth]{../figures/ginger_time_16}
-    \caption{Pressure at time $t = 16$.}
+    \caption{Acoustic pressure at time $t = 16$.}
   \end{subfigure}%
@@
 simulations of hyperbolic PDEs. As researchers in numerical analysis and
-scientific computing, our goals were to create a framework that is extendable,
+scientific computing, our goals were to create a framework that is extensible,
 easy to use, and fast (in this particular order). Making use of Julia's strengths,

### 2021-08-11 paper/paper.tex
remove trailing / in links and obsolete TODO notes

@@
 \end{enumerate}
-Additionally, we are greedy\footnote{See \url{https://julialang.org/blog/2012/02/why-we-created-julia/}}
+Additionally, we are greedy\footnote{See \url{https://julialang.org/blog/2012/02/why-we-created-julia} (accessed 2021-08-11)}
 and wish to include all these features within a single code.
@@
 big data\footnote{Celeste project: \url{https://github.com/jeff-regier/Celeste.jl},
-see \url{https://juliacomputing.com/case-studies/celeste/} (accessed 2021-07-22)},
+see \url{https://juliacomputing.com/case-studies/celeste} (accessed 2021-08-11)},
 machine learning \cite{innes2018fashionable}, and scientific machine learning
@@
 tools, we encourage researchers to try out Julia and stay for a while. The Julia
-community (on the Julia Discourse forum\footnote{\url{https://discourse.julialang.org/}}
+community (on the Julia Discourse forum\footnote{\url{https://discourse.julialang.org}}
 or Slack/Zulip workspaces) is usually welcoming and helpful, both for new and
@@
 
-\todo{ % TODO
-\begin{itemize}
-  \item Full paper: 5--10 pages
-  \item Prepare repro repo including scripts for \trixi and FLUXO
-  \item Gregor: Other aspects of other codes that we like more in \trixi?
-  \item Are there other aspects we should discuss?
-  \item Check affiliations
-\end{itemize}
-\begin{quote}
-  Compared to an extended abstract, a full paper presents more of the background
-  and context motivating the work. It compares the work to other approaches taken
-  in the field and gives some additional insights on the conference contribution.
-  Use cases back up the work by showing how it can be used.
-\end{quote}}
-
 \end{document}

### 2021-08-12 paper/paper.tex
Write section 4.2.2

@@
 \subsubsection{External libraries can be integrated relatively easily}
+\label{sec:external-libraries}
 
@@
 
-\subsubsection{MPI-based parallel simulations need to be established}
+\subsubsection{Simulations with distributed-memory parallelism}
 
-MPI: To be evaluated but the basic tools are already there \cite{byrne2021mpi}
+While Julia offers its own approach to distributed computing via the \lstinline{Distributed} package
+in the standard library, its performance and scalability are currently limited \cite{byrne2021mpi}.
+As an alternatively, the package MPI.jl utilizes the Message Passing Interface (MPI) for data
+exchange \cite{byrne2021mpi}. It provides an API similar to MPI's C interface and allows to either
+wrap MPI binaries installed through Julia's package manager or to make use of an existing MPI
+installation.  Most Julia projects focusing on distributed-memory simulations seem to be using
+MPI.jl, such as ClimateMachine.jl\footnote{\url{https://github.com/CliMA/ClimateMachine.jl}} or
+ImplicitGlobalGrid.jl\footnote{\url{https://github.com/eth-cscs/ImplicitGlobalGrid.jl}}.  In
+general, the performance of MPI programs written in Julia is to be comparable to C programs
+\cite{hunold2020benchmarking}, and parallel simulations with Julia have been shown to scale
+to more than 5,000 GPUs with MPI.jl \cite{omlin2020solving}.
 
-\todo{TODO: Michael, could you please write something here? You are much more
-experienced with MPI and HPC. Be aware of \cite{omlin2020solving, byrne2021mpi}
-to not step on anybodies toes...} %TODO Michael
+However, there are also some practical challenges associated with running MPI-based parallel
+simulations with Julia. For starters, the issue of compilation times as described in the previous
+section is exacerbated, as MPI runs are by default not interactive and thus compilation results cannot be cached between MPI sessions. Since this
+requires to recompile everything for each subsequent execution of an MPI-parallel Julia program,
+this is particularly annoying during development. Partial relief can be obtained by using
+\texttt{tmpi}\footnote{\url{https://github.com/Azrael3000/tmpi}}, which allows to simultaneously
+interact with multiple interactive Julia REPL sessions via the terminal multiplexer \texttt{tmux}.
+However, this approach does not scale beyond a few MPI ranks and is only available on
+Unix-like operating systems. The use of PackageCompiler.jl to reduce compilation times as much as
+possible becomes thus mandatory. Another issue encountered when running Julia with many parallel
+processes is that Julia operates with a lot of small files during startup and precompilation,
+putting heavy pressure on the parallel file system. This can be partially alleviated by doing
+precompilation serially, and there are tools like
+\texttt{Spindle}\footnote{\url{https://github.com/hpc/Spindle}}, which help to solve similar
+problems that occur when running Python programs in parallel \cite{frings2013parallelloading}.
+Finally, the ease of using external libraries as desribed in Section~\ref{sec:external-libraries} is
+somewhat lost when working with dependencies that themselves are parallelized with MPI. In this
+case it is necessary to manually specify the paths to locally compiled libraries that have been
+built against the respective system MPI installation, since there is generally no binary
+compatibility between different MPI implementations. There are ongoing efforts in the Julia HPC
+community to work around this limitation by using the
+\texttt{WI4MPI}\footnote{\url{https://github.com/cea-hpc/wi4mpi}} wrapper package.
+
+Even though the aforementioned difficulties do not prevent massively parallel simulations with
+Julia, they raise the entry barrier for new users and developers. The overarching issue, however, is
+that at the time of writing, there is little precedence in terms of openly available tutorials,
+citable publications, or highly scalable example applications that can be used to learn about best
+practices for MPI-based parallism in Julia. Thus while in general it is justified to be cautiously
+optimistic that most issues can be overcome, it is too early to make a final assessment on the
+suitability of Julia for highly parallel simulations.
 

### 2021-08-12 paper/paper.tex
Add sentence on own time integration schemes
Update Trixi version

@@
 mesh refinement (AMR), and positivity preserving limiting are implemented using
-callbacks.
+callbacks. In case full control over the time step loop is required, e.g., to experiment with new
+features that are not easily realized with the existing API, \trixi also implements its own time
+integration schemes that mimic the interface of OrdinaryDiffEq.jl and that can be used as drop-in
+replacements.
 
@@
 a simulation setup for \trixi are referred to as ``elixirs''. As of version
-v0.3.55 (August 2021), \trixi comes with approximately 200 example elixirs. A basic elixir
-might look as follows:
+v0.3.57 (August 2021), \trixi comes with approximately 200 example elixirs. A basic elixir
+looks as follows:
 

### 2021-08-12 paper/paper.tex
Add sentence on simplicity over abstractions
Slightly rewrite performance section

@@
 difference methods, and variants of finite volume methods.
+And even if a new feature requires modifications to methods provided by \trixi: Due to favoring
+simplicity over abstractions in the code, it remains comparatively straightforward to modify
+existing implementations.
 
@@
 
-Note that we do not want to state that Julia is generically faster than Fortran,
-C, or C++. In contrast, we would like to emphasize that Julia can be at least
-as fast as code written in these traditional scientific computing languages.
-Together, the members of our team have strong backgrounds in C, C++, and
-Fortran. Nevertheless, our Julia code is faster than the established HPC Fortran
-code FLUXO for this example. The reason is not that Fortran or Julia are generically
-faster; in contrast, we expect to be able to achieve similar performance from either
-language. Instead, \trixi owes its performance optimizations in part to the code
-introspection and profiling tools available in Julia. Similar tools used to optimize the
-performance in other languages are usually not as easy to use as their Julia
-equivalents (if they exist at all).
+While our Julia code is faster than the established HPC Fortran code FLUXO for this non-trivial
+example, we do not claim that Julia is generally faster than Fortran, C, or C++.
+Instead, we would like to emphasize that Julia code can be \emph{at least
+as fast} as code written in these traditional scientific computing languages.
+Further, we expect to be able to achieve similar performance with either language, and that
+remaining performance differences are due to how much time and effort has been spent to optimize
+each code.  \trixi owes its performance optimizations in part to the code introspection and
+profiling tools available in Julia. Similar tools used to optimize the performance in other
+languages are often not as easy to use as their Julia counterparts.
 

### 2021-08-12 paper/paper.tex
Somewhat simplify statement

@@
 researchers by making it easy to set up unit and regression tests, since a
-testing framework is included in the standard library and default package
-generation setups prepare the file structure accordingly. This allows test-driven
+testing framework is included in the standard library. This facilitates test-driven
 development and continuous integration (CI), which makes code restructuring

### 2021-08-12 paper/paper.tex
Slightly rephrase conclusions to match new section on MPI parallelism

@@
 in Julia to high-performance computing applications still needs to be demonstrated.
-However, we do not expect this to remain an unsolvable problem since Julia is not
+Yet, we do not consider this an unsolvable problem, since Julia is not
 generically slower than traditional compiled HPC programming languages. Nevertheless,
-it appears to be more complicated to scale Julia to HPC while still keeping the
-code as simple, clean, and flexible as it is now.
+it appears to be more complicated to scale Julia to distributed systems without losing some of the
+simplicity and flexibility it offers for serial or shared-memory parallel computations.
 
 Learning a new programming language naturally requires some time and effort.
-Nevertheless, it has paid off in our experience and we do not want to miss the
-many useful features of \trixi enabled by Julia and its ecosystem. Thus, while there is
+In our experience, however, it has paid off and we do not want to miss the
+many useful features of \trixi enabled by Julia and its ecosystem. While there is
 no need to switch to Julia if people are satisfied with their existing
-tools, we encourage researchers to try out Julia and stay for a while. The Julia
+tools, we encourage researchers to try out Julia for scientific computing and to stay for a while.
+The Julia
 community (on the Julia Discourse forum\footnote{\url{https://discourse.julialang.org}}

### 2021-08-12 paper/paper.tex
Remove last statement from conclusions since it is not substantiated in the main part of the paper
Apply suggestions from code review
Thanks for the suggestions, @ranocha!

Co-authored-by: Hendrik Ranocha <ranocha@users.noreply.github.com>

@@
 a simulation setup for \trixi are referred to as ``elixirs''. As of version
-v0.3.57 (August 2021), \trixi comes with approximately 200 example elixirs. A basic elixir
+v0.3.55 (August 2021), \trixi comes with approximately 200 example elixirs. A basic elixir
 looks as follows:
@@
 as fast} as code written in these traditional scientific computing languages.
-Further, we expect to be able to achieve similar performance with either language, and that
-remaining performance differences are due to how much time and effort has been spent to optimize
-each code.  \trixi owes its performance optimizations in part to the code introspection and
+Further, we expect to be able to achieve similar performance with either language
+by spending enough time and effort to optimize the respective code.
+\trixi owes its performance optimizations in part to the code introspection and
 profiling tools available in Julia. Similar tools used to optimize the performance in other
@@
 ImplicitGlobalGrid.jl\footnote{\url{https://github.com/eth-cscs/ImplicitGlobalGrid.jl}}.  In
-general, the performance of MPI programs written in Julia is to be comparable to C programs
+general, the performance of MPI programs written in Julia is comparable to C programs
 \cite{hunold2020benchmarking}, and parallel simulations with Julia have been shown to scale
-to more than 5,000 GPUs with MPI.jl \cite{omlin2020solving}.
+to more than 5{,}000 GPUs with MPI.jl \cite{omlin2020solving}.
 
@@
 citable publications, or highly scalable example applications that can be used to learn about best
-practices for MPI-based parallism in Julia. Thus while in general it is justified to be cautiously
+practices for MPI-based parallism in Julia. Thus, while in general it is justified to be cautiously
 optimistic that most issues can be overcome, it is too early to make a final assessment on the
@@
 or Slack/Zulip workspaces) is usually welcoming and helpful, both for new and
-experienced users. Making use of existing learning material and asking questions
-appears to be important since a new programming style is necessary to make full
-use of all of Julia's strengths.
+experienced users.
 

### 2021-08-12 paper/paper.tex
Update paper/paper.tex
Co-authored-by: Michael Schlottke-Lakemper <michael@sloede.com>

@@
 And even if a new feature requires modifications to methods provided by \trixi: Due to favoring
-simplicity over abstractions in the code, it remains comparatively straightforward to modify
+simplicity over excessively generic code, it remains comparatively straightforward to modify
 existing implementations.

### 2021-08-12 paper/paper.tex
fix typos

@@
 problems that occur when running Python programs in parallel \cite{frings2013parallelloading}.
-Finally, the ease of using external libraries as desribed in Section~\ref{sec:external-libraries} is
+Finally, the ease of using external libraries as described in Section~\ref{sec:external-libraries} is
 somewhat lost when working with dependencies that themselves are parallelized with MPI. In this
@@
 citable publications, or highly scalable example applications that can be used to learn about best
-practices for MPI-based parallism in Julia. Thus, while in general it is justified to be cautiously
-optimistic that most issues can be overcome, it is too early to make a final assessment on the
+practices for MPI-based parallelism in Julia. Thus, while in general it is justified to be cautiously
+optimistic that most issues can be overcome, it is too early to make a final assessment of the
 suitability of Julia for highly parallel simulations.

### 2021-08-12 paper/paper.tex
fix Fisher et al. reference
Consistency fix

@@
 In particular, it has a focus on entropy-conservative and -dissipative methods
-\cite{tadmor1987numerical, lefloch2002fully, fisher2013discretely,
+\cite{tadmor1987numerical, lefloch2002fully, fisher2013high,
 ranocha2018comparison, chen2017entropy}. Currently, \trixi offers the
@@
 while for the ideal MHD equations it is kinetic and magnetic energy preserving~\cite{hindenlang2019}. We note that both ideal MHD implementations
-require additional computational effort due to nonconservative terms necessary for the numerical approximation
+require additional computational effort due to non-conservative terms necessary for the numerical approximation
 to remain entropy consistent~\cite{bohm2020}. Additionally, the weak form DG solver type is recovered if one uses a central flux

### 2021-08-12 paper/paper.tex
frontend/backend -> front end/back end (seems to be the most used form)

@@
 
-Such a difference between the programming languages used in the front- and
-backend leads to a well-known barrier between a user and the developers of a library.
+Such a difference between the programming languages used in the front end and
+back end leads to a well-known barrier between a user and the developers of a library.
 Julia was designed from the beginning to make it possible to write both simple

### 2021-08-13 paper/paper.tex
Add "the" for improved reading flow

@@
 Trixi.jl is extensible, easy to use, and fast. We describe the main design choices
-that enable these features and compare Trixi.jl with an established open
+that enable these features and compare Trixi.jl with a mature open
 source Fortran code for hyperbolic PDEs using the same numerical methods.
@@
 and related problems. The focus of our research ranges from specific applications
-in CFD to general multi-physics coupling strategies to the development and analysis of high-order numerical methods
+in CFD to general multi-physics coupling strategies to the development and analysis of the high-order numerical methods
 our simulations are built upon. In addition, we are academics and involved in teaching
@@
 and easy-to-use framework of high-order methods for hyperbolic PDEs. Next, we
-compare the serial performance with an established high-performance computing (HPC) Fortran code in
+compare the serial performance with a mature high-performance computing (HPC) Fortran code in
 Section~\ref{sec:performance-comparison}, demonstrating that Julia is not generically
@@
 faster at the cost of simplicity and readability. Nevertheless, \trixi is already quite
-fast, even compared to established high-performance open source software,
+fast, even compared to mature high-performance open source software,
 as described in detail in Section~\ref{sec:performance-comparison}. In this
@@
 
-While our Julia code is faster than the established HPC Fortran code FLUXO for this non-trivial
+While our Julia code is faster than the mature HPC Fortran code FLUXO for this non-trivial
 example, we do not claim that Julia is generally faster than Fortran, C, or C++.
@@
 the number of students and researchers working with \trixi, and serial performance
-comparisons with an established Fortran code. Having developed, from scratch, \trixi for a bit
+comparisons with a mature Fortran code. Having developed, from scratch, \trixi for a bit
 more than one year allows us to give an assessment of Julia for simulation-focused

### 2021-08-13 paper/paper.tex
Improve word flow
Shorten introduction

@@
 We present Trixi.jl, a Julia package for adaptive high-order numerical simulations
-of hyperbolic partial differential equations (PDEs). Utilizing Julia's strengths,
+of hyperbolic partial differential equations. Utilizing Julia's strengths,
 Trixi.jl is extensible, easy to use, and fast. We describe the main design choices
 that enable these features and compare Trixi.jl with a mature open
-source Fortran code for hyperbolic PDEs using the same numerical methods.
+source Fortran code that uses the same numerical methods.
 We conclude with an assessment of Julia for simulation-focused scientific
@@
   \item extensible for research and development,
-  \item easy to apply/understand for students/collaborators,
+  \item easy to understand and use for students and collaborators,
   \item fast enough for applied 3D problems.

### 2021-08-13 paper/paper.tex
Remove unused abbreviations
Improve reading flow

@@
 The physical system is specified by the fluxes $f^j$ and the source term $s$.
-In addition, suitable initial and boundary conditions (ICs, BCs) are required.
-\trixi also handles non-conservative PDE terms as in the shallow water equations or magnetohydrodynamics equations
+In addition, suitable initial and boundary conditions are required.
+\trixi also handles non-conservative PDE terms such as the shallow water equations or the magnetohydrodynamics equations
 with divergence cleaning.

### 2021-08-13 paper/paper.tex
Remove parens since sentence makes no sense without them
established -> mature

--- 2021-08-13 paper/paper.tex
@@
 and related problems. The focus of our research ranges from specific applications
-in CFD to general multi-physics coupling strategies to the development and analysis of the high-order numerical methods
+in CFD to general multi-physics coupling strategies to the development and analysis of high-order numerical methods
 our simulations are built upon. In addition, we are academics and involved in teaching

### 2021-08-13 paper/paper.tex
library -> package

--- 2021-08-13 paper/paper.tex
@@
 The design of \trixi was guided by extensibility, ease of use, and efficiency
-(in this particular order). Thus, it is likely possible to make (some parts) even
+(in this particular order). Thus, it is likely possible to make some parts even
 faster at the cost of simplicity and readability. Nevertheless, \trixi is already quite
@@
 \trixi do not need to compile anything or learn parameter file syntax. Instead, they can
-just load the library and write their simulation setups in Julia. Scripts that describe
+just load the package and write their simulation setups in Julia. Scripts that describe
 a simulation setup for \trixi are referred to as ``elixirs''. As of version

### 2021-08-13 paper/paper.tex
Emphasize "batteries included"

@@
 for the compressible Euler equations,
-to make it easier to use. Nevertheless, users are not restricted
+to make it easy to use out of the box. Nevertheless, users are not restricted
 to the physics models bundled in \trixi. To set up a new type of \lstinline{equations},

### 2021-08-13 paper/paper.tex
Physics is singular
Remove "traditional" to be slightly less judgmental

@@
 design with just-ahead-of-time compilation, these fluxes can be inlined into the
-library functions of \trixi. Thus, the physics can be separated completely from
-the \lstinline{solver} but remain computationally efficient. In particular, this allows
+library functions of \trixi. Thus, the physics is completely separated from
+the \lstinline{solver} but remains computationally efficient. In particular, this allows
 a user to reuse the same numerical fluxes for discontinuous Galerkin methods, finite
@@
 
-In traditional monolithic code bases, it is often necessary to implement a new
+In monolithic code bases, it is often necessary to implement a new
 feature for all combinations of possible (compile time) options. This effort

### 2021-08-13 paper/paper.tex
Avoid two uses of "select" in one sentence

@@
 discontinuous Galerkin methods on simplex elements via StartUpDG.jl\footnote{\url{https://github.com/jlchan/StartUpDG.jl}} in \trixi.
-This is feasible because the modular design of \trixi gives users the flexibility to select
+This is feasible because the modular design of \trixi gives users the flexibility to pick
 a subset of the available meshes/methods, selected via multiple dispatch,

### 2021-08-13 paper/paper.tex
Remove unnecessary parens

@@
 a subset of the available meshes/methods, selected via multiple dispatch,
-to test their newly implemented feature(s).
+to test their newly implemented features.
 

### 2021-08-13 paper/paper.tex
"the" -> "those" for emphasis

@@
 can make it difficult to experiment with new ideas. In contrast, Julia's dynamic
-nature, multiple dispatch, and just-ahead-of-time compilation allow us to implement only the features
+nature, multiple dispatch, and just-ahead-of-time compilation allow us to implement only those features
 that are strictly necessary. For example, if a user wants to simulate a new

### 2021-08-13 paper/paper.tex
"features" -> "capabilities" to reduce occurrences of "feature"

@@
 to achieve entropy conservation/dissipation or kinetic energy preservation that are used in \trixi.
-At the same time, both codes have features the other does not, \eg,
+At the same time, both codes have capabilities the other does not, \eg,
 support for parabolic equations and multi-node parallelism in FLUXO,

### 2021-08-13 paper/paper.tex
"final time t" -> "final time T" for consistency

@@
 by a combination of sine and cosine functions~\cite{chan2019efficient}. To integrate up to the
-final time $t=1.0$, both codes use the five-stage, fourth-order low-storage explicit
+final time $T=1.0$, both codes use the five-stage, fourth-order low-storage explicit
 Runge-Kutta method of Carpenter and Kennedy~\cite{CarpenterKennedy1994}.

### 2021-08-13 paper/paper.tex
run -> obtained
"[3, 15]" -> "from 3 to 15" to avoid confusion with citation notation

@@
 FLUXO was compiled using the Intel compiler suite v18.0.3 and \trixi used Julia v1.6.1.
-The serial performance results were run on a single core.
+The serial performance results were obtained on a single core.
 We ran each configuration of a particular physical setup, polynomial degree, and solver type
@@
 To assess the performance of the two codes, we vary the polynomial degree
-(\lstinline{polydeg}) in the range $[3,15]$.
+(\lstinline{polydeg}) from 3 to 15.
 This corresponds to considering DG approximations of (increasing) spatial accuracy, from fourth up to sixteenth order.

### 2021-08-13 paper/paper.tex
Remove parens as sentence makes no sense without
Introduce MHD abbreviation also in the main text (before it was only introduced in the feature list, where it is also required but might be overlooked)

@@
 a manufactured solution for the 3D compressible Euler equations and Alfv\'{e}n wave propagation for
-the 3D ideal MHD equations~\cite{gassner2009,altmann2012}. These initial conditions, available in FLUXO and
+the 3D ideal magnetohydrodynamics (MHD) equations~\cite{gassner2009,altmann2012}. These initial conditions, available in FLUXO and
 \trixi, are typically used for the purpose of demonstrating the high-order accuracy and convergence properties
@@
 (\lstinline{polydeg}) from 3 to 15.
-This corresponds to considering DG approximations of (increasing) spatial accuracy, from fourth up to sixteenth order.
+This corresponds to considering DG approximations of increasing spatial accuracy, from fourth up to sixteenth order.
 The metric we use to analyze and compare FLUXO and \trixi simulations is a performance index (PID) that

### 2021-08-13 paper/paper.tex
Move statement about compiler speeds to the appropriate section
Add plural "s" for better reading flow
Simplify statement

@@
 for each code five times, taking the smallest value from each for the comparison results.
+We ran the same PID tests in FLUXO compiled with GCC v6.4.0 but found that
+the version compiled using Intel was between \SI{5}{\percent} to \SI{23}{\percent} faster.
+Thus, only results from the Intel compiler are included.
 
@@
 and more than 1.5x faster for the ideal MHD equations than the Fortran code FLUXO.
-We ran the same PID tests in FLUXO compiled with GCC v6.4.0 but found that
-the version compiled using Intel was between \SI{5}{\percent} to \SI{23}{\percent} faster.
-Thus, only results from the Intel compiler are included.
 It is also possible to see the additional computational effort for the entropy-conservative flux
@@
 Further, we expect to be able to achieve similar performance with either language
-by spending enough time and effort to optimize the respective code.
+by spending enough time and effort to optimize the respective codes.
 \trixi owes its performance optimizations in part to the code introspection and
@@
 without need for curvilinear coordinates, the Cartesian \lstinline{TreeMesh} can
-be used. Depending on the relative computational effort of the discretization
-(such as the polynomial degree and the choice of the volume terms), the
+be used. Depending on the particular choice of discretization methods, the
 Cartesian \lstinline{TreeMesh} can be \SI{10}{\percent} to \SI{25}{\percent}

### 2021-08-13 paper/paper.tex
Remove superfluous "often"
Introduce "Pkg" as the package manager name

@@
 
-Many scientists implementing numerical methods often have no formal training
+Many scientists implementing numerical methods have no formal training
 in software development. Julia and its ecosystem support these
@@
 
-In addition, the package manager fosters a unified form of semantic versioning
+In addition, the Julia package manager \lstinline{Pkg} fosters a unified form of semantic versioning
 across the Julia ecosystem, which lays the basis for the following three
@@
 
-The package manager makes it easy to reproduce the exact runtime environment, including
+The package manager \lstinline{Pkg} makes it easy to reproduce the exact runtime environment, including
 binary dependencies, used to generate numerical results. We have used this feature for all of our papers based on

### 2021-08-13 paper/paper.tex
Avoid potential confusion of "this one" referring to reference [45]
Refer to Sec 4.2.2

@@
 binary dependencies, used to generate numerical results. We have used this feature for all of our papers based on
-\trixi \cite{schlottkelakemper2021purely, ranocha2021preventing}, including this
-one \cite{ranocha2021adaptiveRepro}.
+\trixi \cite{schlottkelakemper2021purely, ranocha2021preventing}, including the present manuscript
+\cite{ranocha2021adaptiveRepro}.
 This facilitates code sharing and reproducible research in computational science,
@@
 availability of HDF5.jl\footnote{\url{https://github.com/JuliaIO/HDF5.jl}}.
-There is still work to be done how to improve the experience in combination
+There is still work to be done on how to improve the experience in combination
 with MPI on HPC clusters, but some promising
-approaches exist \cite{byrne2021mpi}.
+approaches exist (see Section~\ref{sec:distributed-memory}).
 
@@
 \subsubsection{Simulations with distributed-memory parallelism}
+\label{sec:distributed-memory}
 

### 2021-08-13 paper/paper.tex
Remove erroneous plural "s"
Fix typo

@@
 session and plot the results on a notebook can take between 30 seconds and a
-minute. Having compiled the code, the second simulation and plot takes less than
+minute. Having compiled the code, the second simulation and plot take less than
 0.05 seconds. Thus, switching to Julia requires adapting the workflow compared
@@
 in the standard library, its performance and scalability are currently limited \cite{byrne2021mpi}.
-As an alternatively, the package MPI.jl utilizes the Message Passing Interface (MPI) for data
+As an alternative, the package MPI.jl utilizes the Message Passing Interface (MPI) for data
 exchange \cite{byrne2021mpi}. It provides an API similar to MPI's C interface and allows to either

### 2021-08-13 paper/paper.tex
Add reference to Spack and MPI.jl paper

@@
 community to work around this limitation by using the
-\texttt{WI4MPI}\footnote{\url{https://github.com/cea-hpc/wi4mpi}} wrapper package.
+WI4MPI\footnote{\url{https://github.com/cea-hpc/wi4mpi}} wrapper package or the
+Spack package manager \cite{gamblin2015spack,byrne2021mpi}.
 

### 2021-08-13 paper/paper.tex
Final fixes by Michael (#18)
* Add "the" for improved reading flow

* Improve word flow

* Shorten introduction

* Remove unused abbreviations

* Improve reading flow

* Remove parens since sentence makes no sense without them

* library -> package

* Emphasize "batteries included"

* Physics is singular

* Remove "traditional" to be slightly less judgmental

* Avoid two uses of "select" in one sentence

* Remove unnecessary parens

* "the" -> "those" for emphasis

* "features" -> "capabilities" to reduce occurrences of "feature"

* "final time t" -> "final time T" for consistency

* run -> ob

@@
 The physical system is specified by the fluxes $f^j$ and the source term $s$.
-In addition, suitable initial and boundary conditions are required.
-\trixi also handles non-conservative PDE terms such as the shallow water equations or the magnetohydrodynamics equations
+In addition, suitable initial and boundary conditions (ICs, BCs) are required.
+\trixi also handles non-conservative PDE terms as in the shallow water equations or magnetohydrodynamics equations
 with divergence cleaning.

### 2021-08-13 paper/paper.tex
phrase not too uncommon sounds awkward

@@
 to change initial/boundary conditions
-is not too uncommon, two-point numerical fluxes are also just functions with a
+is quite common, two-point numerical fluxes are also just functions with a
 specified signature. Thus, users can implement numerical fluxes in their

### 2021-08-13 paper/paper.tex
restructure final sentence of Sectino 2.3.4. One should not begin a sentence with the conjunction And

@@
 difference methods, and variants of finite volume methods.
-And even if a new feature requires modifications to methods provided by \trixi: Due to favoring
+Due to favoring
 simplicity over excessively generic code, it remains comparatively straightforward to modify
-existing implementations.
+existing \trixi implementations if a new feature requires modifications to methods.
 

### 2021-08-13 paper/paper.tex
small edits to Section 3.1

@@
 the 3D ideal magnetohydrodynamics (MHD) equations~\cite{gassner2009,altmann2012}. These initial conditions, available in FLUXO and
-\trixi, are typically used for the purpose of demonstrating the high-order accuracy and convergence properties
+\trixi, are typically used to demonstrate the high-order accuracy and convergence properties
 of the frameworks.
@@
 \end{equation}
-where $5$ is the number of Runge-Kutta stages per time step for the selected time-stepping method. In other words, the PID acts
-as a metric for the run time required for the right hand side (RHS) evaluation of each DOF.
+where $5$ is the number of Runge-Kutta stages per time step for the selected time-stepping method. In other words, the PID
+measures the run time required for the right hand side (RHS) evaluation of each DOF.
 

### 2021-08-13 paper/paper.tex
change for example to e.g. in Section 4

@@
 Julia was designed from scratch to be suitable for numerical computing
-\cite{bezanson2017julia}. Its usefulness has been demonstrated for example for
+\cite{bezanson2017julia}. Its usefulness has been demonstrated, \eg, for
 GPU programming \cite{besard2018juliagpu, omlin2020solving},

### 2021-08-13 paper/paper.tex
split one sentence into two in Section 4.1.1

@@
 Julia is not generically slower than traditional high-performance programming
-languages such as C, C++, and Fortran. A minor exception is due to
+languages like C, C++, and Fortran. A minor exception is due to
 Julia being built on top of LLVM, which does not perform the same optimizations as other
@@
 not use fused multiply-add (FMA) instructions by default, while GCC and Intel
-compilers do, and thus one needs to opt-in to these optimizations explicitly.
-Relying on Julia's code manipulation techniques, this can be achieved by using the macro
+compilers do. Thus, one must opt-in to these optimizations explicitly.
+Relying on Julia's code manipulation techniques, this can be achieved with the macro
 \lstinline{@muladd}\footnote{\url{https://github.com/SciML/MuladdMacro.jl}}.

### 2021-08-13 paper/paper.tex
slight change to subsection name of 4.1.4
small edits to the MPI Section 4.2.2

@@
 
-\subsubsection{External libraries can be integrated relatively easily}
+\subsubsection{External libraries can be integrated with relative ease}
 \label{sec:external-libraries}
@@
 As an alternative, the package MPI.jl utilizes the Message Passing Interface (MPI) for data
-exchange \cite{byrne2021mpi}. It provides an API similar to MPI's C interface and allows to either
+exchange \cite{byrne2021mpi}. It provides an API similar to MPI's C interface and allows one to either
 wrap MPI binaries installed through Julia's package manager or to make use of an existing MPI
@@
 
-However, there are also some practical challenges associated with running MPI-based parallel
-simulations with Julia. For starters, the issue of compilation times as described in the previous
-section is exacerbated, as MPI runs are by default not interactive and thus compilation results cannot be cached between MPI sessions. Since this
-requires to recompile everything for each subsequent execution of an MPI-parallel Julia program,
-this is particularly annoying during development. Partial relief can be obtained by using
-\texttt{tmpi}\footnote{\url{https://github.com/Azrael3000/tmpi}}, which allows to simultaneously
+However, some practical challenges associated with running MPI-based parallel
+simulations with Julia remain. For starters, the issue of compilation times, described in the previous
+section, is exacerbated as MPI runs are by default not interactive and compilation results cannot be cached between MPI sessions. As such, this
+requires code to recompile for each subsequent execution of an MPI-parallel Julia program,
+which is particularly annoying during development. Partial relief can be obtained by using
+\texttt{tmpi}\footnote{\url{https://github.com/Azrael3000/tmpi}} that allows the user to simultaneously
 interact with multiple interactive Julia REPL sessions via the terminal multiplexer \texttt{tmux}.
@@
 Unix-like operating systems. The use of PackageCompiler.jl to reduce compilation times as much as
-possible becomes thus mandatory. Another issue encountered when running Julia with many parallel
-processes is that Julia operates with a lot of small files during startup and precompilation,
+possible thus becomes mandatory. Another issue encountered when running Julia with many parallel
+processes is that Julia operates with many small files during startup and precompilation,
 putting heavy pressure on the parallel file system. This can be partially alleviated by doing
-precompilation serially, and there are tools like
+precompilation in serial, with tools like
 \texttt{Spindle}\footnote{\url{https://github.com/hpc/Spindle}}, which help to solve similar

### 2021-08-13 paper/paper.tex
updated Spindle sentence

@@
 putting heavy pressure on the parallel file system. This can be partially alleviated by doing
-precompilation in serial, with tools like
-\texttt{Spindle}\footnote{\url{https://github.com/hpc/Spindle}}, which help to solve similar
+precompilation in serial. There are also tools like
+\texttt{Spindle}\footnote{\url{https://github.com/hpc/Spindle}} that help solve similar
 problems that occur when running Python programs in parallel \cite{frings2013parallelloading}.

### 2021-08-24 paper/paper.tex
avoid using amcros in captions and section titles

@@
 
-\section{Capabilities and design of \trixi}
+\section{Capabilities and design of Trixi.jl}
 \label{sec:design-of-trixi}
@@
 
-\subsection{Main features of \trixi}
+\subsection{Main features of Trixi.jl}
 
@@
 
-\subsubsection{\trixi is a library}
+\subsubsection{Trixi.jl is a library}
 
@@
            different DG discretizations of the 3D compressible Euler equations
-           in FLUXO (Fortran) and \trixi (Julia).}
+           in FLUXO (Fortran) and Trixi.jl (Julia).}
   \label{fig:PID-Euler}

### 2022-01-05 paper/paper.tex
Allow using pre-generated PDFs from pgfplots
mention number of DOFs
additional performance reference

@@
 
+% Allow external generation of images to speed up compilation
+\usepackage{tikz}
+\usetikzlibrary{external}
+\tikzexternalize[prefix=tikz/]
+
+% Work around the fact that openjournal's whedon does not provide the pgfplots package
+% Xref: https://github.com/JuliaCon/proceedings-review/issues/76#issuecomment-913454361
+\newif\ifregenerateplots
+% \regenerateplotstrue  % Enable this if plots need to be regenerated
+\regenerateplotsfalse % Enable this to build PDF with whedon
+
+\ifregenerateplots
 % Plotting in tikz images
@@
 \pgfplotsset{compat=1.16}
-% Allow external generation of images to speed up compilation
-% \usetikzlibrary{external}
-% \tikzexternalize[prefix=../figures/tikz/]
+\fi
 
@@
 For the problem setup we consider a periodic box of the domain $[-1,1]^3$ with four elements in each
-spatial direction. To make the mesh curvilinear, the interior of the box is heavily warped
+spatial direction. This results in $64 (\text{\lstinline{polydeg}}+1)^3$ degrees
+of freedom (DG nodes) when polynomials of degree \lstinline{polydeg} are used.
+To make the mesh curvilinear, the interior of the box is heavily warped
 by a combination of sine and cosine functions~\cite{chan2019efficient}. To integrate up to the
@@
   \begin{subfigure}{\linewidth}
+    \tikzsetnextfilename{runtime_euler_absolute}
     \begin{tikzpicture}[
@@
   \begin{subfigure}{\linewidth}
+    \tikzsetnextfilename{runtime_euler_relative}
     \begin{tikzpicture}[
@@
   \begin{subfigure}{\linewidth}
+    \tikzsetnextfilename{runtime_mhd_absolute}
     \begin{tikzpicture}[
@@
   \begin{subfigure}{\linewidth}
+    \tikzsetnextfilename{runtime_mhd_relative}
     \begin{tikzpicture}[
@@
 expensive in
-terms of computational cost, and require special care to optimize their performance~\cite{hendrik_blog}.
+terms of computational cost, and require special care to optimize their performance~\cite{hendrik_blog,ranocha2021efficient}.
 We reiterate that both codes implement the same numerical nodal DG methods on curvilinear meshes used in these tests.
@@
 example, we do not claim that Julia is generally faster than Fortran, C, or C++.
-Instead, we would like to emphasize that Julia code can be \emph{at least
-as fast} as code written in these traditional scientific computing languages.
+Instead, we would like to emphasize that well-written Julia code can be \emph{at least
+as fast} as code written in these traditional scientific computing languages,
+as demonstrated also by several microbenchmarks \cite{bezanson2017julia}.
 Further, we expect to be able to achieve similar performance with either language

### 2022-01-05 paper/paper.tex
add some missing hyphens

@@
 main loop. In particular, various tasks such as input/output operations, adaptive
-mesh refinement (AMR), and positivity preserving limiting are implemented using
+mesh refinement (AMR), and positivity-preserving limiting are implemented using
 callbacks. In case full control over the time step loop is required, e.g., to experiment with new
@@
 to implement anything for curvilinear coordinates or shock-capturing approaches.
-In addition, it allows a user to extend \trixi step by step with new capabilities.
-For example, there are ongoing efforts to incorporate summation by parts finite
+In addition, it allows a user to extend \trixi step-by-step with new capabilities.
+For example, there are ongoing efforts to incorporate summation-by-parts finite
 difference methods via SummationByPartsOperators.jl \cite{ranocha2021sbp} and

### 2022-01-05 paper/paper.tex
visualization of IC and mesh for PID runs

@@
 of freedom (DG nodes) when polynomials of degree \lstinline{polydeg} are used.
-To make the mesh curvilinear, the interior of the box is heavily warped
-by a combination of sine and cosine functions~\cite{chan2019efficient}. To integrate up to the
+To make the mesh curvilinear, the interior of the box is heavily warped by
+a mapping adapted from \cite{chan2019efficient}.
+Specifically, we map Cartesian reference coordinates $(\xi, \eta, \zeta)$ to
+physical coordinates $(x, y, z)$ via the transformation
+\begin{equation}
+\begin{aligned}
+  y &= \eta  + 0.15 \bigl( \cos(1.5 \pi \xi) \cos(0.5 \pi \eta) \cos(0.5 \pi \zeta) \bigr),
+  \\
+  x &= \xi   + 0.15 \bigl( \cos(0.5 \pi \xi) \cos(2 \pi y) \cos(0.5 \pi \zeta) \bigr),
+  \\
+  z &= \zeta + 0.15 \bigl( \cos(0.5 \pi x) \cos(\pi y) \cos(0.5 \pi \zeta) \bigr).
+\end{aligned}
+\end{equation}
+To integrate up to the
 final time $T=1.0$, both codes use the five-stage, fourth-order low-storage explicit
@@
 of the frameworks.
+The curved mesh and the initial density for the compressible Euler problem are
+visualized in Figure~\ref{fig:pid_euler_mesh}.
+
+\begin{figure}[!h]
+  \includegraphics[width=\linewidth]{../figures/pid_euler_mesh}
+  \caption{Initial density of the compressible Euler problem on the curved mesh
+           used for the performance benchmarks.}
+  \label{fig:pid_euler_mesh}
+\end{figure}
 

### 2022-01-05 pid_runs/elixir_euler_plot.jl

@@ -1,0 +1,85 @@
+
+using OrdinaryDiffEq
+using Trixi
+
+###############################################################################
+# semidiscretization of the compressible Euler equations
+
+equations = CompressibleEulerEquations3D(1.4)
+
+initial_condition = initial_condition_convergence_test
+source_term=source_terms_convergence_test
+
+
+###############################################################################
+# Get the DG approximation space
+
+# For ECKEP flux split form
+#volume_flux = flux_ranocha
+#solver = DGSEM(polydeg=3, surface_flux=flux_ranocha,
+#               volume_integral=VolumeIntegralFluxDifferencing(volume_flux))
+
+# For standard DG weak form + HLL
+solver = DGSEM(polydeg=3, surface_flux=flux_hll,
+               volume_integral=VolumeIntegralWeakForm())
+
+###############################################################################
+# Get the curved quad mesh from a file
+
+# Mapping as described in https://arxiv.org/abs/2012.12040
+function mapping(xi, eta, zeta)
+  eta =
+
+  y = eta + 0.15 * (cos(1.5 * pi * xi) * cos(0.5 * pi * eta) * cos(0.5 * pi * zeta))
+
+  x = xi + 0.15 * (cos(0.5 * pi * xi) * cos(2 * pi * y) * cos(0.5 * pi * zeta))
+
+  z = zeta + 0.15 * (cos(0.5 * pi * x) * cos(pi * y) * cos(0.5 * pi * zeta))
+
+  return SVector(x, y, z)
+end
+
+cells_per_dimension = (4, 4, 4)
+mesh = StructuredMesh(cells_per_dimension, mapping)
+
+###############################################################################
+# create the semi discretization object
+
+semi = SemidiscretizationHyperbolic(mesh, equations, initial_condition, solver,
+                                    source_terms=source_term)
+
+###############################################################################
+# ODE solvers, callbacks etc.
+
+tspan = (0.0, 0.0)
+ode = semidiscretize(semi, tspan)
+
+summary_callback = SummaryCallback()
+
+analysis_interval = 100
+analysis_callback = AnalysisCallback(semi, interval=analysis_interval)
+
+save_solution = SaveSolutionCallback(interval=100,
+                                     save_initial_solution=true,
+                                     save_final_solution=false)
+
+stepsize_callback = StepsizeCallback(cfl=0.5)
+
+callbacks = CallbackSet(summary_callback,
+                        analysis_callback,
+                        save_solution,
+                        stepsize_callback)
+
[…]

### 2022-01-05 paper/paper.tex
comment on parallelization
change caption of mesh plot to indicate that it is a slice

@@
 
+Currently, \trixi provides shared-memory parallelization via multithreading.
+Distributed-memory parallelization with MPI is available for some mesh types
+but still work in progress.
+
 Some of the main features of \trixi are demonstrated in the following figures.
@@
 \begin{figure}[!h]
+\vspace{-0.5em}
   \begin{subfigure}{0.53\linewidth}
@@
   \includegraphics[width=\linewidth]{../figures/pid_euler_mesh}
-  \caption{Initial density of the compressible Euler problem on the curved mesh
-           used for the performance benchmarks.}
+  \caption{Initial density of the compressible Euler problem on a slice of the
+           periodic, curved mesh used for the performance benchmarks.}
   \label{fig:pid_euler_mesh}

### 2022-01-05 paper/paper.tex
Revisions based on the review by John F. Gibson (#21)
* mention number of DOFs

* additional performance reference

* add some missing hyphens

* visualization of IC and mesh for PID runs

* comment on parallelization

* change caption of mesh plot to indicate that it is a slice

* suggestions from Andrew

Co-authored-by: Andrew Winters <andrew.ross.winters@liu.se>

Co-authored-by: Andrew Winters <andrew.ross.winters@liu.se>

@@
 Currently, \trixi provides shared-memory parallelization via multithreading.
-Distributed-memory parallelization with MPI is available for some mesh types
-but still work in progress.
+Distributed-memory parallelization with MPI is available for some mesh types;
+increasing the capability of this feature is subject to ongoing work.
 
@@
 spatial direction. This results in $64 (\text{\lstinline{polydeg}}+1)^3$ degrees
-of freedom (DG nodes) when polynomials of degree \lstinline{polydeg} are used.
+of freedom (DG nodes) per equation when polynomials of degree \lstinline{polydeg} are used.
 To make the mesh curvilinear, the interior of the box is heavily warped by

### 2022-01-05 paper/paper.tex
adjust figures on p. 2
more details on Cartesian reference coordinates
reword parallelization feature
Co-authored-by: Michael Schlottke-Lakemper <michael@sloede.com>

@@
 Currently, \trixi provides shared-memory parallelization via multithreading.
-Distributed-memory parallelization with MPI is available for some mesh types;
-increasing the capability of this feature is subject to ongoing work.
+Initial parallelization with MPI is available for some mesh types,
+but full support for distributed-memory parallelism is subject to ongoing work.
 
@@
   \label{fig:jet}
+\vspace{-0.5em}
 \end{figure}
@@
 \begin{figure}[!h]
+\vspace{-0.5em}
   \begin{subfigure}{0.475\linewidth}
@@
 a mapping adapted from \cite{chan2019efficient}.
-Specifically, we map Cartesian reference coordinates $(\xi, \eta, \zeta)$ to
-physical coordinates $(x, y, z)$ via the transformation
+Specifically, we map Cartesian reference coordinates $(\xi, \eta, \zeta) \in [-1, 1]^3$
+to physical coordinates $(x, y, z)$ via the transformation
 \begin{equation}

### 2022-01-07 paper/paper.tex
clarify nonconservative terms

@@
 In addition, suitable initial and boundary conditions (ICs, BCs) are required.
-\trixi also handles non-conservative PDE terms as in the shallow water equations or magnetohydrodynamics equations
-with divergence cleaning.
+\trixi also handles non-conservative PDE terms as in the shallow water equations
+or magnetohydrodynamics equations with divergence cleaning, where source terms
+can depend on derivatives of~$u$.
 

### 2022-01-11 paper/paper.tex
layout adjustment
try to adapt layout to work for the old LaTeX version of whedon
Revert "try to adapt layout to work for the old LaTeX version of whedon"
This reverts commit d40d64e194b3c206fe96dbe0230a474c69aa2035.
some comments on performance difference

@@
 \begin{figure}[!h]
-\vspace{-0.5em}
+\vspace{-0.7em}
   \begin{subfigure}{0.53\linewidth}
@@
   \label{fig:jet}
-\vspace{-0.5em}
+\vspace{-0.7em}
 \end{figure}
@@
   \label{fig:kelvin_helmholtzs}
+\vspace{-0.7em}
 \end{figure}
@@
 \begin{figure}[!h]
-\vspace{-0.5em}
+\vspace{-0.7em}
   \begin{subfigure}{0.475\linewidth}
@@
 profiling tools available in Julia. Similar tools used to optimize the performance in other
-languages are often not as easy to use as their Julia counterparts.
+languages are often not as easy to use as their Julia counterparts. Therefore,
+we developed some additional performance improvements in \trixi and ported them
+to FLUXO later \cite{ranocha2021efficient}. Moreover, \trixi uses slightly
+different implementations of computationally intensive kernels and stores a bit
+less information in RAM, recomputing some terms instead. Additionally, the
+manufactured solution of the compressible Euler equations uses source terms
+containing sine and cosine terms, which are computed together by \lstinline{sincos}
+in \trixi but require individual calls to \lstinline{sin} and \lstinline{cos} in
+FLUXO.
 

### 2022-01-11 paper/paper.tex
reword FMA part
make figures a bit smaller to keep page limit
link to Base muladd

@@
 \vspace{-0.7em}
-  \begin{subfigure}{0.53\linewidth}
+  \begin{subfigure}{0.52\linewidth}
     \includegraphics[width=\textwidth]{../figures/jet_density}
@@
   \hspace*{\fill}
-  \begin{subfigure}{0.47\linewidth}
+  \begin{subfigure}{0.46\linewidth}
     \includegraphics[width=\textwidth]{../figures/jet_mesh}
@@
 \begin{figure}[!h]
-  \begin{subfigure}{0.53\linewidth}
+  \begin{subfigure}{0.52\linewidth}
     \includegraphics[width=\textwidth]{../figures/kelvin_helmholtz_density_t2}
@@
   \hspace*{\fill}
-  \begin{subfigure}{0.47\linewidth}
+  \begin{subfigure}{0.46\linewidth}
     \includegraphics[width=\textwidth]{../figures/kelvin_helmholtz_mesh_t2}
@@
   \\
-  \begin{subfigure}{0.53\linewidth}
+  \begin{subfigure}{0.52\linewidth}
     \includegraphics[width=\textwidth]{../figures/kelvin_helmholtz_density_t3}
@@
   \hspace*{\fill}
-  \begin{subfigure}{0.47\linewidth}
+  \begin{subfigure}{0.46\linewidth}
     \includegraphics[width=\textwidth]{../figures/kelvin_helmholtz_mesh_t3}
@@
 \vspace{-0.7em}
-  \begin{subfigure}{0.475\linewidth}
+  \begin{subfigure}{0.46\linewidth}
     \includegraphics[width=\textwidth]{../figures/ginger_time_00}
@@
   \hspace*{\fill}
-  \begin{subfigure}{0.475\linewidth}
+  \begin{subfigure}{0.46\linewidth}
     \includegraphics[width=\textwidth]{../figures/ginger_time_08}
@@
   \\
-  \begin{subfigure}{0.475\linewidth}
+  \begin{subfigure}{0.46\linewidth}
     \includegraphics[width=\textwidth]{../figures/ginger_time_16}
@@
   \hspace*{\fill}
-  \begin{subfigure}{0.475\linewidth}
+  \begin{subfigure}{0.46\linewidth}
     \includegraphics[width=\textwidth]{../figures/ginger_mesh}
@@
 Julia being built on top of LLVM, which does not perform the same optimizations as other
-C/C++/Fortran compilers. In particular, LLVM does
-not use fused multiply-add (FMA) instructions by default, while GCC and Intel
+C/C++/Fortran compilers. In particular, most LLVM frontends (including Julia) do
+not enable fused multiply-add (FMA) contraction by default, while GCC and Intel
 compilers do. Thus, one must opt-in to these optimizations explicitly.
 Relying on Julia's code manipulation techniques, this can be achieved with the macro
-\lstinline{@muladd}\footnote{\url{https://github.com/SciML/MuladdMacro.jl}}.
+\lstinline{@muladd}\footnote{\url{https://github.com/SciML/MuladdMacro.jl}},
+which makes it convenient to use Julia's function \lstinline{muladd}\footnote{\url{https://docs.julialang.org/en/v1/base/math/\#Base.muladd}}.
 

### 2022-01-11 paper/paper.tex
rewrite FMA contraction part

@@
 Julia is not generically slower than traditional high-performance programming
-languages like C, C++, and Fortran. A minor exception is due to
-Julia being built on top of LLVM, which does not perform the same optimizations as other
-C/C++/Fortran compilers. In particular, most LLVM frontends (including Julia) do
-not enable fused multiply-add (FMA) contraction by default, while GCC and Intel
-compilers do. Thus, one must opt-in to these optimizations explicitly.
+languages like C, C++, and Fortran. A minor exception in the context of scientific
+computing relying on floating point operations is that Julia does not enable
+fused multiply-add (FMA) contraction by default, while GCC and Intel compilers
+do\footnote{An interesting coincidence is that not only Julia but most LLVM-based
+frontends tend not to enable FMA contraction by default.}.
+Thus, one must opt-in to these optimizations explicitly.
 Relying on Julia's code manipulation techniques, this can be achieved with the macro

### 2022-01-11 paper/paper.tex
improve FMA contraction part using suggestions of Simon Byrne

--- 2022-01-11 paper/paper.tex
@@
 Julia is not generically slower than traditional high-performance programming
-languages like C, C++, and Fortran. A minor exception is due to
-Julia being built on top of LLVM, which does not perform the same optimizations as other
-C/C++/Fortran compilers. In particular, LLVM does
-not use fused multiply-add (FMA) instructions by default, while GCC and Intel
-compilers do. Thus, one must opt-in to these optimizations explicitly.
-Relying on Julia's code manipulation techniques, this can be achieved with the macro
-\lstinline{@muladd}\footnote{\url{https://github.com/SciML/MuladdMacro.jl}}.
+languages like C, C++, and Fortran. A minor exception in the context of scientific
+computing relying on floating point operations is that Julia does not perform
+automatic fused multiply-add (FMA) contraction, that is replacing a multiplication
+followed by an addition with a single FMA instruction, as it is inconsistent
+with its strict floating point semantics. In other compilers, this optimization
+is controlled by a global compiler option (\lstinline{-ffp-contract} in GCC and
+Clang) and often enabled at reasonable optimization levels such as \lstinline{-O2}
+for GCC and Intel compilers\footnote{An interesting coincidence is that not only
+Julia but most LLVM-based frontends tend not to enable FMA contraction by default.}.
+In Julia, FMA contraction is controlled more locally: the
+\lstinline{muladd}\footnote{\url{https://docs.julialang.org/en/v1/base/math/\#Base.muladd}}
+intrinsic function allows the compiler to evaluate a multiply-add using the most
+efficient method. Relying on Julia's code manipulation techniques, the
+\lstinline{@muladd}\footnote{\url{https://github.com/SciML/MuladdMacro.jl}}
+macro makes it convenient to use the \lstinline{muladd} function by
+syntactically rewriting expressions to insert \lstinline{muladd} in appropriate
+locations.
 

### 2022-01-11 paper/paper.tex
tune figures to keep page limit
Apply suggestions from code review
Co-authored-by: Michael Schlottke-Lakemper <michael@sloede.com>

@@
 \begin{figure*}[htbp]
-  \includegraphics[width=\linewidth]{../figures/trixi_global_overview}
+\centering
+  \includegraphics[width=0.9\linewidth]{../figures/trixi_global_overview}
   \caption{Schematic overview of the basic components in \trixi and how they
@@
 \begin{figure}[!h]
-  \includegraphics[width=\linewidth]{../figures/pid_euler_mesh}
+\centering
+  \includegraphics[width=0.9\linewidth]{../figures/pid_euler_mesh}
   \caption{Initial density of the compressible Euler problem on a slice of the
@@
 to FLUXO later \cite{ranocha2021efficient}. Moreover, \trixi uses slightly
-different implementations of computationally intensive kernels and stores a bit
-less information in RAM, recomputing some terms instead. Additionally, the
+different implementations of computationally intensive kernels and stores marginally
+less information in memory, recomputing some terms instead. Additionally, the
 manufactured solution of the compressible Euler equations uses source terms
@@
 computing relying on floating point operations is that Julia does not perform
-automatic fused multiply-add (FMA) contraction, that is replacing a multiplication
+automatic fused multiply-add (FMA) contraction, that is, replacing a multiplication
 followed by an addition with a single FMA instruction, as it is inconsistent

### 2022-01-11 paper/paper.tex
move FMA discussion to the appendix

--- 2022-01-11 paper/paper.tex
@@
 languages like C, C++, and Fortran. A minor exception in the context of scientific
-computing relying on floating point operations is that Julia does not enable
-fused multiply-add (FMA) contraction by default, while GCC and Intel compilers
-do\footnote{An interesting coincidence is that not only Julia but most LLVM-based
-frontends tend not to enable FMA contraction by default.}.
-Thus, one must opt-in to these optimizations explicitly.
-Relying on Julia's code manipulation techniques, this can be achieved with the macro
-\lstinline{@muladd}\footnote{\url{https://github.com/SciML/MuladdMacro.jl}},
-which makes it convenient to use Julia's function \lstinline{muladd}\footnote{\url{https://docs.julialang.org/en/v1/base/math/\#Base.muladd}}.
+computing relying on floating point operations is that Julia does not perform
+automatic fused multiply-add (FMA) contraction, which can be remedied by
+explicitly using the
+\lstinline{muladd}\footnote{\url{https://docs.julialang.org/en/v1/base/math/\#Base.muladd}}
+function or the more convenient
+\lstinline{@muladd}\footnote{\url{https://github.com/SciML/MuladdMacro.jl}}
+macro (for more details see the appendix).
 
@@
 
+\appendix
+
+Julia does not perform automatic fused multiply-add (FMA) contraction, that is
+replacing a multiplication followed by an addition with a single FMA instruction,
+as it is inconsistent with its strict floating point semantics.
+In other compilers, this optimization is controlled by a global compiler option
+(\lstinline{-ffp-contract} in GCC and Clang) and often enabled at reasonable
+optimization levels such as \lstinline{-O2} for GCC and Intel compilers\footnote{%
+An interesting coincidence is that not only Julia but most LLVM-based frontends
+tend not to enable FMA contraction by default.}.
+In Julia, FMA contraction is controlled more locally: the
+\lstinline{muladd}%\footnote{\url{https://docs.julialang.org/en/v1/base/math/\#Base.muladd}}
+intrinsic function allows the compiler to evaluate a multiply-add using the most
+efficient method. Relying on Julia's code manipulation techniques, the
+\lstinline{@muladd}%\footnote{\url{https://github.com/SciML/MuladdMacro.jl}}
+macro makes it convenient to use the \lstinline{muladd} function by
+syntactically rewriting expressions to insert \lstinline{muladd} in appropriate
+locations.
+
 \end{document}

### 2022-01-11 paper/paper.tex

@@
 
-Julia does not perform automatic fused multiply-add (FMA) contraction, that is
+Julia does not perform automatic fused multiply-add (FMA) contraction, i.e.,
 replacing a multiplication followed by an addition with a single FMA instruction,
 as it is inconsistent with its strict floating point semantics.
-In other compilers, this optimization is controlled by a global compiler option
-(\lstinline{-ffp-contract} in GCC and Clang) and often enabled at reasonable
+In other compilers, this optimization can be controlled by a global compiler option
+(\lstinline{-ffp-contract} in GCC and Clang) and is often enabled at reasonable
 optimization levels such as \lstinline{-O2} for GCC and Intel compilers\footnote{%
@@
 In Julia, FMA contraction is controlled more locally: the
-\lstinline{muladd}%\footnote{\url{https://docs.julialang.org/en/v1/base/math/\#Base.muladd}}
+\lstinline{muladd} %\footnote{\url{https://docs.julialang.org/en/v1/base/math/\#Base.muladd}}
 intrinsic function allows the compiler to evaluate a multiply-add using the most
 efficient method. Relying on Julia's code manipulation techniques, the
-\lstinline{@muladd}%\footnote{\url{https://github.com/SciML/MuladdMacro.jl}}
+\lstinline{@muladd} %\footnote{\url{https://github.com/SciML/MuladdMacro.jl}}
 macro makes it convenient to use the \lstinline{muladd} function by

### 2022-01-14 paper/paper.tex
try H from package float
Revert "try H from package float"
This reverts commit 181570596d2fa048598a8e27c7abd9ac7566a427.
reorder texts and figures in the source
WIP
try sloppy
\captionsetup[subfigure]{font=footnotesize}
Revert "\captionsetup[subfigure]{font=footnotesize}"
This reverts commit e7b6579a2896b28f20d8ed721f0b4bbf15b9cc78.
shorten subcaptions of fig:pressure_waves

@@
   \label{fig:jet}
-\vspace{-0.7em}
+\vspace{-1.4em}
 \end{figure}
@@
 classical Kelvin-Helmholtz flow instability problem.
-\begin{figure}[!h]
+
+Figure~\ref{fig:pressure_waves} shows the approximation
+of acoustic perturbation equations~\cite{ewert2003acoustic} wave scattering on a
+curvilinear and unstructured domain. The 2D quadrilateral mesh used in the simulation
+(Figure~\ref{fig:ginger_mesh}) was generated with
+HOHQMesh.jl\xspace\footnote{\url{https://github.com/trixi-framework/HOHQMesh.jl}}.
+
+\begin{figure}[!ht]
   \begin{subfigure}{0.52\linewidth}
@@
 \end{figure}
-
-Figure~\ref{fig:pressure_waves} shows the approximation
-of acoustic perturbation equations~\cite{ewert2003acoustic} wave scattering on a
-curvilinear and unstructured domain. The 2D quadrilateral mesh used in the simulation
-(Figure~\ref{fig:ginger_mesh}) was generated with
-HOHQMesh.jl\xspace\footnote{\url{https://github.com/trixi-framework/HOHQMesh.jl}}.
-\begin{figure}[!h]
+%
+\begin{figure}[!ht]
 \vspace{-0.7em}
@@
     \includegraphics[width=\textwidth]{../figures/ginger_time_00}
-    \caption{Acoustic pressure at time $t = 0$.}
+    \caption{Acoustic pressure at $t = 0$.}
   \end{subfigure}%
@@
     \includegraphics[width=\textwidth]{../figures/ginger_time_08}
-    \caption{Acoustic pressure at time $t = 8$.}
+    \caption{Acoustic pressure at $t = 8$.}
   \end{subfigure}%
@@
     \includegraphics[width=\textwidth]{../figures/ginger_time_16}
-    \caption{Acoustic pressure at time $t = 16$.}
+    \caption{Acoustic pressure at $t = 16$.}
   \end{subfigure}%
@@
   \caption{Numerical solutions of pressure wave scattering for the acoustic
-                perturbation equations at three points in time as well as the unstructured,
-                curvilinear quadrilateral mesh.}
+                perturbation equations at three points in time $t$ as well as
+                the unstructured, curvilinear quadrilateral mesh.}
   \label{fig:pressure_waves}
+\vspace{-0.7em}
 \end{figure}
@@
 
-\begin{figure*}[htbp]
+\begin{figure*}[th]
 \centering
@@
   \label{fig:trixi_global_overview}
+\vspace{-0.7em}
 \end{figure*}
